---
title: "1_Data_Preparation"
---

This markdown file contains the code for filtering the downloaded raw datasets to the Greater London study area. The filtered data sets are saved as follows:

1.  `lsoa_london` : LSOA boundaries filtered for London only
2.  `greenspace_london_2ha` : green space polygons filtered for any intersecting the Greater London boundary, further filtered for sizes bigger than 2ha only (as informed by literature)
3.  `london_lsoa_pop_weighted_centroids.shp` : LSOA population weighted centroids filtered for Greater London only
4.  `gtfs_rail.zip` : converted rail timetable data into GTFS format.

## Load Packages

```{r, message=FALSE, warning=FALSE}
# Load packages
library(sf)
library(dplyr)
library(remotes)
# library(UK2GTFS)  # install and load in below
```

## 1. London LSOA Data

```{r}
# LSOA for whole Great Britian
lsoa_gb <- st_read("data/raw_data/Lower_layer_Super_Output_Areas_December_2021_Boundaries_EW_BFC_V10/LSOA_2021_EW_BFC_V10.shp")

# London Boundary 
london_boundary <- st_read("data/raw_data/gla/London_GLA_Boundary.shp")

# ensure both have the same CRS
london_boundary <- st_transform(london_boundary, st_crs(lsoa_gb))
```

```{r}
# centroids of LSOAs across GB
lsoa_centroids <- st_centroid(lsoa_gb)
```

```{r}
# centroids within the London boundary
lsoa_london <- lsoa_gb[st_within(lsoa_centroids, london_boundary, sparse = FALSE), ]
```

```{r}
# check number of LSOAs after clipped
nrow(lsoa_london) 

# check geom type
table(st_geometry_type(lsoa_london))
```

```{r}
# save as shapefile
st_write(lsoa_london, "data/lsoa_london/lsoa_london.shp")
```

## 2. London Green Space Polygons \> 2ha

```{r}
# Greenspace polygons
greenspace_gb <- st_read("data/raw_data/OS-Open-Greenspace-GB/GB_GreenspaceSite/GB_GreenspaceSite.shp") %>%
  st_transform(st_crs(london_boundary))
```

```{r}
# st_intersects to make sure green space polygons that intersects the London boundary is fully kept
greenspace_london <- greenspace_gb %>% 
    st_filter(london_boundary, .predicate = st_intersects)
```

```{r}
# remove polygon of sizes smaller than 2ha
greenspace_london_2ha <- greenspace_london %>%
  mutate(
    # area
    area_m2 = as.numeric(st_area(.)),
    # hectares 
    area_ha = area_m2 / 10000
  ) %>%
  # filter for hectare sizes of larger or equals to 2ha
  filter(area_ha >= 2)
```

```{r}
# save as shapefile
st_write(greenspace_london_2ha, "data/greenspace_london_2ha/greenspace_london_2ha.shp")
```

## 3. London LSOA Population Weighted Centroids

```{r}
# population weighted centroids
pop_weighted_centroids <- st_read("data/raw_data/LSOA_PopCentroids_EW_2021_V4_-7256415914866551835")
```

```{r}
# use london boundary to filter
# ensure consistent CRS
pop_weighted_centroids <- st_transform(pop_weighted_centroids,st_crs(london_boundary))

# filter pop centroids to London only
london_pop_weighted_centroids <- pop_weighted_centroids %>% 
  st_filter(london_boundary, .predicate = st_intersects)
```

```{r}
# save as shapefile
st_write(london_pop_weighted_centroids, "data/london_lsoa_pop_weighted_centroids/london_lsoa_pop_weighted_centroids.shp")
```

## 4. Rail GTFS

```{r}
# install UK2GTFS remotely from github
remotes::install_github("ITSleeds/UK2GTFS")

# Load the UK2GTFS package for converting to GTFS
library(UK2GTFS)
```

```{r}
# to take data from National Rail Portal
# save username and password
install.packages("usethis")
library(usethis)

usethis::edit_r_environ()
```

```{r}
# donload the data from the National Rail Data Portal and save as zip file
nrdp_timetable("data/raw_data/timetable.zip")

path_in <- "data/raw_data/timetable.zip"
# convert into GTFS
gtfs_rail <- atoc2gtfs(path_in = path_in,
                       ncores = 5)

# save as gtfs zip file for r5r 
gtfs_write(gtfs_rail,
           folder = "data/r5r",
           name = "gtfs_rail"
)
```
