---
title: "1_Data_Filter_London"
---

This markdown file contans the code for filtering the downloaded raw data sets to the Greater London study area. These are pre-filtered for further data preprocessing in Google Earth Engine. The filtered data sets are saved as follows:

1.  `lsoa_london` : LSOA boundaries filtered for London only
2.  `greenspace_london` : green space polygons filtered for any intersecting the Greater London boundary
3.  `greenspace_london_2ha` : green space polygons filtered for any intersecting the Greater London boundary, and sizes bigger than 2ha only (as informed by literature)

## Load Packages

```{r, message=FALSE, warning=FALSE}
# Load packages
library(sf)
library(dplyr)
library(tmap)
```

## 1. London LSOA Data 

```{r}
# LSOA for whole Great Britian
lsoa_gb <- st_read("data/raw_data/Lower_layer_Super_Output_Areas_December_2021_Boundaries_EW_BFC_V10/LSOA_2021_EW_BFC_V10.shp")

# London Boundary 
london_boundary <- st_read("data/raw_data/gla/London_GLA_Boundary.shp")

# Ensure both have the same CRS
london_boundary <- st_transform(london_boundary, st_crs(lsoa_gb))
```

```{r}
# Centroids of LSOAs across GB
lsoa_centroids <- st_centroid(lsoa_gb)
```

```{r}
# Centroids within the London boundary
lsoa_london <- lsoa_gb[st_within(lsoa_centroids, london_boundary, sparse = FALSE), ]
```

```{r}
# Check number of LSOAs after clipped
nrow(lsoa_london) 

# Check geom type
table(st_geometry_type(lsoa_london))
```

```{r}
qtm(lsoa_london)
```

```{r}
# Save as shapefile
st_write(lsoa_london, "data/lsoa_london/lsoa_london.shp")
```

## 2. London Green Space Polygons

```{r}
greenspace_gb <- st_read("data/OS-Open-Greenspace-GB/GB_GreenspaceSite/GB_GreenspaceSite.shp") %>%
  st_transform(st_crs(london_boundary))
```

```{r}
# st_intersects to make sure the whole park is kept despite overspilling the London boundary
greenspace_london <- greenspace_gb %>% 
    st_filter(london_boundary, .predicate = st_intersects)
```

```{r}
qtm(greenspace_london)
```

```{r}
# Save as shapefile
st_write(greenspace_london, "data/greenspace_london/greenspace_london.shp")
```

## 3. London Green Space Polygons \> 2ha 

```{r}
# Removed polygon sizes smaller than 2ha
greenspace_london_2ha <- greenspace_london %>%
  mutate(
    area_m2 = as.numeric(st_area(.)),
    area_ha = area_m2 / 10000
  ) %>%
  filter(area_ha >= 2)
```

```{r}
st_write(greenspace_london_2ha, "data/greenspace_london_2ha/greenspace_london_2ha.shp")
```
