---
title: "4_Cooling_Effective_UGS_Accessibility"
---

This markdown file contains the code for ......

```{r}
# remove objects from workspace and clear memory
rm(list = ls())    
gc()
```

## Load Packages

```{r, message=FALSE, warning=FALSE}
library(sf)
library(dplyr)
```

## 1. Park Cooling Quantification

### 1.1. Data Loading and Preparation

```{r}
# load greenspace buffer lst shapefile from GEE export
greenspace <- st_read("data/greenspace_buffer_lst/greenspace_buffer_lst.shp")
```

```{r}
# check column names
greenspace %>% colnames()
```

```{r}
# selected needed columns
greenspace_clean <- greenspace %>% 
  select(id, function_, area_ha, park_lst,
         # all of the 20 buffer rings in ascending order
         all_of(paste0("ring_", seq(30, 600, 30))),
         geometry)
```

```{r}
# check any NA values
greenspace_clean %>%
  st_drop_geometry() %>%
  summarise(across(everything(), ~sum(is.na(.)))) %>%
  t()
```

### 1.2. Cubic Polynomial Fitting Model and Park Cooling Intensity (PCI)

| Name                         | Formula                         |
|------------------------------|---------------------------------|
| Cublic Polynomial Fitting    | **T(x) = ax³ + bx² + cx + d**   |
| Park Cooling Intensity (PCI) | **PCI = L × T_L − ∫₀ᴸ T(x) dx** |

```{r}
# Ring distances in metres — x values for polynomial fitting
x_vals <- seq(30, 600, 30)
ring_cols <- paste0("ring_", seq(30, 600, 30))

results <- list()

# loop through each greenspace
for (i in 1:nrow(greenspace_clean)) {
  # for each greenspace
  park <- greenspace_clean[i, ]
  # take LST values for each ring
  y_vals <- as.numeric(unlist(park[, ring_cols] %>% st_drop_geometry(), use.names = FALSE))
  # take park LST values
  park_lst <- park$park_lst
  
  # Cubic Polynomial Fitting:
  # T(x) = ax3 + bx2 + cx + d
  poly_fit <- lm(y_vals ~ poly(x_vals, degree = 3, raw = TRUE))
  
  # Coefficients
  coefs <- coef(poly_fit)
  a <- coefs[4]
  b <- coefs[3]
  c <- coefs[2]
  d <- coefs[1]
  
  # R squared 
  # less error with higher r2 (https://www.nature.com/articles/s41599-023-02209-5)
  r2 <- summary(poly_fit)$r.squared
  
  # Find derivative T'(x) = 3ax2 + 2bx + c (for slope of temp distance curve)
  # so when T'(x) = 0 --> that is a inflection point where temp stop decreasing/ increasing
  disc <- (2*b)^2 - 4*(3*a)*c
  # if no inflection point, return to NA values (no valid cooling limit)
  if (is.na(disc) || disc < 0) {
    d_max <- NA
    T_L   <- NA
    PCI   <- NA
    dTmax <- NA
  } else {
    root1 <- (-(2*b) + sqrt(disc)) / (2 * 3*a)
    root2 <- (-(2*b) - sqrt(disc)) / (2 * 3*a)
    
    valid_roots <- c(root1, root2)
    # Keep only estimated turning points within your observed GEE ring range
    valid_roots <- valid_roots[valid_roots >= 30 & valid_roots <= 600]
    
    if (length(valid_roots) == 0) {
      d_max <- NA
      T_L   <- NA
      PCI   <- NA
      dTmax <- NA
    } else {
      
      # Check whether each turning point is a peak or a dip
      # T''(x) = 6ax + 2b
      second_derivative <- 6 * a * valid_roots + 2 * b
      
      # Keep only peaks
      peak_roots <- valid_roots[second_derivative < 0]
      
      if (length(peak_roots) == 0) {
        d_max <- NA
        T_L   <- NA
        PCI   <- NA
        dTmax <- NA
      } else {
        
        # Take the first peak aka the initial inflection point
        d_max <- min(peak_roots)
        
        # Calculate fitted LST at d_max
        T_L <- a * d_max^3 + b * d_max^2 + c * d_max + d
        
        # Integral of fitted T(x) from 0 to d_max
        integral <- (a/4 * d_max^4) +
          (b/3 * d_max^3) + 
          (c/2 * d_max^2) + 
          (d   * d_max)
        
        # PCI
        PCI <- (d_max * T_L) - integral
        
        # Maximum temperature difference
        dTmax <- T_L - park_lst
      }
    }
  }
  
  # store results
  results[[i]] <- data.frame(
    id    = park$id,
    r2    = round(r2, 3),
    d_max = round(d_max, 1),
    T_L   = round(T_L, 2),
    dTmax = round(dTmax, 2),
    PCI   = round(PCI, 1)
  )
}

# combine results of all parks
pci_results <- bind_rows(results)

# check
cat("Total parks processed:", nrow(pci_results), "\n")
cat("Parks with valid PCI:", sum(!is.na(pci_results$PCI)), "\n")
cat("Parks with no inflection point:", sum(is.na(pci_results$PCI)), "\n")

summary(pci_results$PCI)
summary(pci_results$r2)
```

```{r}
# join calculated PCI back to greenspace_clean
greenspace_pci <- greenspace_clean %>%
  left_join(pci_results, by = "id")
```

### 1.3. Identify Cooling Effective UGS

```{r}
# identify cooling effective parks
cooling_effective <- greenspace_pci %>%
  filter(
    !is.na(dTmax),
    !is.na(PCI),
    # filter temp difference > 0
    dTmax > 0,
    # filter accumulated cooling > 0
    PCI > 0,
    # 0.7 as a useful threshold(https://www.sciencedirect.com/science/article/pii/S0169743998002044) as it is said less error with higher r2 (https://www.nature.com/articles/s41599-023-02209-5)
    r2 >= 0.7
  )
```

Number of cooling effective UGS:

```{r}
nrow(cooling_effective)
```

## 2. Multimodal Rapid Realistic Routing (r5r)
