---
title: "4_Cooling_Effective_UGS_Accessibility"
---

This markdown file contains the full working code for ......

```{r}
# remove objects from R environment and clear memory
rm(list = ls())    
gc()
```

## Load Packages

```{r, message=FALSE, warning=FALSE}
library(sf)
library(dplyr)
library(ggplot2)
library(tidyr)
```

## 1. Park Cooling Quantification

### 1.1. Data Loading and Preparation

```{r}
# load greenspace buffer lst shapefile exported from GEE
greenspace <- st_read("data/greenspace_buffer_lst/greenspace_buffer_lst.shp")
```

```{r}
# check column names
greenspace %>% colnames()
```

```{r}
# select needed columns only
greenspace_clean <- greenspace %>% 
  select(id, function_, area_ha, park_lst,
         # all of the 20 buffer rings in ascending order
         all_of(paste0("ring_", seq(30, 900, 30))),
         geometry)
```

```{r}
# check any NA values
greenspace_clean %>%
  st_drop_geometry() %>%
  summarise(across(everything(), ~sum(is.na(.))))
```

### 1.2. Cubic Polynomial Curve and Park Cooling Intensity (PCI)

| Name | Formula | Literature Support |
|----|----|----|
| Cublic Polynomial Curve | T(x) = ax³ + bx² + cx + d | Du *et al.*, 2022; Cai *et al.*, 2023; Yang *et al.*, 2024 |
| Accumulated PCI | PCI = L × T_L − ∫₀ᴸ T(x) dx | Du *et al.*, 2022; Zhang *et al.*, 2024 |
| Normalised PCI | \\(PCI = \\frac{PCI\_{acc}}{L\_{max} \\times TD}\\) | Zhang *et al.*, 2024 |
| Park Cooling Gradient (PCG) | \\(PCG = \\frac{PCI\_{acc}}{L\_{max}}\\) | Yang *et al.*, 2024; Zhang *et al.*, 2024 |

```{r}
# Ring distances in metres — x values for polynomial fitting
x_vals <- seq(30, 900, 30)
ring_cols <- paste0("ring_", seq(30, 900, 30))

# create an empty list to store results for later
results <- list()

# for loop for each greenspace
for (i in 1:nrow(greenspace_clean)) {
  # for each greenspace
  park <- greenspace_clean[i, ]
  # LST values from buffer ring
  y_vals <- as.numeric(unlist(park[, ring_cols] %>% st_drop_geometry(), use.names = FALSE))
  # park LST values
  park_lst <- as.numeric(park$park_lst)
  
  # fit a cubic polynomial curve: T(x) = ax3 + bx2 + cx + d
  poly_fit <- lm(y_vals ~ poly(x_vals, degree = 3, raw = TRUE))
  
  # coefficients
  coefs <- coef(poly_fit)
  a <- coefs[4]
  b <- coefs[3]
  c <- coefs[2]
  d <- coefs[1]
  
  # calculate R squared 
  # it is said there is found less error with higher r2 (https://www.nature.com/articles/s41599-023-02209-5)
  r2 <- summary(poly_fit)$r.squared
  
  # inflection/ turning point of curve --> find where temp stop decreasing/ increasing
  # find derivative: T'(x) = 3ax2 + 2bx + c 
  disc <- (2*b)^2 - 4*(3*a)*c
  # parks with no valid inflection/ turning point, return as NA
  if (is.na(disc) || disc < 0) {
    Lmax           <- NA
    T_L             <- NA
    PCI_accumulated <- NA
    PCI             <- NA
    PCG             <- NA
    temp_diff           <- NA
  } else {
   
    # distances where temp stops increasing/ decreasing
    root1 <- (-(2*b) + sqrt(disc)) / (2 * 3*a)
    root2 <- (-(2*b) - sqrt(disc)) / (2 * 3*a)
    
    # only keep inflection points within ring range 30 - 900m
    valid_roots <- c(root1, root2)
    valid_roots <- valid_roots[valid_roots >= 30 & valid_roots <= 900]
    
    if (length(valid_roots) == 0) {
      Lmax           <- NA
      T_L             <- NA
      PCI_accumulated <- NA
      PCI             <- NA
      PCG             <- NA
      temp_diff           <- NA
    } else {
      
      # identify whether inflection point is at highest point of curve
      # T''(x) = 6ax + 2b
      second_derivative <- 6 * a * valid_roots + 2 * b
      
      # Keep only peaks
      peak_roots <- valid_roots[second_derivative < 0]
      # if it s not a peak, return as NA
      if (length(peak_roots) == 0) {
        Lmax           <- NA
        T_L             <- NA
        PCI_accumulated <- NA
        PCI             <- NA
        PCG             <- NA
        temp_diff           <- NA
      } else {
        
        # select the initial inflection point (first peak) as cooling distance
        Lmax <- min(peak_roots)
        
        # calculate fitted LST at the Lmax
        T_L <- a * Lmax^3 + b * Lmax^2 + c * Lmax + d
        
        # calculate integral of fitted T(x) from 0 to Lmax
        # aka summarise the temp pattern from park geom to the cooling distance
        integral <- (a/4 * Lmax^4) +
          (b/3 * Lmax^3) + 
          (c/2 * Lmax^2) + 
          (d   * Lmax)
        
        # calculate PCI accumulated
        PCI_accumulated <- (Lmax * T_L) - integral
        
        # PCI
        PCI <- PCI_accumulated / (Lmax * T_L)
        
        # calculate PCG (Park cooling gradient)
        # PCG = accumulated cooling area/ maximum cooling distance (https://www.sciencedirect.com/science/article/pii/S2210670723007205)
        PCG <- PCI_accumulated / Lmax
        
        # calculate max temperature difference
        temp_diff <- T_L - park_lst
      }
    }
  }
  
  # store results
  results[[i]] <- data.frame(
    id = park$id,                           # park ID
    r2 = round(r2, 3),                      # R squared, how well the curve fits the data
    Lmax = round(Lmax, 1),                # estimated cooling distance
    T_L = round(T_L, 2),                    # temp at Lmax
    temp_diff = round(temp_diff, 2),                # difference between surrounding temp and park temp
    PCI_accumulated = round(PCI_accumulated, 1),   # accumulated cooling area, old PCI formula
    PCI = round(PCI, 4),                    # normalised PCI
    PCG = round(PCG, 2)                     # park cooling gradient
  )
}

# combine results
pci_results <- bind_rows(results)

# check
cat("Total parks:", nrow(pci_results), "\n")
cat("Parks with valid PCI:", sum(!is.na(pci_results$PCI)), "\n")
cat("Parks with no inflection point:", sum(is.na(pci_results$PCI)), "\n")

summary(pci_results$PCI_accumulated)
summary(pci_results$PCI)
summary(pci_results$PCG)
summary(pci_results$r2)
```

```{r}
# join PCI results back to greenspace_clean
greenspace_pci <- greenspace_clean %>%
  left_join(pci_results, by = "id")
```

### 1.3. Identify Cooling Effective UGS

```{r}
# identify cooling effective parks
cooling_effective <- greenspace_pci %>%
  filter(
    # filter away all NA
    !is.na(temp_diff),
    !is.na(PCI),
    !is.na(PCG),
    # filter temp difference > 0
    temp_diff > 0,
    # filter normalised cooling > 0
    PCI > 0,
    # filter PCG > 0
    PCG > 0,
    # 0.7 as a useful threshold(https://www.sciencedirect.com/science/article/pii/S0169743998002044) as it is said less error with higher r2 (https://www.nature.com/articles/s41599-023-02209-5)
    r2 >= 0.7
  )
```

```{r}
# check number of cooling effective UGS: 
nrow(cooling_effective)
```

```{r}
cooling_effective %>%
  st_drop_geometry() %>%
  summarise(
    cooling_effective_parks = n(),
    mean_PCI = mean(PCI, na.rm = TRUE),
    median_PCI = median(PCI, na.rm = TRUE),
    mean_PCG = mean(PCG, na.rm = TRUE),
    median_PCG = median(PCG, na.rm = TRUE),
    min_PCG = min(PCG, na.rm = TRUE),
    max_PCG = max(PCG, na.rm = TRUE),
    mean_dmax = mean(Lmax, na.rm = TRUE),
    median_dmax = median(Lmax, na.rm = TRUE)
  )
```

```{r}
# plot map for checks
ggplot(greenspace_pci) +
  geom_sf(aes(fill = PCG), colour = NA) +
  scale_fill_viridis_c(
    name = "PCG",
    na.value = "grey80"
  ) +
  labs(
    title = "Park Cooling Gradient across Green Spaces"
  ) +
  theme_void() +
  theme(
    plot.title = element_text(hjust = 0.5)
  )
```

### 1.4. Limitations

Did not remove pixels of water bodies, vegetation that potentially overlap with the buffer rings.

## 2. Multimodal Rapid Realistic Routing (r5r)
