---
title: "4_Quantify_Urban_Heat_Park_Cooling"
---

This markdown contains the code for mapping spatial distribution of LST and SUHII (Section 1). The calculation for identifying cooling-effective urban green spaces and mapping approaches are in Section 2.

```{r}
# remove objects from R environment and clear memory
rm(list = ls())    
gc()
```

## Load Packages

```{r, message=FALSE, warning=FALSE}
library(sf)
library(dplyr)
library(ggplot2)
library(tmap)
library(cowplot)
library(terra)
library(tidyterra)
library(scales)
```

## Data Loading

```{r}
# greater london boundary 
london_boundary <- st_read("data/gla/London_GLA_Boundary.shp")

# london lst raster
london_lst <- rast("data/London_LST_median.tif")

# london lsoa lst and SUHII
lsoa_lst <- st_read("data/London_LSOA_LST_SUHI/London_LSOA_LST_SUHI.shp")%>% 
  st_transform(.,27700)

# greenspace >= 2ha
greenspace2ha <- st_read("data/greenspace_london_2ha/greenspace_london_2ha.shp")

# greenspace and greenspace buffer lst
greenspace <- st_read("data/greenspace_buffer_lst/greenspace_buffer_lst.shp") %>% 
  st_transform(.,27700)
```

## 1. LST and Urban Heat

**Map: Mean LST and Green Spaces Polygons**

```{r}
# ensure consistent CRS 
greenspace2ha <- st_transform(greenspace2ha, crs(london_lst))

# plot lst raster with UGS 
lst_map <- ggplot() +
  # lst raster 
  geom_spatraster(data = london_lst) +
  # UGS polygons
  geom_sf(
    data = greenspace2ha,
    # create a legend for the UGS polygons 
    aes(colour = "Selected\nUGS"),
    fill = NA,
    linewidth = 0.3,
    key_glyph = "polygon"
  ) +
  # set legend colours for lst raster
  scale_fill_gradientn(
    colours = c(
      "#3F5F8C",
      "#6F9FC4", 
      "#BFD9D3",
      "#F2E8B8",
      "#E7B46A", 
      "#C96A4A",
      "#9E3B32"
    ),
    name = "LST (°C)",
    na.value = "transparent"
  ) +
  # set UGS with black borders
  scale_colour_manual(
    name = NULL,
    values = c("Selected\nUGS" = "black")
  ) +
  # arrange lst colour legend bar above UGS legend
  guides(
    fill = guide_colourbar(order = 1),
    colour = guide_legend(
      order = 2,
      override.aes = list(
        fill = NA,
        colour = "black",
        linewidth = 0.7
      )
    )
  ) +
  coord_sf(expand = FALSE) +
  theme_minimal() +
  theme(
    panel.grid = element_blank(),
    axis.text = element_blank(),
    axis.title = element_blank(),
    axis.ticks = element_blank(),
    legend.position = "right",
    legend.text = element_text(size = 10),
    legend.key.size = unit(0.6, "cm"),
    plot.title = element_text(hjust = 0.5)
  )

lst_map
```

```{r}
# save as png
ggsave(
  filename = "figures/Fig3_Spatial_Distribution_LST.png",
  plot = lst_map,
  width = 11.69,
  height = 8.27,
  units = "in",
  dpi = 300
)
```

**Map: SUHII = LST (LSOA) - LST (London)**

```{r}
# Fix geometry
lsoa_lst <- st_make_valid(lsoa_lst)

# use NYT colours for SUHII 
diff_cols <- c(
  "#006D6F",
  "#268D8B",
  "#58AAA5",
  "#91C9C2",
  "#C8E4DF",
  "#F4F1E9",
  "#F5D8CD",
  "#F2B3A1",
  "#EA8973",
  "#DD604C",
  "#B9362A"
)

# define SUHII breaks
# given the data range of -9.98 to 8.53, so use -10 and 10 range
lims <- c(-10, 10)

# colours group into 2°C intervals
breaks_seq <- seq(-10, 10, by = 2)

# plot
map_plot <- ggplot(lsoa_lst) +
  # SUHII
  geom_sf(
    aes(fill = lst_differ),
    colour = NA,
    linewidth = 0
  ) +
  # apply colour breaks as defined in above
  scale_fill_stepsn(
    colours = diff_cols,
    breaks = breaks_seq,
    limits = lims,
    oob = scales::squish,
    name = "SUHII (°C)"
  ) +
  coord_sf(datum = NA) +
  theme_void() +
  theme(
    legend.position = "none",
    plot.title = element_text(
      hjust = 0.5,
      face = "bold",
      size = 14
    ),
    plot.margin = margin(5, 5, 5, 5)
  )

# create a histogram legend and make each bar = 0.1°C
# colours group into 2°C intervals
hist_legend <- ggplot(
  st_drop_geometry(lsoa_lst),
  aes(x = lst_differ)
) +
  # plot SUHII values distribution with 0.1°C bins 
  geom_histogram(
    aes(fill = after_stat(x)),
    binwidth = 0.1,
    boundary = -10,
    colour = NA
  ) +
  # make a dark grey line at where SUHII = 0 
  geom_vline(
    xintercept = 0,
    linewidth = 0.5,
    colour = "grey30"
  ) +
  # make lines at each 2°C interval colour groups 
  geom_vline(
    xintercept = breaks_seq[breaks_seq != 0],
    linetype = "dashed",
    linewidth = 0.3,
    colour = "grey40"
  ) +
  # ensure histogram colours consistent with SUHII mapping colours
  scale_fill_stepsn(
    colours = diff_cols,
    breaks = breaks_seq,
    limits = lims,
    oob = scales::squish,
    guide = "none"
  ) +
  # histogram x axis
  scale_x_continuous(
    limits = lims,
    breaks = seq(-10, 10, by = 4),
    labels = scales::label_number(accuracy = 1)
  )+
  # histogram y axis 
  scale_y_continuous(
    expand = expansion(mult = c(0, 0.05))
  )+
  # histogram x and y axis names
  labs(
    x = "SUHII (°C)",
    y = "LSOA Count"
  ) +
  # histogram font sizes and other settings
  theme_classic(base_size = 10) +
  theme(
    axis.title = element_text(size = 10),
    axis.text = element_text(size = 8),
    plot.background = element_rect(
      fill = "white",
      colour = NA
    ),
    panel.background = element_rect(
      fill = "white",
      colour = NA
    ),
    plot.margin = margin(3, 3, 3, 3)
  )

# put histogram next to map
suhii_tm <- ggdraw() +
  draw_plot(
    map_plot,
    x = 0,
    y = 0,
    width = 0.9,
    height = 1
  ) +
  draw_plot(
    hist_legend,
    x = 0.72,
    y = 0.10,
    width = 0.25,
    height = 0.22
  )

suhii_tm
```

```{r}
# save as png
ggsave(
  "figures/Fig9_Spatial_Distribution_SUHII.png",
  plot = suhii_tm,
  width = 11.69,
  height = 8.27,
  units = "in",
  dpi = 300,
  bg = "white"
)
```

## 2. Park Cooling Quantification

### 2.1. Data Preparation

```{r}
# check greenspace column names
greenspace %>% colnames()
```

```{r}
# select needed columns only
greenspace_clean <- greenspace %>% 
  select(id, function_, area_ha, park_lst,
         # all of the 20 buffer rings in ascending order
         all_of(paste0("ring_", seq(30, 900, 30))),
         geometry)

# check any NA values
greenspace_clean %>%
  st_drop_geometry() %>%
  select(park_lst, 
         # take all of the buffer ring mean lst
         all_of(paste0("ring_", seq(30, 900, 30)))) %>%
  summarise(
    across(
      everything(),
      # -9999 was previously imputed in GEE scripts representing no values 
      ~ sum(. == -9999, na.rm = TRUE)
    )
  )
```

### 2.2. Cubic Polynomial Curve and Park Cooling Intensity (PCI)

+-----------------------------+--------------------------------------------------------------+-----------------------------------------------------------------------------------------+
| Name                        | Formula                                                      | Literature Support                                                                      |
+=============================+==============================================================+=========================================================================================+
| Cublic Polynomial Curve     | +---------------------+                                      | Du *et al.*, 2022; Cai *et al.*, 2023; Xiong, Xie and Yang, 2024; Yang *et al.*, 2024   |
|                             | | ```                 |                                      |                                                                                         |
|                             | | T(r)=ar^3+br^2+cr+d |                                      |                                                                                         |
|                             | | ```                 |                                      |                                                                                         |
|                             | +---------------------+                                      |                                                                                         |
+-----------------------------+--------------------------------------------------------------+-----------------------------------------------------------------------------------------+
| PCI                         | +----------------------------------------------------------+ | Peng *et al.*, 2021; Du *et al.*, 2022; Xiong, Xie and Yang, 2024; Zhang *et al.*, 2024 |
|                             | | ```                                                      | |                                                                                         |
|                             | | PCI=\frac{L{\times}T_L-{{\int}_0^L}T(r)dr}{L{\times}T_L} | |                                                                                         |
|                             | | ```                                                      | |                                                                                         |
|                             | +----------------------------------------------------------+ |                                                                                         |
+-----------------------------+--------------------------------------------------------------+-----------------------------------------------------------------------------------------+
| Park Cooling Gradient (PCG) | +----------------------------------------------+             | Peng *et al.*, 2021; Xiong, Xie and Yang, 2024                                          |
|                             | | ```                                          |             |                                                                                         |
|                             | | PCG=\frac{L{\times}T_L-{{\int}_0^L}T(r)dr}{L |             |                                                                                         |
|                             | | ```                                          |             |                                                                                         |
|                             | +----------------------------------------------+             |                                                                                         |
+-----------------------------+--------------------------------------------------------------+-----------------------------------------------------------------------------------------+

```{r}
# buffer ring from 30m to 900m each 30m intervals as x values for polynomial fitting
x_vals <- seq(30, 900, 30)
ring_cols <- paste0("ring_", seq(30, 900, 30))

# create an empty list to store results for later
results <- list()

# for loop for all greenspace
for (i in 1:nrow(greenspace_clean)) {
  # for each greenspace
  park <- greenspace_clean[i, ]
  # take mean LST values of buffer ring 
  y_ring_lst <- as.numeric(unlist(park[, ring_cols] %>% st_drop_geometry(), use.names = FALSE))
  # take park LST values
  park_lst <- as.numeric(park$park_lst)
  

  # fit a cubic polynomial curve through the LST values:
  # T(x) = ax3 + bx2 + cx + d
  poly_curve <- lm(y_ring_lst ~ poly(x_vals, degree = 3, raw = TRUE))
  
  # extract polynomial coefficients
  coefs <- coef(poly_curve)
  d <- coefs[1]
  c <- coefs[2]
  b <- coefs[3]
  a <- coefs[4]
  
  # calculate R2 of the fitted curve to check how well it fits 
  # higher R2 should find less error 
  # https://www.nature.com/articles/s41599-023-02209-5
  r2 <- summary(poly_curve)$r.squared
  
  
  
  # find turning point of curve
  
  # first derivative is:
  # T'(x) = 3*a*x^2 + 2*b*x + c
  
  # turning points at where slope = 0:
  # 3*a*x^2 + 2*b*x + c = 0
  
  # quadratic equation:
  # A*x^2 + B*x + C = 0
  A <- 3 * a
  B <- 2 * b
  C <- c

  # use discriminant of the derivative equation to find x: 
  disc <- B^2 - 4 * A * C
  
  # parks with no turning point found, return as NA
  if (is.na(disc) || disc < 0) {
    L <- NA
    TL <- NA
    accumulated_cooling <- NA
    PCI <- NA
    PCG <- NA
    max_lst_diff <- NA
  } else {
   
    # find 2 possible turning points
    root1 <- (-B + sqrt(disc)) / (2 * A)
    root2 <- (-B - sqrt(disc)) / (2 * A)
    
    # only keep turning points within the range 30 - 900m
    valid_roots <- c(root1, root2)
    valid_roots <- valid_roots[valid_roots >= 30 & valid_roots <= 900]
    
    if (length(valid_roots) == 0) {
      L <- NA
      TL <- NA
      accumulated_cooling <- NA
      PCI <- NA
      PCG <- NA
      max_lst_diff <- NA
    } else {
      
      # to identify whether turning point is a local maximum or minimum
      # use second derivative: T''(x) = 6ax + 2b
      second_derivative <- 6 * a * valid_roots + 2 * b
      
      # keep only peak of local maximum
      peak_roots <- valid_roots[second_derivative < 0]
      # if not a peak, return as NA
      if (length(peak_roots) == 0) {
        L <- NA
        TL <- NA
        accumulated_cooling <- NA
        PCI <- NA
        PCG <- NA
        max_lst_diff <- NA
      } else {
        
        # using the initial turning point (first peak) calculate cooling distance
        # L = estimated maximum cooling distance
        L <- min(peak_roots)
        
        # calculate fitted LST at L
        # TL = fitted LST at the max cooling distance (L)
        TL <- a * L^3 + b * L^2 + c * L + d
        
        # calculate fitted_lst_integral from 0 to L
        fitted_lst_integral <- (a/4 * L^4) +
          (b/3 * L^3) + 
          (c/2 * L^2) + 
          (d   * L)
        
        # calculate accumulated PCI
        accumulated_cooling <- (L * TL) - fitted_lst_integral
        
        # PCI
        PCI <- accumulated_cooling / (L * TL)
        
        # calculate PCG
        PCG <- accumulated_cooling / L
        
        # calculate max temperature difference 
        max_lst_diff <- TL - park_lst
      }
    }
  }
  
  # store results
  results[[i]] <- data.frame(
    id = park$id, # park ID
    r2 = round(r2, 3), # R squared, how well the curve fits the data
    L = round(L, 1), # estimated cooling distance
    TL = round(TL, 2), # temp at L
    max_lst_diff = round(max_lst_diff, 2), # difference between surrounding temp and park temp
    accumulated_cooling = round(accumulated_cooling, 1), # accumulated cooling
    PCI = round(PCI, 4), # PCI
    PCG = round(PCG, 2) # PCG
  )
}

# combine results
pci_results <- bind_rows(results)
```

```{r}
# check
cat("Total parks:", nrow(pci_results), "\n")
cat("Parks with valid PCI:", sum(!is.na(pci_results$PCI)), "\n")
cat("Parks with no turning point:", sum(is.na(pci_results$PCI)), "\n")

summary(pci_results$accumulated_cooling)
summary(pci_results$PCI)
summary(pci_results$PCG)
summary(pci_results$r2)
```

```{r}
# join PCI results back to greenspace_clean
greenspace_pci <- greenspace_clean %>%
  left_join(pci_results, by = "id")
```

### 2.3. Identify Cooling Effective UGS

```{r}
# identify cooling effective parks
cooling_effective <- greenspace_pci %>%
  filter(
    # filter away all NA
    !is.na(L),
    !is.na(max_lst_diff),
    !is.na(PCI),
    !is.na(PCG),
    # filter temp difference > 0
    max_lst_diff > 0,
    # filter normalised cooling > 0
    PCI > 0,
    # filter PCG > 0
    PCG > 0
  )
```

```{r}
# check number of cooling effective UGS
nrow(cooling_effective)
```

```{r}
# check stats
cooling_effective %>%
  st_drop_geometry() %>%
  summarise(
    # number of cooling effective ugs
    cooling_effective_parks = n(),
    # lst difference
    min_max_lst_diff = min(max_lst_diff, na.rm = TRUE),
    max_max_lst_diff = max(max_lst_diff, na.rm = TRUE),
    mean_max_lst_diff = mean(max_lst_diff, na.rm = TRUE),
    median_max_lst_diff = median(max_lst_diff, na.rm = TRUE),
    # PCI
    min_PCI = min(PCI, na.rm = TRUE),
    max_PCI = max(PCI, na.rm = TRUE),
    mean_PCI = mean(PCI, na.rm = TRUE),
    median_PCI = median(PCI, na.rm = TRUE),
    # PCG
    min_PCG = min(PCG, na.rm = TRUE),
    max_PCG = max(PCG, na.rm = TRUE),
    mean_PCG = mean(PCG, na.rm = TRUE),
    median_PCG = median(PCG, na.rm = TRUE),
    # L
    mean_L = mean(L, na.rm = TRUE),
    median_L = median(L, na.rm = TRUE)
  )
```

### 2.4. Save as Shapefile

```{r, eval=FALSE}
# create a folder for the shapefiles
dir.create("data/cooling_effective_UGS", recursive = TRUE, showWarnings = FALSE)

# Save as shapefile
st_write(
  cooling_effective,
  "data/cooling_effective_UGS/cooling_effective_UGS.shp",
  delete_dsn = TRUE
)
```

### 2.5. Limitations

Did not remove pixels of water bodies, vegetation that potentially overlap with the buffer rings.

### 2.6. PCI Visualisation 

**Map: Cooling Effective UGS by PCI**

```{r}
# Fix geometries
london_boundary <- st_make_valid(london_boundary)
greenspace_pci <- st_make_valid(greenspace_pci)
cooling_effective <- st_make_valid(cooling_effective)

# Match CRS
london_boundary <- st_transform(london_boundary, st_crs(greenspace_pci))
cooling_effective <- st_transform(cooling_effective, st_crs(greenspace_pci))

# Static tmap mode
tmap_mode("plot")

# Create 4 equal-interval PCI breaks using cooling-effective UGS only
pci_breaks <- seq(
  min(cooling_effective$PCI, na.rm = TRUE),
  max(cooling_effective$PCI, na.rm = TRUE),
  length.out = 5
)

# Teal palette for plotting
# low PCI = light teal, high PCI = dark teal
pci_teals <- c(
  "#B8DDDA",
  "#72C0BB",
  "#2FA5A0",
  "#008F8C"
)

# Add to cooling-effective UGS only
cooling_effective_map <- cooling_effective %>%
  mutate(
    PCI_bin = cut(
      PCI,
      breaks = pci_breaks,
      include.lowest = TRUE
    )
  )
```

```{r}
# ggplot map
pci_map_gg <- ggplot() +
  geom_sf(
    # gla boundary
    data = london_boundary,
    fill = "white",
    color = "grey30",
    linewidth = 0.4
  ) +
  # all selected ugs 
  geom_sf(
    data = greenspace_pci,
    fill = "#E3E3E3",
    color = NA
  ) +
  # cooling effective ugs and colour by PCI
  geom_sf(
    data = cooling_effective_map,
    aes(fill = PCI_bin),
    color = NA
  ) +
  scale_fill_manual(
    values = pci_teals,
    guide = "none"
  ) +
  theme_void() +
  theme(
    plot.background = element_rect(fill = "white", color = NA),
    panel.background = element_rect(fill = "white", color = NA)
  )

# histogram legend for PCI values
hist_legend_pci <- ggplot(
  st_drop_geometry(cooling_effective),
  aes(x = PCI)
) +
  geom_histogram(
    aes(fill = after_stat(x)),
    binwidth = diff(range(pci_breaks)) / 40,
    boundary = pci_breaks[1],
    color = NA
  ) +
  geom_vline(
    xintercept = pci_breaks,
    linetype = "dashed",
    linewidth = 0.35,
    color = "black"
  ) +
  scale_fill_stepsn(
    colours = pci_teals,
    breaks = pci_breaks,
    limits = range(pci_breaks),
    oob = scales::squish,
    guide = "none"
  ) +
  scale_x_continuous(
    limits = range(pci_breaks),
    breaks = pci_breaks,
    labels = scales::label_number(accuracy = 0.01),
    expand = c(0, 0)
  ) +
  scale_y_continuous(
   expand = expansion(mult = c(0, 0.05))
  ) +
  labs(
    x = "PCI",
    y = "UGS Count"
  ) +
  theme_classic(base_size = 9) +
  theme(
    axis.title = element_text(size = 9),
    axis.text = element_text(size = 7),
    plot.background = element_rect(fill = "white", color = NA),
    panel.background = element_rect(fill = "white", color = NA),
    plot.margin = margin(3, 3, 3, 3)
  )

# combine plot together
pci_map <- ggdraw() +
  draw_plot(
    pci_map_gg,
    x = 0,
    y = 0,
    width = 0.9,
    height = 1
  ) +
  draw_plot(
    hist_legend_pci,
    x = 0.72,
    y = 0.10,
    width = 0.25,
    height = 0.22
  )

pci_map
```

**Map cooling-effective UGS by PCI filtered for public park or garden and playing field**:

```{r}
# select UGS function and needed columns
cooling_effective_UGS_clean <- cooling_effective %>%
  # select needed columns
  select(id, function_, PCI, geometry) %>%
  # filter park functions to Public Park or Garden and Playing Field only
  # reasons refer to above table
  filter(function_ %in% c("Public Park Or Garden", "Playing Field"))
```

```{r}
summary(cooling_effective_UGS_clean)
```

```{r}
cooling_effective_UGS_clean %>%
  st_drop_geometry() %>%
  mutate(
    PCI_bin = cut(
      PCI,
      breaks = c(0.00, 0.01, 0.02, 0.03),
      include.lowest = TRUE
    )
  ) %>%
  count(PCI_bin)
```

```{r}
# Cooling-effective UGS that were removed by function filtering
cooling_effective_filtered_out <- cooling_effective %>%
  filter(!function_ %in% c("Public Park Or Garden", "Playing Field"))
```

```{r}
# make breaks same as previous map
pci_breaks <- c(0.00, 0.03, 0.06, 0.09, 0.11)

# labels
pci_labels <- c(
  "0.00–0.03",
  "0.03–0.06",
  "0.06–0.09",
  "0.09–0.11"
)

# colours
pci_teals <- c(
  "0.00–0.03" = "#B8DDDA",
  "0.03–0.06" = "#72C0BB",
  "0.06–0.09" = "#2FA5A0",
  "0.09–0.11" = "#008F8C"
)

# Create plotting variable and PCI 
cooling_effective_UGS_clean <- cooling_effective_UGS_clean %>%
  mutate(
    # Keeps any very small values at 0 and values above 0.11
    # inside the displayed range, matching oob = squish
    PCI_plot = scales::squish(
      PCI,
      range = range(pci_breaks)
    ),

    PCI_bin = cut(
      PCI_plot,
      breaks = pci_breaks,
      labels = pci_labels,
      include.lowest = TRUE,
      right = TRUE
    )
  )
```

```{r}
# plot 
pci_map_gg_clean <- ggplot() +
  geom_sf(
    data = london_boundary,
    fill = "white",
    color = "grey30",
    linewidth = 0.4
  ) +
  geom_sf(
    # not cooling effective as grey first
    data = greenspace_pci,
    fill = "#E3E3E3",
    color = NA
  ) +
  geom_sf(
    # cooling effective but removed for functional purposes in darker grey
    data = cooling_effective_filtered_out, 
    fill = "#9E9E9E", 
    color = NA
  ) +
  geom_sf(
    # the final cooling effective ugs 
    data = cooling_effective_UGS_clean,
    aes(fill = PCI_bin),
    color = NA
  ) +
  scale_fill_manual(
    values = pci_teals,
    guide = "none"
  ) +
  theme_void() +
  theme(
    plot.background = element_rect(fill = "white", color = NA),
    panel.background = element_rect(fill = "white", color = NA)
  )


# histogram legend
hist_legend_pci_clean <- ggplot(
  st_drop_geometry(cooling_effective_UGS_clean), 
  aes(x = PCI)
) +
  geom_histogram(
    aes(fill = after_stat(x)),
    binwidth = 0.003,
    boundary = 0,
    color = NA
) +
  geom_vline(
    xintercept = pci_breaks,
    linetype = "dashed",
    linewidth = 0.35,
    color = "black"
  ) +
  scale_fill_stepsn(
    colours = pci_teals,
    breaks = pci_breaks,
    limits = range(pci_breaks),
    oob = scales::squish,
    guide = "none"
  ) +
  scale_x_continuous(
    limits = range(pci_breaks),
    breaks = pci_breaks,
    labels = scales::label_number(accuracy = 0.01),
    expand = c(0, 0)
  ) +
  scale_y_continuous(
    limits = c(0, 100), # ensure same range (plotted before and the y axis was 60)
    breaks = c(0, 30, 60, 90),
    expand = c(0, 0)
  ) +
  labs(
    x = "PCI",
    y = "UGS Count"
  ) +
  theme_classic(base_size = 9) +
  theme(
    axis.title = element_text(size = 9),
    axis.text = element_text(size = 7),
    plot.background = element_rect(fill = "white", color = NA),
    panel.background = element_rect(fill = "white", color = NA),
    plot.margin = margin(3, 3, 3, 3)
  )

# combine plots together
pci2_map <- ggdraw() +
  draw_plot(
    pci_map_gg_clean,
    x = 0, y = 0, width = 0.9, height = 1
  ) +
  draw_plot(
    hist_legend_pci_clean,
    x = 0.72, y = 0.10, width = 0.25, height = 0.22
  )

pci2_map
```

```{r}
# cowplot combine two map objects together in one output 
combined_pci_maps <- plot_grid(
  pci_map,
  pci2_map,
  labels     = c("a", "b"),
  label_size = 14,
  ncol       = 1
)

combined_pci_maps
```

```{r}
# save as PNG
ggsave(
  filename = "figures/Fig10_Spatial_Distribution_PCI.png",
  plot = combined_pci_maps,
  width = 7.5,
  height = 10,
  units = "in",
  dpi = 300,
  bg = "white"
)
```
