---
title: "4_Cooling_Effective_UGS_Accessibility"
---

This markdown file contains the full working code for ......

```{r}
# remove objects from R environment and clear memory
rm(list = ls())    
gc()
```

## Load Packages

```{r, message=FALSE, warning=FALSE}
library(sf)
library(here)
library(dplyr)
library(ggplot2)
library(tidyverse)
library(tidyr)
library(terra)
library(tidyterra)
library(viridis)
library(tmap)
library(cowplot)
library(scales)
library(RColorBrewer)
```

## Data Loading

**Greater London Boundary:**

```{r}
london_boundary <- st_read("data/gla/London_london_boundary.shp")
```

**London LST Raster:**

```{r}
london_lst <- rast("data/London_LST_median.tif")
```

**London LSOA LST and SUHII:**

```{r}
lsoa_lst <- st_read("data/London_LSOA_LST_SUHI/London_LSOA_LST_SUHI.shp")%>% 
  st_transform(.,27700)
```

**Green Space \> 2ha:**

```{r}
greenspace2ha <- st_read("data/greenspace_london_2ha/greenspace_london_2ha.shp")
```

**Green Space and Green Space Buffer LST:**

```{r}
greenspace <- st_read("data/greenspace_buffer_lst/greenspace_buffer_lst.shp") %>% 
  st_transform(.,27700)
```

## 1. Urban Heat

**Map: Mean LST and Green Spaces Polygons**

```{r}
# Ensure consistent CRS 
greenspace2ha <- st_transform(greenspace2ha, crs(london_lst))

# Plot
lst_map <- ggplot() +
  geom_spatraster(data = london_lst) +
  geom_sf(
    data = greenspace2ha,
    aes(colour = "Selected\nUGS"),
    fill = NA,
    linewidth = 0.3,
    key_glyph = "polygon"
  ) +
  scale_fill_gradientn(
    colours = c(
      "#2F6FA8",  # stronger blue
      "#4FA3D9",  # brighter blue
      "#8FD0E8",  # clearer light blue
      "#FFF08A",  # brighter yellow
      "#FDB056",  # stronger orange
      "#F26A3D",  # stronger orange-red
      "#C93A32"   # deeper red
    ),
    name = "LST (°C)",
    na.value = "transparent"
  ) +
  scale_colour_manual(
    name = NULL,
    values = c("Selected\nUGS" = "black")
  ) +
  guides(
    fill = guide_colourbar(order = 1),
    colour = guide_legend(
      order = 2,
      override.aes = list(
        fill = NA,
        colour = "black",
        linewidth = 0.7
      )
    )
  ) +
  coord_sf(expand = FALSE) +
  theme_minimal() +
  theme(
    panel.grid = element_blank(),
    axis.text = element_blank(),
    axis.title = element_blank(),
    axis.ticks = element_blank(),
    legend.position = "right",
    legend.text = element_text(size = 10),
    legend.key.size = unit(0.6, "cm"),
    plot.title = element_text(hjust = 0.5)
  )

lst_map
```

```{r}
# save as png
ggsave(
  filename = "figures/Fig2_Mean_LST_Greenspace_2ha_map.png",
  plot = lst_map,
  width = 11.69,
  height = 8.27,
  units = "in",
  dpi = 300
)
```

**Map: SUHII = LST (LSOA) - LST (London)**

```{r}
# Fix geometry
lsoa_lst <- st_make_valid(lsoa_lst)

# set colours
# blue = cooler than London mean, red = hotter than London mean
diff_cols <- rev(brewer.pal(11, "RdBu"))


# define SUHII breaks
# data range: -9.98 to 8.53 so use -10 and 10 range
lims <- c(-10, 10)

# colours group into 2°C intervals
breaks_seq <- seq(-10, 10, by = 2)

# map plot
map_plot <- ggplot(lsoa_lst) +
  geom_sf(
    aes(fill = lst_differ),
    color = NA,
    linewidth = 0
  ) +
  scale_fill_stepsn(
    colours = diff_cols,
    breaks = breaks_seq,
    limits = lims,
    oob = squish,
    name = "SUHII"
  ) +
  coord_sf(datum = NA) +
  theme_void() +
  theme(
    legend.position = "none",
    plot.title = element_text(
      hjust = 0.5,
      face = "bold",
      size = 14
    ),
    plot.margin = margin(5, 5, 5, 5)
  )

# histogram legend plot - each bar = 0.1°C
# colours group into 2°C intervals
hist_legend <- ggplot(
  st_drop_geometry(lsoa_lst),
  aes(x = lst_differ)
) +
  geom_histogram(
    aes(fill = after_stat(x)),
    binwidth = 0.1,
    boundary = -10,
    color = NA
  ) +
  geom_vline(
    xintercept = breaks_seq,
    linetype = "dashed",
    linewidth = 0.35,
    color = "black"
  ) +
  scale_fill_stepsn(
    colours = diff_cols,
    breaks = breaks_seq,
    limits = lims,
    oob = squish,
    guide = "none"
  ) +
  scale_x_continuous(
    limits = lims,
    breaks = seq(-10, 10, by = 4),
    labels = label_number(accuracy = 1)
  ) +
  labs(
    x = "SUHII (°C)",
    y = "LSOA Count"
  ) +
  theme_classic(base_size = 9) +
  theme(
    axis.title = element_text(size = 9),
    axis.text = element_text(size = 7),
    plot.background = element_rect(fill = "white", color = NA),
    panel.background = element_rect(fill = "white", color = NA),
    plot.margin = margin(3, 3, 3, 3)
  )


# put histogram next to map
suhii_tm <- ggdraw() +
  draw_plot(
    map_plot,
    x = 0,
    y = 0,
    width = 0.9,
    height = 1
  ) +
  draw_plot(
    hist_legend,
    x = 0.72,
    y = 0.10,
    width = 0.25,
    height = 0.22
  )

suhii_tm
```

```{r}
# save as png
ggsave(
  "figures/Fig3_LSOA_SUHII.png",
  plot = suhii_tm,
  width = 11.69,
  height = 8.27,
  units = "in",
  dpi = 300,
  bg = "white"
)
```

## 2. Park Cooling Quantification

### 2.1. Data Preparation

```{r}
# check greenspace column names
greenspace %>% colnames()
```

```{r}
# select needed columns only
greenspace_clean <- greenspace %>% 
  select(id, function_, area_ha, park_lst,
         # all of the 20 buffer rings in ascending order
         all_of(paste0("ring_", seq(30, 900, 30))),
         geometry)
```

```{r}
# check any NA values
greenspace_clean %>%
  st_drop_geometry() %>%
  summarise(across(everything(), ~sum(is.na(.))))
```

### 2.2. Cubic Polynomial Curve and Park Cooling Intensity (PCI)

+-----------------------------+--------------------------------------------------------------+-----------------------------------------------------------------------------------------+
| Name                        | Formula                                                      | Literature Support                                                                      |
+=============================+==============================================================+=========================================================================================+
| Cublic Polynomial Curve     | +---------------------+                                      | Du *et al.*, 2022; Cai *et al.*, 2023; Xiong, Xie and Yang, 2024; Yang *et al.*, 2024   |
|                             | | ```                 |                                      |                                                                                         |
|                             | | T(r)=ar^3+br^2+cr+d |                                      |                                                                                         |
|                             | | ```                 |                                      |                                                                                         |
|                             | +---------------------+                                      |                                                                                         |
+-----------------------------+--------------------------------------------------------------+-----------------------------------------------------------------------------------------+
| PCI                         | +----------------------------------------------------------+ | Peng *et al.*, 2021; Du *et al.*, 2022; Xiong, Xie and Yang, 2024; Zhang *et al.*, 2024 |
|                             | | ```                                                      | |                                                                                         |
|                             | | PCI=\frac{L{\times}T_L-{{\int}_0^L}T(r)dr}{L{\times}T_L} | |                                                                                         |
|                             | | ```                                                      | |                                                                                         |
|                             | +----------------------------------------------------------+ |                                                                                         |
+-----------------------------+--------------------------------------------------------------+-----------------------------------------------------------------------------------------+
| Park Cooling Gradient (PCG) | +----------------------------------------------+             | Peng *et al.*, 2021; Xiong, Xie and Yang, 2024                                          |
|                             | | ```                                          |             |                                                                                         |
|                             | | PCG=\frac{L{\times}T_L-{{\int}_0^L}T(r)dr}{L |             |                                                                                         |
|                             | | ```                                          |             |                                                                                         |
|                             | +----------------------------------------------+             |                                                                                         |
+-----------------------------+--------------------------------------------------------------+-----------------------------------------------------------------------------------------+

```{r}
# Ring distances in metres — x values for polynomial fitting
x_vals <- seq(30, 900, 30)
ring_cols <- paste0("ring_", seq(30, 900, 30))

# create an empty list to store results for later
results <- list()

# for loop for each greenspace
for (i in 1:nrow(greenspace_clean)) {
  # for each greenspace
  park <- greenspace_clean[i, ]
  # mean LST values of buffer ring 
  y_ring_lst <- as.numeric(unlist(park[, ring_cols] %>% st_drop_geometry(), use.names = FALSE))
  # park LST values
  park_lst <- as.numeric(park$park_lst)
  
  # fit a cubic polynomial curve: T(x) = ax3 + bx2 + cx + d
  poly_curve <- lm(y_ring_lst ~ poly(x_vals, degree = 3, raw = TRUE))
  
  # coefficients
  coefs <- coef(poly_curve)
  d <- coefs[1]
  c <- coefs[2]
  b <- coefs[3]
  a <- coefs[4]
  
  # calculate R squared 
  # it is said there is found less error with higher r2 (https://www.nature.com/articles/s41599-023-02209-5)
  r2 <- summary(poly_curve)$r.squared
  
  # turning point of curve --> find where temp stop decreasing/ increasing
  # find derivative: T'(x) = 3ax2 + 2bx + c 
  disc <- (2*b)^2 - 4*(3*a)*c
  # parks with no valid turning point, return as NA
  if (is.na(disc) || disc < 0) {
    L <- NA
    # LST value at the first turning point
    TL <- NA
    accumulated_cooling <- NA
    PCI <- NA
    PCG <- NA
    max_lst_diff <- NA
  } else {
   
    # distances where temp stops increasing/ decreasing
    root1 <- (-(2*b) + sqrt(disc)) / (2 * 3*a)
    root2 <- (-(2*b) - sqrt(disc)) / (2 * 3*a)
    
    # only keep turning points within ring range 30 - 900m
    valid_roots <- c(root1, root2)
    valid_roots <- valid_roots[valid_roots >= 30 & valid_roots <= 900]
    
    if (length(valid_roots) == 0) {
      L <- NA
      TL <- NA
      accumulated_cooling <- NA
      PCI <- NA
      PCG <- NA
      max_lst_diff <- NA
    } else {
      
      # identify whether turning point is at highest point of curve
      # derivative: T''(x) = 6ax + 2b
      second_derivative <- 6 * a * valid_roots + 2 * b
      
      # keep only peaks
      peak_roots <- valid_roots[second_derivative < 0]
      # if it s not a peak, return as NA
      if (length(peak_roots) == 0) {
        L <- NA
        TL <- NA
        accumulated_cooling <- NA
        PCI <- NA
        PCG <- NA
        max_lst_diff <- NA
      } else {
        
        # select the initial turning point (first peak) as cooling distance
        L <- min(peak_roots)
        
        # calculate fitted LST at the L
        TL <- a * L^3 + b * L^2 + c * L + d
        
        # calculate actual_accumulated_lst from 0 to L
        actual_accumulated_lst <- (a/4 * L^4) +
          (b/3 * L^3) + 
          (c/2 * L^2) + 
          (d   * L)
        
        # calculate accumulated PCI
        accumulated_cooling <- (L * TL) - actual_accumulated_lst
        
        # PCI
        PCI <- accumulated_cooling / (L * TL)
        
        # calculate PCG
        PCG <- accumulated_cooling / L
        
        # calculate max temperature difference
        max_lst_diff <- TL - park_lst
      }
    }
  }
  
  # store results
  results[[i]] <- data.frame(
    id = park$id, # park ID
    r2 = round(r2, 3), # R squared, how well the curve fits the data
    L = round(L, 1), # estimated cooling distance
    TL = round(TL, 2), # temp at L
    max_lst_diff = round(max_lst_diff, 2), # difference between surrounding temp and park temp
    accumulated_cooling = round(accumulated_cooling, 1), # accumulated cooling
    PCI = round(PCI, 4), # PCI
    PCG = round(PCG, 2) # PCG
  )
}

# combine results
pci_results <- bind_rows(results)

# check
cat("Total parks:", nrow(pci_results), "\n")
cat("Parks with valid PCI:", sum(!is.na(pci_results$PCI)), "\n")
cat("Parks with no turning point:", sum(is.na(pci_results$PCI)), "\n")

summary(pci_results$accumulated_cooling)
summary(pci_results$PCI)
summary(pci_results$PCG)
summary(pci_results$r2)
```

```{r}
# join PCI results back to greenspace_clean
greenspace_pci <- greenspace_clean %>%
  left_join(pci_results, by = "id")
```

### 2.3. Identify Cooling Effective UGS

```{r}
# identify cooling effective parks
cooling_effective <- greenspace_pci %>%
  filter(
    # filter away all NA
    !is.na(L),
    !is.na(max_lst_diff),
    !is.na(PCI),
    !is.na(PCG),
    # filter temp difference > 0
    max_lst_diff > 0,
    # filter normalised cooling > 0
    PCI > 0,
    # filter PCG > 0
    PCG > 0
  )
```

```{r}
# check number of cooling effective UGS
nrow(cooling_effective)
```

```{r}
# check stats
cooling_effective %>%
  st_drop_geometry() %>%
  summarise(
    # number of cooling effective ugs
    cooling_effective_parks = n(),
    # lst difference
    min_max_lst_diff = min(max_lst_diff, na.rm = TRUE),
    max_max_lst_diff = max(max_lst_diff, na.rm = TRUE),
    mean_max_lst_diff = mean(max_lst_diff, na.rm = TRUE),
    median_max_lst_diff = median(max_lst_diff, na.rm = TRUE),
    # PCI
    min_PCI = min(PCI, na.rm = TRUE),
    max_PCI = max(PCI, na.rm = TRUE),
    mean_PCI = mean(PCI, na.rm = TRUE),
    median_PCI = median(PCI, na.rm = TRUE),
    # PCG
    min_PCG = min(PCG, na.rm = TRUE),
    max_PCG = max(PCG, na.rm = TRUE),
    mean_PCG = mean(PCG, na.rm = TRUE),
    median_PCG = median(PCG, na.rm = TRUE),
    # L
    mean_L = mean(L, na.rm = TRUE),
    median_L = median(L, na.rm = TRUE)
  )
```

**Map: Cooling Effective UGS by PCI**

```{r, eval=FALSE}
# not to run, keep code

# Match CRS
london_boundary <- st_transform(london_boundary, st_crs(greenspace_pci))
cooling_effective <- st_transform(cooling_effective, st_crs(greenspace_pci))

# Create 4 equal-interval PCI breaks using cooling-effective UGS only
pci_breaks <- seq(
  min(cooling_effective$PCI, na.rm = TRUE),
  max(cooling_effective$PCI, na.rm = TRUE),
  length.out = 5
)

# Clean legend labels
pci_labels <- paste0(
  sprintf("%.2f", pci_breaks[-length(pci_breaks)]),
  " - ",
  sprintf("%.2f", pci_breaks[-1])
)

# Blue palette: low PCI = light blue, high PCI = dark blue
pci_blues <- c(
  "#C6DBEF", 
  "#6BAED6", 
  "#3182BD", 
  "#08519C"
)

# Add PCI classes to cooling-effective UGS only
cooling_effective_map <- cooling_effective %>%
  mutate(
    PCI_bin = cut(
      PCI,
      breaks = pci_breaks,
      labels = pci_labels,
      include.lowest = TRUE
    )
  )

# Plot map: all greenspaces grey underneath, cooling-effective UGS on top
pci_map <- ggplot() +
  geom_sf(
    data = london_boundary,
    fill = "white",
    colour = "grey30",
    linewidth = 0.4
  ) +
  geom_sf(
    data = greenspace_pci,
    aes(fill = "< 0.00"),
    colour = NA
  ) +
  geom_sf(
    data = cooling_effective_map,
    aes(fill = PCI_bin),
    colour = NA
  ) +
  scale_fill_manual(
    name = "PCI",
    values = c(
      "< 0.00" = "grey80",
      setNames(pci_blues, pci_labels)
    ),
    breaks = c("< 0.00", pci_labels),
    drop = FALSE
  ) +
  coord_sf(datum = NA) +
  theme_void() +
  theme(
    legend.position = "right",
    legend.title = element_text(size = 12, face = "bold"),
    legend.text = element_text(size = 10),
    legend.key.size = unit(0.45, "cm"),
    plot.background = element_rect(fill = "white", color = NA)
  )

pci_map
```

```{r}
# Fix geometries
london_boundary <- st_make_valid(london_boundary)
greenspace_pci <- st_make_valid(greenspace_pci)
cooling_effective <- st_make_valid(cooling_effective)

# Match CRS
london_boundary <- st_transform(london_boundary, st_crs(greenspace_pci))
cooling_effective <- st_transform(cooling_effective, st_crs(greenspace_pci))

# Static tmap mode
tmap_mode("plot")

# Create 4 equal-interval PCI breaks using cooling-effective UGS only
pci_breaks <- seq(
  min(cooling_effective$PCI, na.rm = TRUE),
  max(cooling_effective$PCI, na.rm = TRUE),
  length.out = 5
)

# Blue palette: low PCI = light blue, high PCI = dark blue
pci_blues <- c(
  "#C6DBEF", 
  "#6BAED6", 
  "#3182BD", 
  "#08519C"
)

# Add PCI classes to cooling-effective UGS only
cooling_effective_map <- cooling_effective %>%
  mutate(
    PCI_bin = cut(
      PCI,
      breaks = pci_breaks,
      include.lowest = TRUE
    )
  )

# ggplot2 map — color = NA means NO borders, guaranteed
pci_map_gg <- ggplot() +
  geom_sf(
    data = london_boundary,
    fill = "white",
    color = "grey30",
    linewidth = 0.4
  ) +
  geom_sf(
    data = greenspace_pci,
    fill = "grey80",
    color = NA          # no border
  ) +
  geom_sf(
    data = cooling_effective_map,
    aes(fill = PCI_bin),
    color = NA          # no border
  ) +
  scale_fill_manual(values = pci_blues, guide = "none") +
  theme_void() +
  theme(
    plot.background = element_rect(fill = "white", color = NA),
    panel.background = element_rect(fill = "white", color = NA)
  )

# Histogram legend for PCI values
hist_legend_pci <- ggplot(
  st_drop_geometry(cooling_effective),
  aes(x = PCI)
) +
  geom_histogram(
    aes(fill = after_stat(x)),
    bins = 40,
    color = NA
  ) +
  geom_vline(
    xintercept = pci_breaks,
    linetype = "dashed",
    linewidth = 0.35,
    color = "black"
  ) +
  scale_fill_stepsn(
    colours = pci_blues,
    breaks = pci_breaks,
    limits = range(pci_breaks),
    oob = scales::squish,
    guide = "none"
  ) +
  scale_x_continuous(
    limits = range(pci_breaks),
    breaks = pci_breaks,
    labels = scales::label_number(accuracy = 0.01),
    expand = c(0, 0)
  ) +
  labs(
    x = "PCI",
    y = "UGS Count"
  ) +
  theme_classic(base_size = 9) +
  theme(
    axis.title = element_text(size = 9),
    axis.text = element_text(size = 7),
    plot.background = element_rect(fill = "white", color = NA),
    panel.background = element_rect(fill = "white", color = NA),
    plot.margin = margin(3, 3, 3, 3)
  )

# plot together
pci_map <- ggdraw() +
  draw_plot(
    pci_map_gg,
    x = 0, 
    y = 0,
    width = 0.9, 
    height = 1
  ) +
  draw_plot(
    hist_legend_pci,
    x = 0.72, 
    y = 0.10,
    width = 0.25, 
    height = 0.22
  )

pci_map
```

```{r}
# save as PNG
ggsave(
  filename = "figures/Fig4_Cooling_Effective_UGS_PCI_map.png",
  plot = pci_map,
  width = 11.69,
  height = 8.27,
  units = "in",
  dpi = 300,
  bg = "white"
)
```

### 2.4. Save as Shapefile

```{r}
# create a folder
dir.create("data/cooling_effective_UGS", recursive = TRUE, showWarnings = FALSE)

# Save as shapefile
st_write(
  cooling_effective,
  "data/cooling_effective_UGS/cooling_effective_UGS.shp",
  delete_dsn = TRUE
)
```

### 2.5. Limitations

Did not remove pixels of water bodies, vegetation that potentially overlap with the buffer rings.
