---
title: "5_Multimodal_Accessibility_r5r"
---

This markdown file contains the code for taking origin and destination points, building r5r routing network for obtaining travel time matrices to derive accessibility indicators (minimum travel time and cumulative opportunity).

```{r}
# remove objects from R environment and clear memory
rm(list = ls())    
gc()
```

## Load Packages

```{r, message=FALSE, warning=FALSE}
library(sf)
library(here)
library(dplyr)
library(ggplot2)
library(tidyverse)
library(tidytransit)
library(gtfstools)
```

## Data Loading

```{r}
# greater london boundary
london_boundary <- st_read("data/gla/London_GLA_Boundary.shp")

# london LSOA boundaries
london_lsoa <- st_read("data/lsoa_london/lsoa_london.shp")

# greenspaces > 2ha
greenspace2ha <- st_read("data/greenspace_london_2ha/greenspace_london_2ha.shp")

# cooling-effective greenspace shapefile
# exported from 4_Quantify_Urban_Heat_Park_Cooling.Rmd
cooling_effective_UGS <- st_read("data/cooling_effective_UGS/cooling_effective_UGS.shp") %>%
  st_transform(.,27700)

# OS open greenspace access points
greenspace_access_points <- st_read("data/raw_data/OS-Open-Greenspace-GB/GB_AccessPoint/GB_AccessPoint.shp") 

# london lsoa population weighted centroids
london_lsoa_pop_weighted_centroids <- st_read("data/london_lsoa_pop_weighted_centroids/london_lsoa_pop_weighted_centroids.shp")
```

## 1. Origins

The loaded data `london_lsoa_pop_weighted_centroids` contains the origns to be used in r5r routing.

**Map 1. London LSOA population weighted centroids**

```{r}
# plot
london_centroids_map <- ggplot() +
  geom_sf(
    data = london_lsoa,
    fill = NA,
    colour = "grey80"
  ) +
  geom_sf(
    data = london_boundary,
    fill = NA,
    colour = "grey30"
  ) +
  geom_sf(
    data = london_lsoa_pop_weighted_centroids,
    aes(fill = "Population Weighted Centroids"),
    shape = 21,
    colour = "black",
    stroke = 0.08,
    size = 0.8,
    alpha = 0.8
  ) +
  scale_fill_manual(
    values = c("Population Weighted Centroids" = "#C96F57"),
    name = NULL
  ) +
  guides(
    fill = guide_legend(
      override.aes = list(
        shape = 21,
        colour = "black",
        size = 3,
        stroke = 0.2,
        alpha = 0.8
      )
    )
  ) +
  coord_sf(datum = NA) +
  theme_void() +
  theme(
    plot.background = element_rect(fill = "white", colour = NA),
    legend.position = "bottom",
    legend.justification = "center",
    legend.text = element_text(size = 10),
    legend.margin = margin(t = 2, b = 2)
  )

london_centroids_map
```

## 2. Destinations

```{r}
# check column names
names(cooling_effective_UGS)
```

### 2.1. Filter Function of Cooling Effective UGS

+----------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| Support                          | Reason                                                                                                                                                                                                                                         |
+==================================+================================================================================================================================================================================================================================================+
| Green Infrastructure Framework a | 1.2 Accessible Green Infrastructure (AGI) defined accessible greenspace used in Version 2.2 of the Green Infrastructure Mapping: <https://designatedsites.naturalengland.org.uk/GreenInfrastructure/UserGuide/GreenInfrastructureModules.aspx> |
+----------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| Green Infrastructure Framework b | Reasons (allotments and golf course): <https://designatedsites.naturalengland.org.uk/GreenInfrastructure/Glossary.aspx>                                                                                                                        |
+----------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| Smith et al. (2026)              | Greenspaces in the categories 'allotments or community growing spaces', 'other sports facility', 'tennis court', 'bowling green' and 'golf course' are removed                                                                                 |
|                                  |                                                                                                                                                                                                                                                |
|                                  | <https://www.sciencedirect.com/science/article/pii/S1618866725005552?__cf_chl_f_tk=kEs2lY5mgF92ex7FIa4v7E.LiQsjcC9.0QWhg03lBFk-1783176049-1.0.1.1-jCcO8_o0U3UK2CNrm0z8GHQdQ8xAGZDVW20H9a2L8.A>                                                 |
+----------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+

+----------------------------------------+--------------------+--------------------------------------------------------------------------------------------------+-----------------------------------------------------+
| Function                               |                    | Reason                                                                                           | Support                                             |
+========================================+====================+==================================================================================================+=====================================================+
| Cemetery                               | Accessible, but No |                                                                                                  | Green Infrastructure Framework                      |
+----------------------------------------+--------------------+--------------------------------------------------------------------------------------------------+-----------------------------------------------------+
| Golf Course                            | No                 | May require club membership and or payment of a fee to access (Green Infrastructure Framework b) | Green Infrastructure Framework, Smith et al. (2026) |
+----------------------------------------+--------------------+--------------------------------------------------------------------------------------------------+-----------------------------------------------------+
| Other Sports Facility                  | No                 |                                                                                                  | Green Infrastructure Framework, Smith et al. (2026) |
+----------------------------------------+--------------------+--------------------------------------------------------------------------------------------------+-----------------------------------------------------+
| Tennis Court                           | No                 |                                                                                                  | Green Infrastructure Framework, Smith et al. (2026) |
+----------------------------------------+--------------------+--------------------------------------------------------------------------------------------------+-----------------------------------------------------+
| Religious Grounds                      | Accessible, but No |                                                                                                  | Green Infrastructure Framework                      |
+----------------------------------------+--------------------+--------------------------------------------------------------------------------------------------+-----------------------------------------------------+
| Allotments or Community Growing Spaces | No                 | Normally locked and only accessible to key holders (Green Infrastructure Framework b)            | Green Infrastructure Framework, Smith et al. (2026) |
+----------------------------------------+--------------------+--------------------------------------------------------------------------------------------------+-----------------------------------------------------+

```{r}
# select UGS function and needed columns
cooling_effective_UGS_clean <- cooling_effective_UGS %>%
  # select needed columns
  select(id, functn_, geometry) %>%
  # filter park functions to Public Park or Garden and Playing Field only
  # reasons refer to above table
  filter(functn_ %in% c("Public Park Or Garden", "Playing Field"))

# check
nrow(cooling_effective_UGS_clean)
names(cooling_effective_UGS_clean)
unique(cooling_effective_UGS_clean$functn_)
```

### 2.2. Join Filtered Cooling Effective UGS with Access Points

```{r}
# check before joining
nrow(greenspace_access_points)
names(greenspace_access_points)
```

```{r}
# join by greenspace site ID
cooling_effective_access_points <- greenspace_access_points %>%
  inner_join(
    cooling_effective_UGS_clean %>%
      st_drop_geometry() %>%
      select(id, functn_),
    by = c("refToGSite" = "id")
  )
```

```{r}
# check number of access points
nrow(cooling_effective_access_points)
```

**Map 2: Cooling Effective Greenspace Access Points**

```{r}
# ensure consist CRS
london_boundary <- st_transform(london_boundary, st_crs(cooling_effective_UGS_clean))
greenspace2ha <- st_transform(greenspace2ha, st_crs(cooling_effective_UGS_clean))
cooling_effective_access_points <- st_transform(cooling_effective_access_points, st_crs(cooling_effective_UGS_clean))

access_point_map <- ggplot() +
  # greater london boundary
  geom_sf(
    data = london_boundary,
    fill = "white",
    colour = "grey30"
  ) +
  # greenspaces >= 2ha layer as light grey
  geom_sf(
    data = greenspace2ha,
    fill = "grey85",
    colour = NA
  ) +
  # cooling effective UGS layer
  geom_sf(
    data = cooling_effective_UGS_clean,
    fill = "#B8DDDA", # light teal used in PCI maps
    colour = NA
  ) +
  # plot on top the access points of all cooling effective UGS
  geom_sf(
    data = cooling_effective_access_points,
    aes(fill = "Access Points"),
    shape = 25,  # upside-down triangle
    colour = "black",
    stroke = 0.15,
    size = 0.8,
    alpha = 0.9
  ) +
  # set colour for access points 
  scale_fill_manual(
    values = c("Access Points" = "#008F8C"),   # dark teal
    name = NULL
  ) +
  guides(
    fill = guide_legend(
      override.aes = list(
        shape = 25,
        colour = "black",
        size = 3,
        stroke = 0.25,
        alpha = 0.9
      )
    )
  ) +
  coord_sf(datum = NA) +
  theme_void() +
  theme(
    plot.background = element_rect(fill = "white", colour = NA),
    legend.position = "bottom",
    legend.justification = "center",
    legend.text = element_text(size = 10),
    legend.margin = margin(t = 2, b = 2)
  )

access_point_map
```

## 3. Prepare Origin and Destinations for Routing

| Routing O-D  | Objects                            |
|--------------|------------------------------------|
| Origins      | london_lsoa_pop_weighted_centroids |
| Destinations | cooling_effective_access_points    |

```{r}
# check column names
names(london_lsoa_pop_weighted_centroids)
names(cooling_effective_access_points)
```

r5r routing requires lon lat coordinates, so first step is to transform them:

```{r}
#origins
# transform to EPSG 4326
origins_sf <- london_lsoa_pop_weighted_centroids %>%
  st_transform(4326)

# extract lon and lat coordinates from origins
origins_coords <- st_coordinates(origins_sf)

# create an origins df  
origins <- origins_sf %>%
  st_drop_geometry() %>%
  transmute(
    id = as.character(LSOA21CD),
    lon = origins_coords[, 1],
    lat = origins_coords[, 2]
  )


# destinations
# transform to EPSG 4326
destinations_sf <- cooling_effective_access_points %>%
  st_transform(4326)

# extract lon and lat coordinates from destinations
destinations_coords <- st_coordinates(destinations_sf)

# create an destinations df
destinations <- destinations_sf %>%
  st_drop_geometry() %>%
  transmute(
    id = as.character(id),
    lon = destinations_coords[, 1],
    lat = destinations_coords[, 2]
  )
```

```{r}
# check 
names(origins)
names(destinations)

head(origins)
head(destinations)

# check if they all return TRUE 
all(c("id", "lon", "lat") %in% names(origins))
all(c("id", "lon", "lat") %in% names(destinations))
```

**Save origins and destinations objects as RDS**

```{r}
# save as RDS 
saveRDS(origins, "data/origins_destinations/origins.rds")
saveRDS(destinations, "data/origins_destinations/destinations.rds")
```

**Map (a) Origins and (b) Destinations in one plot**

```{r}
origin_destination_map <- cowplot::plot_grid(
  london_centroids_map,
  access_point_map,
  labels = c("a", "b"),
  label_size = 14,
  label_fontface = "bold",
  label_x = c(0.02, 0.02),
  label_y = c(0.92, 0.92),
  hjust = 0,
  vjust = 1,
  ncol = 2,
  align = "h"
)

origin_destination_map
```

```{r}
# save as PNG
ggsave(
  filename = "figures/Fig5_Origin_Destinations.png",
  plot = origin_destination_map,
  width = 11.69, # keep A4 size width
  height = 5.5,
  units = "in",
  dpi = 300,
  bg = "white"
)
```

## 4. r5r

### 4.1. GTFS Data Checks

```{r}
# Load GTFS Data

# GTFS Bus Data
gtfs_bus <- read_gtfs("data/r5r/itm_london_gtfs.zip")

# GTFS Rail
gtfs_rail <- read_gtfs("data/r5r/gtfs_rail.zip")
```

```{r}
# check Bus GTFS Data

# route types
# 3 = bus, 1 = subway/tube, 2 = rail, 0 = tram
table(gtfs_bus$routes$route_type)

# service providers (agency)
unique(gtfs_bus$agency$agency_name)
```

```{r}
# check Rail GTFS Data

# check all rail agencies/ providers
unique(gtfs_rail$agency$agency_name)
```

```{r}
# within Rail GTFS
# check specifically what TfL Rail includes
gtfs_rail$routes %>%
  left_join(gtfs_rail$agency, by = "agency_id") %>%
  filter(grepl("TfL Rail", agency_name)) %>%
  select(route_short_name, route_long_name, route_type)

# check Elizabeth Line stops 
gtfs_rail$stops %>%
  filter(grepl("Abbey Wood|Shenfield|Paddington|Liverpool Street|Whitechapel|Canary Wharf|Custom House|Woolwich", stop_name, ignore.case = TRUE)) %>%
  select(stop_name, stop_id) %>%
  arrange(stop_name)

# check number of trips TfL Rail operates
gtfs_rail$trips %>%
  left_join(gtfs_rail$routes, by = "route_id") %>%
  left_join(gtfs_rail$agency, by = "agency_id") %>%
  filter(grepl("TfL Rail", agency_name)) %>%
  nrow()
```

```{r}
# Check date ranges for both bus and rail GTFS
cat("Bus GTFS\n")
cat("Start date:", as.character(min(gtfs_bus$calendar$start_date)), "\n")
cat("End date:  ", as.character(max(gtfs_bus$calendar$end_date)), "\n")

cat("\nRail GTFS\n")
cat("Start date:", as.character(min(gtfs_rail$calendar$start_date)), "\n")
cat("End date:  ", as.character(max(gtfs_rail$calendar$end_date)), "\n")
```

### 4.2. Build r5r network

#### 4.2.1. Setup

```{r}
install.packages('rJavaEnv')
library(rJavaEnv)

# check version of Java currently installed
rJavaEnv::java_check_version_rjava()

# install Java 21
rJavaEnv::java_quick_install(version = 21)
```

**Read in Origins and Destinations again:**

```{r}
origins1 <- readRDS("data/origins_destinations/origins.rds")
destinations1 <- readRDS("data/origins_destinations/destinations.rds")
```

```{r, eval=FALSE}
# check again after loading the saved RDS
nrow(origins1)
nrow(destinations1)
```

#### 4.2.2. Build r5r network

Note\* this section can be skipped once r5r network has been built, using the built network only requires loading the r5r folder within data directory. This section's code chunks are all labelled for `eval=FALSE`.

```{r, eval=FALSE}
# check package version
packageVersion("r5r")
```

```{r, eval=FALSE}
# increase Java memory
options(java.parameters = "-Xmx12G")

# load r5r library - load after loading rJavaEnv
library(r5r)

#path <- system.file("r5r_path", package = "r5r")
r5r_core <- r5r::build_network(data_path = "data/r5r")
```

```{r, eval=FALSE}
# high priority GTFS errors found
# https://ipeagit.github.io/r5r/reference/get_gtfs_errors.html 
# check the error from building network
r5r::get_gtfs_errors(data_path)
head(r5r::get_gtfs_errors(data_path), 20)
```

```{r, eval=FALSE}
# check which GTFS errored
feeds <- c("data/r5r/gtfs_rail.zip", 
           "data/r5r/itm_london_gtfs.zip")  

# check for each GTFS feed
for (f in feeds) {
  # read GTFS 
  gtfs <- read_gtfs(f)
  # skip to next if there is no transfers.txt
  if (is.null(gtfs$transfers)) {
    message(basename(f), ": no transfers.txt, nothing to fix")
    next
  }
  # check for valid stops id within stops.txt
  valid_stops <- gtfs$stops$stop_id
  # check for invalid transfers in from or to stop id 
  invalid_transfers <- gtfs$transfers[
    !(from_stop_id %in% valid_stops) | !(to_stop_id %in% valid_stops)
  ]
  message(basename(f), ": ", nrow(invalid_transfers), " broken transfer rows")
  # keep only transfers when both from and to stop id exists
  if (nrow(invalid_transfers) > 0) {
    gtfs$transfers <- gtfs$transfers[
      from_stop_id %in% valid_stops & to_stop_id %in% valid_stops
    ]
    # rewrite this revised version as new zip file
    fixed_name <- sub("\\.zip$", "_fixed.zip", f)
    write_gtfs(gtfs, fixed_name)
    message("  wrote ", basename(fixed_name))
  }
}
```

```{r, eval=FALSE}
# move the old zip out of r5r folder
file.rename("data/r5r/gtfs_rail.zip", 
            "data/raw_data/r5r_raw/gtfs_rail.zip")

# remove any leftover files from the failed built network
file.remove(list.files("data/r5r", pattern = "network.dat|mapdb", full.names = TRUE))
```

```{r, eval=FALSE}
# rebuild 
r5r_core <- r5r::build_network(data_path = "data/r5r")
```

```{r, eval=FALSE}
# high priority GTFS errors found
# check the error from building network
list.files("data/r5r")

errors <- r5r::get_gtfs_errors("data/r5r")
errors[errors$priority == "HIGH", ]
```

```{r, eval=FALSE}
# check which GTFS errored
feeds <- c("data/r5r/gtfs_rail_fixed.zip", 
           "data/r5r/itm_london_gtfs.zip")

# check for each GTFS feed
for (f in feeds) {
  gtfs <- read_gtfs(f)
  # take valid stops id within stops.txt
  valid_stops <- gtfs$stops$stop_id
  # check for stop id in trips that is not in stops.txt
  invalid_trips <- unique(
    gtfs$stop_times[
      !(stop_id %in% valid_stops)
      ]$trip_id
    )
  message(basename(f), ": ", length(invalid_trips), " trips reference missing stops")
  # if invalid trips exist remove them and 
  # rewrite this revised version as new zip file 
  if (length(invalid_trips) > 0) {
    gtfs <- filter_by_trip_id(gtfs, invalid_trips, keep = FALSE)
    fixed_name <- sub("\\.zip$", "_fix2.zip", f)
    write_gtfs(gtfs, fixed_name)
    message("  wrote ", basename(fixed_name))
  }
}
```

```{r, eval=FALSE}
# move the first errored rail feed outside of the folder
file.rename("data/r5r/gtfs_rail_fixed.zip", "data/raw_data/r5r_raw/gtfs_rail_fixed.zip")

# clear away any leftover files from the failed built network
file.remove(list.files("data/r5r", pattern = "mapdb|gtfs_errors|r5r-log|network.dat", full.names = TRUE))
```

```{r}
# final build of r5r network
r5r_core <- r5r::build_network(data_path = "data/r5r")
```

#### 4.2.3. Calculate Travel Time

+----------------------+------------------------------------------------------------------------------------------+
| Arguements           | Link                                                                                     |
+======================+==========================================================================================+
| Detailed Itineraries | <https://ipeagit.github.io/r5r/reference/detailed_itineraries.html>                      |
|                      |                                                                                          |
|                      | <https://www.rdocumentation.org/packages/r5r/versions/2.2.0/topics/detailed_itineraries> |
+----------------------+------------------------------------------------------------------------------------------+

```{r}
# maximum trip durations
max_trip_duration120 <- 120 
```

```{r}
# can change to just a day (no time) and it will return min time of the entire day
# 07/07/2026 no strike no abnormal events
departure_datetime <- as.POSIXct("07-07-2026 07:00:00", 
                                 format = "%d-%m-%Y %H:%M:%S",
                                 tz = "Europe/London")
```

**Public Transit**

```{r}
# travel by (all kinds) public transport available
mode_transit <- ("TRANSIT")

# estimating the travel time
TTM_PT <- travel_time_matrix(r5r_core,
                             origins = origins1,
                             destinations = destinations1,
                             mode = mode_transit,
                             departure_datetime = departure_datetime,
                             max_trip_duration = max_trip_duration120,
                             time_window = 180,
                             max_rides = 3,
                             # percentile (default is 50) 
                             percentiles = 50)
```

```{r, eval=FALSE}
# check
nrow(TTM_PT)
head(TTM_PT)
```

```{r, eval=FALSE}
# save RDS
saveRDS(TTM_PT, "data/r5r_TTM/r5r_TTM_PT_raw.rds")
```

**Walk**

```{r}
# travel by walking only 
mode_walk <- c("WALK")

# estimating the travel time
TTM_walk <- travel_time_matrix(r5r_core,
                               origins = origins1,
                               destinations = destinations1,
                               mode = mode_walk,
                               departure_datetime = departure_datetime,
                               max_trip_duration = max_trip_duration120,
                               time_window = 180,
                               walk_speed = 5.15,
                               percentiles = 50)
```

```{r, eval=FALSE}
# check
nrow(TTM_walk)
head(TTM_walk)
```

```{r, eval=FALSE}
# save as RDS
saveRDS(TTM_walk, "data/r5r_TTM/r5r_TTM_walk_raw.rds")
```

**Cycle**

```{r, eval=FALSE}
# travel by cycling
mode_cycle <- c("BICYCLE")

# estimating the travel time
TTM_cycle <- travel_time_matrix(r5r_core,
                                origins = origins1,
                                destinations = destinations1,
                                mode = mode_cycle,
                                departure_datetime = departure_datetime,
                                max_trip_duration = max_trip_duration120,
                                time_window = 180,
                                percentiles = 50,
                                # Bike speed (default is 12)
                                bike_speed = 12,
                                # Level of Stress (default is 2)
                                max_lts = 2)
```

```{r, eval=FALSE}
# check
nrow(TTM_cycle)
head(TTM_cycle)
```

```{r, eval=FALSE}
# save as RDS
saveRDS(TTM_cycle, "data/r5r_TTM/r5r_TTM_cycle_raw.rds")
```

```{r, eval=FALSE}
# check
nrow(TTM_PT)
nrow(TTM_walk)
nrow(TTM_cycle)
```

```         
Pairs of O-D
PT = 20023385 
Walking = 3497514 
Cycling = 8288685
```

### 4.3. Accessibility Indicators

**Load the TTM objects**

```{r}
# read in the TTM produced by r5r above
TTM_PT <- readRDS("data/r5r_TTM/r5r_TTM_PT_raw.rds")
TTM_walk <- readRDS("data/r5r_TTM/r5r_TTM_walk_raw.rds")
TTM_cycle <- readRDS("data/r5r_TTM/r5r_TTM_cycle_raw.rds")
```

```{r}
# check cooling effective UGS access point column names
names(cooling_effective_access_points)
```

```{r}
# join access point id to match the parent greenspace site
access_point_find <- cooling_effective_access_points %>% 
  st_drop_geometry() %>%
  transmute(
    id = as.character(id),
    refToGSite = as.character(refToGSite)
  ) %>%
  distinct(id, refToGSite)

# create function
join_to_site <- function(ttm) {
  ttm %>%
    mutate(
      from_id = as.character(from_id),
      to_id = as.character(to_id)
    ) %>%
    # join parent greenspace site with access point id
    left_join(
      access_point_find,
      by = c("to_id" = "id")
    )
}

# apply function to all travel modes
TTM_PT <- join_to_site(TTM_PT)
TTM_walk <- join_to_site(TTM_walk)
TTM_cycle <- join_to_site(TTM_cycle)
```

```{r}
# check
head(TTM_PT)
head(TTM_walk)
head(TTM_cycle)
```

```{r}
# minimum travel time across all of access points of each UGS

# create function
summarise_to_ugs <- function(ttm) {
  ttm %>%
    # keep access points where it has a parent greenspace site matched
    filter(!is.na(refToGSite)) %>%
    # group by origin LSOA from id and parent greenspace site
    group_by(from_id, refToGSite) %>%
    # take the minimum travel time across all access points
    summarise(travel_time_p50 = min(travel_time_p50, na.rm = TRUE),
              .groups = "drop")
}

# apply function
TTM_PT1 <- summarise_to_ugs(TTM_PT)
TTM_walk1 <- summarise_to_ugs(TTM_walk)
TTM_cycle1 <- summarise_to_ugs(TTM_cycle)
```

```{r}
# check
names(TTM_PT1)
names(TTM_walk1)
names(TTM_cycle1)
```

```{r}
# cumulative opportunity score
calc_accessibility_score <- function(lsoa_ugs_ttm, mode_name) {
  lsoa_ugs_ttm %>%
    group_by(from_id) %>%
    # count number of cooling-effective UGS reachable within each threshold (15/60)
    summarise(
      access_score_15 = sum(travel_time_p50 <= 15, na.rm = TRUE),
      access_score_60 = sum(travel_time_p50 <= 60, na.rm = TRUE),
      .groups = "drop"
    ) %>%
    # add travel mode to variable names for easier reading
    rename_with(
      ~ paste0(.x, "_", mode_name),
      .cols = c(access_score_15, access_score_60)
    )
}
```

```{r}
# calculate cumulative accessibility for each travel mode

# Public transit
PT_accessibility_score <- calc_accessibility_score(TTM_PT1, "PT")
# Walking
walking_accessibility_score <- calc_accessibility_score(TTM_walk1, "walk")
# Cycling
cycling_accessibility_score <- calc_accessibility_score(TTM_cycle1, "cycle")
```

```{r}
# check
names(PT_accessibility_score)
names(walking_accessibility_score)
names(cycling_accessibility_score)
```

```{r}
# check 
head(PT_accessibility_score)
nrow(PT_accessibility_score)
summary(PT_accessibility_score)
```

**Join all together**

```{r}
# combine cumulative accessibility across all travel modes
accessibility_score_all <- PT_accessibility_score %>%
  full_join(walking_accessibility_score, by = "from_id") %>%
  full_join(cycling_accessibility_score, by = "from_id") %>%
  # missing values present as 0
  mutate(across(starts_with("access_score"), ~replace_na(.x, 0)))
```

```{r}
# check
nrow(accessibility_score_all)
names(accessibility_score_all)
head(accessibility_score_all)
```

```{r}
# check
summary(accessibility_score_all)
```

Identified two LSOAs to have 0 values, considered acceptable as it may lie inside a CE UGS already.

### 4.4. Minimum Time to Nearest CE UGS

```{r}
# calculate min travel time to nearest cooling-effective park for each origin
calc_min_time <- function(ttm_ugs, mode_name) {
  ttm_ugs %>%
    group_by(from_id) %>%
    # create minimum travel time variable
    summarise(
      !!paste0("travel_time_min_", mode_name) := min(travel_time_p50, na.rm = TRUE),
      .groups = "drop"
    )
}

# use TTM_mode1 which is already summarised to quickest access point for each cooling effective UGS
PT_min_time <- calc_min_time(TTM_PT1, "PT")
walking_min_time <- calc_min_time(TTM_walk1, "walk")
cycling_min_time <- calc_min_time(TTM_cycle1, "cycle")
```

```{r}
# join cumulative opportunity variables with minimum travel time variables
accessibility <- accessibility_score_all %>%
  full_join(PT_min_time, by = "from_id") %>%
  full_join(walking_min_time, by = "from_id") %>%
  full_join(cycling_min_time, by = "from_id")
```

```{r}
# check
nrow(accessibility)
names(accessibility)
head(accessibility)
```

### 4.5. Join LSOA

```{r}
# join multimodal accessibility indicators back with LSOAs
london_lsoa_accessibility <- london_lsoa %>%
  left_join(
    accessibility,
    by = c("LSOA21CD" = "from_id")
  ) %>%
  mutate(across(starts_with("access_score"), ~replace_na(.x, 0)))
```

```{r}
# check
nrow(london_lsoa_accessibility)
names(london_lsoa_accessibility)
head(london_lsoa_accessibility)
```

```{r, eval=FALSE}
# save as geopackage cus names too long
st_write(london_lsoa_accessibility,"data/multimodal_accessibility_lsoa.gpkg")
```
