---
title: "5_Multimodal_Travel_Time"
---

This markdown file contains the code for ......

```{r}
# remove objects from R environment and clear memory
rm(list = ls())    
gc()
```

## Load Packages

```{r, message=FALSE, warning=FALSE}
library(sf)
library(here)
library(dplyr)
library(ggplot2)
library(tidyverse)
library(tidytransit)
```

## Data Loading

**Greater London Boundary**

```{r}
london_boundary <- st_read("data/gla/London_GLA_Boundary.shp")
```

**London LSOA Boundaries**

```{r}
london_lsoa <- st_read("data/lsoa_london/lsoa_london.shp")
```

**Green spaces \> 2ha**

```{r}
greenspace2ha <- st_read("data/greenspace_london_2ha/greenspace_london_2ha.shp")
```

**Cooling Effective UGS shp** (exported from 4_Quantify_Urban_Heat_Park_Cooling.Rmd)\_

```{r}
# Load cooling-effective greenspace shapefile
cooling_effective_UGS <- st_read("data/cooling_effective_UGS/cooling_effective_UGS.shp") %>%
  st_transform(.,27700)
```

**OS Open Greenspace Access Points**

```{r}
greenspace_access_points <- st_read("data/raw_data/OS-Open-Greenspace-GB/GB_AccessPoint/GB_AccessPoint.shp") 
```

**London LSOA Population Weighted Centroids**

```{r}
london_lsoa_pop_weighted_centroids <- st_read("data/london_lsoa_pop_weighted_centroids/london_lsoa_pop_weighted_centroids.shp")
```

## 1. Origins

```{r}
london_centroids_map <- ggplot() +
  geom_sf(
    data = london_lsoa,
    fill = NA,
    colour = "grey80"
  ) +
  geom_sf(
    data = london_boundary,
    fill = NA,
    colour = "grey30"
  ) +
  geom_sf(
    data = london_lsoa_pop_weighted_centroids,
    shape = 21,        # circle with fill + border
    fill = "#C96F57",
    colour = "black",
    stroke = 0.08,     # thin black border
    size = 0.8,
    alpha = 0.8
  ) +
  coord_sf(datum = NA) +
  theme_void() +
  theme(
    plot.background = element_rect(fill = "white", colour = NA)
  )

london_centroids_map
```

## 2. Destinations

```{r}
# check column names
names(cooling_effective_UGS)
```

### 2.1. Filter Function of Cooling Effective UGS

+----------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| Literature Support               | Reason                                                                                                                                                                                                                                         |
+==================================+================================================================================================================================================================================================================================================+
| Green Infrastructure Framework a | 1.2 Accessible Green Infrastructure (AGI) defined accessible greenspace used in Version 2.2 of the Green Infrastructure Mapping: <https://designatedsites.naturalengland.org.uk/GreenInfrastructure/UserGuide/GreenInfrastructureModules.aspx> |
+----------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| Green Infrastructure Framework b | Reasons (allotments and golf course): <https://designatedsites.naturalengland.org.uk/GreenInfrastructure/Glossary.aspx>                                                                                                                        |
+----------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| Smith et al. (2026)              | Greenspaces in the categories 'allotments or community growing spaces', 'other sports facility', 'tennis court', 'bowling green' and 'golf course' are removed                                                                                 |
|                                  |                                                                                                                                                                                                                                                |
|                                  | <https://www.sciencedirect.com/science/article/pii/S1618866725005552?__cf_chl_f_tk=kEs2lY5mgF92ex7FIa4v7E.LiQsjcC9.0QWhg03lBFk-1783176049-1.0.1.1-jCcO8_o0U3UK2CNrm0z8GHQdQ8xAGZDVW20H9a2L8.A>                                                 |
+----------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+

+----------------------------------------+--------------------+--------------------------------------------------------------------------------------------------+-----------------------------------------------------+
| Function                               |                    | Reason                                                                                           | Support                                             |
+========================================+====================+==================================================================================================+=====================================================+
| Cemetery                               | Accessible, but No |                                                                                                  | Green Infrastructure Framework                      |
+----------------------------------------+--------------------+--------------------------------------------------------------------------------------------------+-----------------------------------------------------+
| Golf Course                            | No                 | May require club membership and or payment of a fee to access (Green Infrastructure Framework b) | Green Infrastructure Framework, Smith et al. (2026) |
+----------------------------------------+--------------------+--------------------------------------------------------------------------------------------------+-----------------------------------------------------+
| Other Sports Facility                  | No                 |                                                                                                  | Green Infrastructure Framework, Smith et al. (2026) |
+----------------------------------------+--------------------+--------------------------------------------------------------------------------------------------+-----------------------------------------------------+
| Tennis Court                           | No                 |                                                                                                  | Green Infrastructure Framework, Smith et al. (2026) |
+----------------------------------------+--------------------+--------------------------------------------------------------------------------------------------+-----------------------------------------------------+
| Religious Grounds                      | Accessible, but No |                                                                                                  | Green Infrastructure Framework                      |
+----------------------------------------+--------------------+--------------------------------------------------------------------------------------------------+-----------------------------------------------------+
| Allotments or Community Growing Spaces | No                 | Normally locked and only accessible to key holders (Green Infrastructure Framework b)            | Green Infrastructure Framework, Smith et al. (2026) |
+----------------------------------------+--------------------+--------------------------------------------------------------------------------------------------+-----------------------------------------------------+

```{r}
# select UGS function and needed columns
cooling_effective_UGS_clean <- cooling_effective_UGS %>%
  # select needed columns
  select(id, functn_, geometry) %>%
  # filter park functions to Public Park or Garden and Playing Field only
  filter(functn_ %in% c("Public Park Or Garden", "Playing Field"))

# check
nrow(cooling_effective_UGS_clean)
names(cooling_effective_UGS_clean)
unique(cooling_effective_UGS_clean$functn_)
```

### 2.2. Join Filtered Cooling Effective UGS with Access Points

```{r}
nrow(greenspace_access_points)
names(greenspace_access_points)
```

```{r}
# join 
cooling_effective_access_points <- greenspace_access_points %>%
  inner_join(
    cooling_effective_UGS_clean %>%
      st_drop_geometry() %>%
      select(id, functn_),
    by = c("refToGSite" = "id")
  )
```

```{r}
nrow(cooling_effective_access_points)
```

**Map: Cooling Effective Greenspace Access Points**

```{r}
# ensure consist CRS
london_boundary <- st_transform(london_boundary,st_crs(cooling_effective_UGS_clean))

greenspace2ha <- st_transform(greenspace2ha,st_crs(cooling_effective_UGS_clean))

cooling_effective_access_points <- st_transform(cooling_effective_access_points,st_crs(cooling_effective_UGS_clean))


access_point_map <- ggplot() +
  geom_sf(
    data = london_boundary,
    fill = "white",
    colour = "grey30"
  ) +
  geom_sf(
    data = greenspace2ha,
    fill = "grey85",
    colour = NA
  ) +
  geom_sf(
    data = cooling_effective_UGS_clean,
    fill = "#CFE8D5",
    colour = NA
  ) +
  geom_sf(
    data = cooling_effective_access_points,
    shape = 25,           # upside-down triangle
    fill = "#3B5E61",      # triangle fill
    colour = "black",     # triangle border
    stroke = 0.15,        # thin border
    size = 0.8,
    alpha = 0.9
  ) +
  coord_sf(datum = NA) +
  theme_void() +
  theme(
    plot.background = element_rect(fill = "white", colour = NA)
  )

access_point_map
```

## 3. Origin and Destinations for Routing

| Routing O-D  | Objects                            |
|--------------|------------------------------------|
| Origins      | cooling_effective_access_points    |
| Destinations | london_lsoa_pop_weighted_centroids |

```{r}
# check column names now
names(cooling_effective_access_points)
names(london_lsoa_pop_weighted_centroids)
```

Save origins and destinations objects as RDS

```{r}
saveRDS(london_lsoa_pop_weighted_centroids, "data/origins_destinations/origins.rds")
saveRDS(cooling_effective_access_points, "data/origins_destinations/destinations.rds")
```

**Map (a) Origins and (b) Destinations in one plot**

```{r}
origin_destination_map <- cowplot::plot_grid(
  london_centroids_map,
  access_point_map,
  labels = c("a", "b"),
  label_size = 14,
  label_fontface = "bold",
  label_x = c(0.02, 0.02),
  label_y = c(0.92, 0.92),
  hjust = 0,
  vjust = 1,
  ncol = 2,
  align = "h"
)

origin_destination_map
```

```{r}
# save as PNG
ggsave(
  filename = "figures/Fig4_Origin_Destinations.png",
  plot = origin_destination_map,
  width = 11.69, # keep A4 size width
  height = 5.5,
  units = "in",
  dpi = 300,
  bg = "white"
)
```

## 4. r5r

### 4.1. Data Loading for r5r

```{r}
# GTFS Bus Data
gtfs_bus <- read_gtfs("data/r5r/itm_london_gtfs.zip")

# GTFS Rail
gtfs_rail <- read_gtfs("data/r5r/gtfs_rail.zip")
```

### 4.2. Data Checks

**Check Bus Open Data Service GTFS**

```{r}
# Route types
# 3 = bus, 1 = subway/tube, 2 = rail, 0 = tram
table(gtfs_bus$routes$route_type)

# Service providers (agency)
unique(gtfs_bus$agency$agency_name)
```

**Check rail GTFS**

```{r}
# Check all agencies (providers)
unique(gtfs_rail$agency$agency_name)
```

Check what TfL Rail Includes

```{r}
gtfs_rail$routes %>%
  left_join(gtfs_rail$agency, by = "agency_id") %>%
  filter(grepl("TfL Rail", agency_name)) %>%
  select(route_short_name, route_long_name, route_type)

# check lizzy line stops 
gtfs_rail$stops %>%
  filter(grepl("Abbey Wood|Shenfield|Paddington|Liverpool Street|Whitechapel|Canary Wharf|Custom House|Woolwich", stop_name, ignore.case = TRUE)) %>%
  select(stop_name, stop_id) %>%
  arrange(stop_name)

# check number of trips TfL Rail operates
gtfs_rail$trips %>%
  left_join(gtfs_rail$routes, by = "route_id") %>%
  left_join(gtfs_rail$agency, by = "agency_id") %>%
  filter(grepl("TfL Rail", agency_name)) %>%
  nrow()
```

**Check GTFS date ranges**

```{r}
# Check date ranges
cat("Bus GTFS\n")
cat("Start date:", as.character(min(gtfs_bus$calendar$start_date)), "\n")
cat("End date:  ", as.character(max(gtfs_bus$calendar$end_date)), "\n")

cat("\nRail GTFS\n")
cat("Start date:", as.character(min(gtfs_rail$calendar$start_date)), "\n")
cat("End date:  ", as.character(max(gtfs_rail$calendar$end_date)), "\n")
```

### 4.3. Build r5r network

#### 4.3.1. r5r network setup

Install Java Development Kit (JDK)

```{r, eval=FALSE}
install.packages('rJavaEnv')

# check version of Java currently installed (if any) 
rJavaEnv::java_check_version_rjava()

# install Java 21
rJavaEnv::java_quick_install(version = 21)
```

Load Origins and Destinations again:

```{r}
origins <- readRDS("data/origins_destinations/origins.rds")
destinations <- readRDS("data/origins_destinations/destinations.rds")
```

```{r}
# increase Java memory
options(java.parameters = "-Xmx12G")

# load libraries
library(r5r)

#path <- system.file("r5r_path", package = "r5r")
r5r_core <- r5r::build_network(data_path = "data/r5r")

# used setup_r5() previously and got the following: 
# Warning: ! `setup_r5()` is being deprecated in *r5r* v2.3.0 and will be removed in a future release. ℹ Please use `r5r::build_network()` instead.
```

#### 4.3.2. Set Routing Parameters

+----------------------+------------------------------------------------------------------------------------------+
| Arguements           | Link                                                                                     |
+======================+==========================================================================================+
| Detailed Itineraries | <https://ipeagit.github.io/r5r/reference/detailed_itineraries.html>                      |
|                      |                                                                                          |
|                      | <https://www.rdocumentation.org/packages/r5r/versions/2.2.0/topics/detailed_itineraries> |
+----------------------+------------------------------------------------------------------------------------------+

```{r}
# maximum trip durations to be 1 hour (60 min)
# later filter for 15, 30, 60
max_trip_duration60 <- 60 
```

```{r}
# can change to just a day (no time) and it will return min time of the entire day
# 01/07/2026 no strike no abnormal events
departure_datetime <- as.POSIXct("01-07-2026 12:00:00", 
                                 format = "%d-%m-%Y %H:%M:%S",
                                 tz = "Europe/London")
```

**Public Transit**

```{r}
# travel by (all kinds) public transport available
mode_transit <- ("TRANSIT")

# estimating the travel time
TTM_PT <- travel_time_matrix(r5r_core,
                             origins = origins,
                             destinations = destinations,
                             mode = mode_transit,
                             departure_datetime = departure_datetime,
                             max_trip_duration = max_trip_duration,
                             time_window = 120,
                             # percentile is set default for 50 (median) 
                             # but still typed it out to be clear here
                             percentiles = 50)
                             # percentiles = c(25, 50, 75))
```

**Walk**

```{r}
# travel by walking only 
mode_walk <- c("WALK")

# estimating the travel time
TTM_walk <- travel_time_matrix(r5r_core,
                               origins = origins,
                               destinations = destinations,
                               mode = mode_walk,
                               departure_datetime = departure_datetime,
                               max_trip_duration = max_trip_duration,
                               time_window = 120,
                               percentiles = 50)
```

**Cycle**

```{r}
mode_cycle <- c("BICYCLE")

# estimating the travel time
TTM_cycle <- travel_time_matrix(r5r_core,
                                origins = origins,
                                destinations = destinations,
                                mode = TTM_cycle,
                                departure_datetime = departure_datetime,
                                max_trip_duration = max_trip_duration,
                                time_window = 120,
                                percentiles = 50,
                                # Bike speed (default is 12)
                                bike_speed = 12,
                                # Level of Stress (default is 2)
                                max_lts = 2)
```

### 4.3. Minimum Travel Time

### 4.4. Filter Travel Time for 15 min, 30 min and 60 min

```{r}

```
