---
title: "5_Multimodal_Travel_Time"
---

This markdown file contains the code for ......

```{r}
# remove objects from R environment and clear memory
rm(list = ls())    
gc()
```

## Load Packages

```{r, message=FALSE, warning=FALSE}
library(sf)
library(here)
library(dplyr)
library(ggplot2)
library(tidyverse)
```

## Data Loading

**Greater London Boundary**

```{r}
london_boundary <- st_read("data/raw_data/gla/London_london_boundary.shp")
```

**London LSOA Boundaries**

```{r}
london_lsoa <- st_read("data/lsoa_london/lsoa_london.shp")
```

**Green spaces \> 2ha**

```{r}
greenspace2ha <- st_read("data/greenspace_london_2ha/greenspace_london_2ha.shp")
```

**Cooling Effective UGS shp** (exported from 4_Quantify_Urban_Heat_Park_Cooling.Rmd)\_

```{r}
# Load cooling-effective greenspace shapefile
cooling_effective_UGS <- st_read("data/cooling_effective_UGS/cooling_effective_UGS.shp") %>%
  st_transform(.,27700)
```

**OS Open Greenspace Access Points**

```{r}
greenspace_access_points <- st_read("data/raw_data/OS-Open-Greenspace-GB/GB_AccessPoint/GB_AccessPoint.shp") 
```

**London LSOA Population Weighted Centroids**

```{r}
london_lsoa_pop_weighted_centroids <- st_read("data/london_lsoa_pop_weighted_centroids/london_lsoa_pop_weighted_centroids.shp")
```

## 1. Origins

```{r}
london_centroids_map <- ggplot() +
  geom_sf(
    data = london_boundary,
    fill = "white",
    colour = "grey30"
  ) +
  geom_sf(
    data = london_lsoa,
    fill = NA,
    colour = "grey80",
    linewidth = 0.1
  ) +
  geom_sf(
    data = london_lsoa_pop_weighted_centroids,
    shape = 21,        # circle with fill + border
    fill = "#C96F57",
    colour = "black",
    stroke = 0.08,     # thin black border
    size = 0.8,
    alpha = 0.8
  ) +
  coord_sf(datum = NA) +
  theme_void() +
  theme(
    plot.background = element_rect(fill = "white", colour = NA)
  )

london_centroids_map
```

## 2. Destinations

```{r}
# check column names
names(cooling_effective_UGS)
```

### 2.1. Filter Function of Cooling Effective UGS

+-----------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| Literature Support                | Reason                                                                                                                                                                                                                                         |
+===================================+================================================================================================================================================================================================================================================+
| Green Infrastructure Framework a  | 1.2 Accessible Green Infrastructure (AGI) defined accessible greenspace used in Version 2.2 of the Green Infrastructure Mapping: <https://designatedsites.naturalengland.org.uk/GreenInfrastructure/UserGuide/GreenInfrastructureModules.aspx> |
+-----------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| Green Infrastructure Framework b  | Reasons (allotments and golf course): <https://designatedsites.naturalengland.org.uk/GreenInfrastructure/Glossary.aspx>                                                                                                                        |
+-----------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| Smith et al. (2026)               | Greenspaces in the categories 'allotments or community growing spaces', 'other sports facility', 'tennis court', 'bowling green' and 'golf course' are removed                                                                                 |
|                                   |                                                                                                                                                                                                                                                |
|                                   | <https://www.sciencedirect.com/science/article/pii/S1618866725005552?__cf_chl_f_tk=kEs2lY5mgF92ex7FIa4v7E.LiQsjcC9.0QWhg03lBFk-1783176049-1.0.1.1-jCcO8_o0U3UK2CNrm0z8GHQdQ8xAGZDVW20H9a2L8.A>                                                 |
+-----------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+

| Function |   | Reason | Support |
|----|----|----|----|
| Cemetery | Accessible, but No |  | Green Infrastructure Framework |
| Golf Course | No | May require club membership and or payment of a fee to access (Green Infrastructure Framework b) | Green Infrastructure Framework, Smith et al. (2026) |
| Other Sports Facility | No |  | Green Infrastructure Framework, Smith et al. (2026) |
| Tennis Court | No |  | Green Infrastructure Framework, Smith et al. (2026) |
| Religious Grounds | Accessible, but No |  | Green Infrastructure Framework |
| Allotments or Community Growing Spaces | No | Normally locked and only accessible to key holders (Green Infrastructure Framework b) | Green Infrastructure Framework, Smith et al. (2026) |

```{r}
# select UGS function and needed columns
cooling_effective_UGS_clean <- cooling_effective_UGS %>%
  # select needed columns
  select(id, functn_, geometry) %>%
  # filter park functions to Public Park or Garden and Playing Field only
  filter(functn_ %in% c("Public Park Or Garden", "Playing Field"))

# check
nrow(cooling_effective_UGS_clean)
names(cooling_effective_UGS_clean)
unique(cooling_effective_UGS_clean$functn_)
```

### 2.2. Join Filtered Cooling Effective UGS with Access Points

```{r}
nrow(greenspace_access_points)
names(greenspace_access_points)
```

```{r}
# join 
cooling_effective_access_points <- greenspace_access_points %>%
  inner_join(
    cooling_effective_UGS_clean %>%
      st_drop_geometry() %>%
      select(id, functn_),
    by = c("refToGSite" = "id")
  )
```

```{r}
nrow(cooling_effective_access_points)
```

**Map: Cooling Effective Greenspace Access Points**

```{r}
# ensure consist CRS
london_boundary <- st_transform(london_boundary,st_crs(cooling_effective_UGS_clean))

greenspace2ha <- st_transform(greenspace2ha,st_crs(cooling_effective_UGS_clean))

cooling_effective_access_points <- st_transform(cooling_effective_access_points,st_crs(cooling_effective_UGS_clean))


access_point_map <- ggplot() +
  geom_sf(
    data = london_boundary,
    fill = "white",
    colour = "grey30",
    linewidth = 0.3
  ) +
  geom_sf(
    data = greenspace2ha,
    fill = "grey85",
    colour = NA
  ) +
  geom_sf(
    data = cooling_effective_UGS_clean,
    fill = "#CFE8D5",
    colour = NA
  ) +
  geom_sf(
    data = cooling_effective_access_points,
    shape = 25,           # upside-down triangle
    fill = "#3B5E61",      # triangle fill
    colour = "black",     # triangle border
    stroke = 0.15,        # thin border
    size = 0.8,
    alpha = 0.9
  ) +
  coord_sf(datum = NA) +
  theme_void() +
  theme(
    plot.background = element_rect(fill = "white", colour = NA)
  )

access_point_map
```

## 3. Origin and Destinations for Routing

| Routing O-D  | Objects                            |
|--------------|------------------------------------|
| Origins      | cooling_effective_access_points    |
| Destinations | london_lsoa_pop_weighted_centroids |

**Map (a) Origins and (b) Destinations in one plot**

```{r}
origin_destination_map <- cowplot::plot_grid(
  london_centroids_map,
  access_point_map,
  labels = c("a", "b"),
  label_size = 14,
  label_fontface = "bold",
  label_x = c(0.03, 0.03),
  label_y = c(0.74, 0.74),
  hjust = 0,
  vjust = 1,
  ncol = 2,
  align = "hv"
)

origin_destination_map
```

```{r}
# save as PNG
ggsave(
  filename = "figures/Fig4_Origin_Destinations.png",
  plot = origin_destination_map,
  width = 11.69,
  height = 8.27,
  units = "in",
  dpi = 300,
  bg = "white"
)
```
