---
title: "5_Multimodal_Travel_Time"
---

This markdown file contains the code for ......

```{r}
# remove objects from R environment and clear memory
rm(list = ls())    
gc()
```

## Load Packages

```{r, message=FALSE, warning=FALSE}
library(sf)
library(here)
library(dplyr)
library(ggplot2)
library(tidyverse)
library(tidytransit)
library(gtfstools)
```

## Data Loading

**Greater London Boundary**

```{r}
london_boundary <- st_read("data/gla/London_GLA_Boundary.shp")
```

**London LSOA Boundaries**

```{r}
london_lsoa <- st_read("data/lsoa_london/lsoa_london.shp")
```

**Green spaces \> 2ha**

```{r}
greenspace2ha <- st_read("data/greenspace_london_2ha/greenspace_london_2ha.shp")
```

**Cooling Effective UGS shp** (exported from 4_Quantify_Urban_Heat_Park_Cooling.Rmd)\_

```{r}
# Load cooling-effective greenspace shapefile
cooling_effective_UGS <- st_read("data/cooling_effective_UGS/cooling_effective_UGS.shp") %>%
  st_transform(.,27700)
```

**OS Open Greenspace Access Points**

```{r}
greenspace_access_points <- st_read("data/raw_data/OS-Open-Greenspace-GB/GB_AccessPoint/GB_AccessPoint.shp") 
```

**London LSOA Population Weighted Centroids**

```{r}
london_lsoa_pop_weighted_centroids <- st_read("data/london_lsoa_pop_weighted_centroids/london_lsoa_pop_weighted_centroids.shp")
```

## 1. Origins

**London LSOA population weighted centroids**

```{r}
london_centroids_map <- ggplot() +
  geom_sf(
    data = london_lsoa,
    fill = NA,
    colour = "grey80"
  ) +
  geom_sf(
    data = london_boundary,
    fill = NA,
    colour = "grey30"
  ) +
  geom_sf(
    data = london_lsoa_pop_weighted_centroids,
    shape = 21,        # circle with fill + border
    fill = "#C96F57",
    colour = "black",
    stroke = 0.08,     # thin black border
    size = 0.8,
    alpha = 0.8
  ) +
  coord_sf(datum = NA) +
  theme_void() +
  theme(
    plot.background = element_rect(fill = "white", colour = NA)
  )

london_centroids_map
```

```{r}
# map with legend
london_centroids_map <- ggplot() +
  geom_sf(
    data = london_lsoa,
    fill = NA,
    colour = "grey80"
  ) +
  geom_sf(
    data = london_boundary,
    fill = NA,
    colour = "grey30"
  ) +
  geom_sf(
    data = london_lsoa_pop_weighted_centroids,
    aes(fill = "Population Weighted Centroids"),
    shape = 21,
    colour = "black",
    stroke = 0.08,
    size = 0.8,
    alpha = 0.8
  ) +
  scale_fill_manual(
    values = c("Population Weighted Centroids" = "#C96F57"),
    name = NULL
  ) +
  guides(
    fill = guide_legend(
      override.aes = list(
        shape = 21,
        colour = "black",
        size = 3,
        stroke = 0.2,
        alpha = 0.8
      )
    )
  ) +
  coord_sf(datum = NA) +
  theme_void() +
  theme(
    plot.background = element_rect(fill = "white", colour = NA),
    legend.position = "bottom",
    legend.justification = "center",
    legend.text = element_text(size = 10),
    legend.margin = margin(t = 2, b = 2)
  )

london_centroids_map
```

## 2. Destinations

```{r}
# check column names
names(cooling_effective_UGS)
```

### 2.1. Filter Function of Cooling Effective UGS

+----------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| Literature Support               | Reason                                                                                                                                                                                                                                         |
+==================================+================================================================================================================================================================================================================================================+
| Green Infrastructure Framework a | 1.2 Accessible Green Infrastructure (AGI) defined accessible greenspace used in Version 2.2 of the Green Infrastructure Mapping: <https://designatedsites.naturalengland.org.uk/GreenInfrastructure/UserGuide/GreenInfrastructureModules.aspx> |
+----------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| Green Infrastructure Framework b | Reasons (allotments and golf course): <https://designatedsites.naturalengland.org.uk/GreenInfrastructure/Glossary.aspx>                                                                                                                        |
+----------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| Smith et al. (2026)              | Greenspaces in the categories 'allotments or community growing spaces', 'other sports facility', 'tennis court', 'bowling green' and 'golf course' are removed                                                                                 |
|                                  |                                                                                                                                                                                                                                                |
|                                  | <https://www.sciencedirect.com/science/article/pii/S1618866725005552?__cf_chl_f_tk=kEs2lY5mgF92ex7FIa4v7E.LiQsjcC9.0QWhg03lBFk-1783176049-1.0.1.1-jCcO8_o0U3UK2CNrm0z8GHQdQ8xAGZDVW20H9a2L8.A>                                                 |
+----------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+

+----------------------------------------+--------------------+--------------------------------------------------------------------------------------------------+-----------------------------------------------------+
| Function                               |                    | Reason                                                                                           | Support                                             |
+========================================+====================+==================================================================================================+=====================================================+
| Cemetery                               | Accessible, but No |                                                                                                  | Green Infrastructure Framework                      |
+----------------------------------------+--------------------+--------------------------------------------------------------------------------------------------+-----------------------------------------------------+
| Golf Course                            | No                 | May require club membership and or payment of a fee to access (Green Infrastructure Framework b) | Green Infrastructure Framework, Smith et al. (2026) |
+----------------------------------------+--------------------+--------------------------------------------------------------------------------------------------+-----------------------------------------------------+
| Other Sports Facility                  | No                 |                                                                                                  | Green Infrastructure Framework, Smith et al. (2026) |
+----------------------------------------+--------------------+--------------------------------------------------------------------------------------------------+-----------------------------------------------------+
| Tennis Court                           | No                 |                                                                                                  | Green Infrastructure Framework, Smith et al. (2026) |
+----------------------------------------+--------------------+--------------------------------------------------------------------------------------------------+-----------------------------------------------------+
| Religious Grounds                      | Accessible, but No |                                                                                                  | Green Infrastructure Framework                      |
+----------------------------------------+--------------------+--------------------------------------------------------------------------------------------------+-----------------------------------------------------+
| Allotments or Community Growing Spaces | No                 | Normally locked and only accessible to key holders (Green Infrastructure Framework b)            | Green Infrastructure Framework, Smith et al. (2026) |
+----------------------------------------+--------------------+--------------------------------------------------------------------------------------------------+-----------------------------------------------------+

```{r}

# select UGS function and needed columns
cooling_effective_UGS_clean <- cooling_effective_UGS %>%
  # select needed columns
  select(id, functn_, geometry) %>%
  # filter park functions to Public Park or Garden and Playing Field only
  # reasons refer to above table
  filter(functn_ %in% c("Public Park Or Garden", "Playing Field"))

# check
nrow(cooling_effective_UGS_clean)
names(cooling_effective_UGS_clean)
unique(cooling_effective_UGS_clean$functn_)
```

### 2.2. Join Filtered Cooling Effective UGS with Access Points

```{r}
# check
nrow(greenspace_access_points)
names(greenspace_access_points)
```

```{r}
# join 
cooling_effective_access_points <- greenspace_access_points %>%
  inner_join(
    cooling_effective_UGS_clean %>%
      st_drop_geometry() %>%
      select(id, functn_),
    by = c("refToGSite" = "id")
  )
```

```{r}
# check
# number of access points
nrow(cooling_effective_access_points)
```

**Map: Cooling Effective Greenspace Access Points**

```{r, eval=FALSE}
# ensure consist CRS
london_boundary <- st_transform(london_boundary,st_crs(cooling_effective_UGS_clean))

greenspace2ha <- st_transform(greenspace2ha,st_crs(cooling_effective_UGS_clean))

cooling_effective_access_points <- st_transform(cooling_effective_access_points,st_crs(cooling_effective_UGS_clean))


access_point_map <- ggplot() +
  geom_sf(
    data = london_boundary,
    fill = "white",
    colour = "grey30"
  ) +
  geom_sf(
    data = greenspace2ha,
    fill = "grey85",
    colour = NA
  ) +
  geom_sf(
    data = cooling_effective_UGS_clean,
    fill = "#CFE8D5",
    colour = NA
  ) +
  geom_sf(
    data = cooling_effective_access_points,
    shape = 25,           # upside-down triangle
    fill = "#3B5E61",      # triangle fill
    colour = "black",     # triangle border
    stroke = 0.15,        # thin border
    size = 0.8,
    alpha = 0.9
  ) +
  coord_sf(datum = NA) +
  theme_void() +
  theme(
    plot.background = element_rect(fill = "white", colour = NA)
  )

access_point_map
```

```{r}
# map with legend
# ensure consist CRS
london_boundary <- st_transform(london_boundary, st_crs(cooling_effective_UGS_clean))
greenspace2ha <- st_transform(greenspace2ha, st_crs(cooling_effective_UGS_clean))
cooling_effective_access_points <- st_transform(cooling_effective_access_points, st_crs(cooling_effective_UGS_clean))

access_point_map <- ggplot() +
  geom_sf(
    data = london_boundary,
    fill = "white",
    colour = "grey30"
  ) +
  geom_sf(
    data = greenspace2ha,
    fill = "grey85",
    colour = NA
  ) +
  geom_sf(
    data = cooling_effective_UGS_clean,
    fill = "#CFE8D5",
    colour = NA
  ) +
  geom_sf(
    data = cooling_effective_access_points,
    aes(fill = "Access Points"),
    shape = 25,              # upside-down triangle
    colour = "black",
    stroke = 0.15,
    size = 0.8,
    alpha = 0.9
  ) +
  scale_fill_manual(
    values = c("Access Points" = "#3B5E61"),
    name = NULL
  ) +
  guides(
    fill = guide_legend(
      override.aes = list(
        shape = 25,
        colour = "black",
        size = 3,
        stroke = 0.25,
        alpha = 0.9
      )
    )
  ) +
  coord_sf(datum = NA) +
  theme_void() +
  theme(
    plot.background = element_rect(fill = "white", colour = NA),
    legend.position = "bottom",
    legend.justification = "center",
    legend.text = element_text(size = 10),
    legend.margin = margin(t = 2, b = 2)
  )

access_point_map
```

```{r}
# map with legend
# ensure consist CRS
london_boundary <- st_transform(london_boundary, st_crs(cooling_effective_UGS_clean))
greenspace2ha <- st_transform(greenspace2ha, st_crs(cooling_effective_UGS_clean))
cooling_effective_access_points <- st_transform(cooling_effective_access_points, st_crs(cooling_effective_UGS_clean))

access_point_map <- ggplot() +
  geom_sf(
    data = london_boundary,
    fill = "white",
    colour = "grey30"
  ) +
  geom_sf(
    data = greenspace2ha,
    fill = "grey85",
    colour = NA
  ) +
  geom_sf(
    data = cooling_effective_UGS_clean,
    fill = "#B8DDDA",   # light teal
    colour = NA
  ) +
  geom_sf(
    data = cooling_effective_access_points,
    aes(fill = "Access Points"),
    shape = 25,              # upside-down triangle
    colour = "black",
    stroke = 0.15,
    size = 0.8,
    alpha = 0.9
  ) +
  scale_fill_manual(
    values = c("Access Points" = "#008F8C"),   # dark teal
    name = NULL
  ) +
  guides(
    fill = guide_legend(
      override.aes = list(
        shape = 25,
        colour = "black",
        size = 3,
        stroke = 0.25,
        alpha = 0.9
      )
    )
  ) +
  coord_sf(datum = NA) +
  theme_void() +
  theme(
    plot.background = element_rect(fill = "white", colour = NA),
    legend.position = "bottom",
    legend.justification = "center",
    legend.text = element_text(size = 10),
    legend.margin = margin(t = 2, b = 2)
  )

access_point_map
```

## 3. Prepare Origin and Destinations for Routing

| Routing O-D  | Objects                            |
|--------------|------------------------------------|
| Origins      | london_lsoa_pop_weighted_centroids |
| Destinations | cooling_effective_access_points    |

```{r}
# check column names
names(london_lsoa_pop_weighted_centroids)
names(cooling_effective_access_points)
```

```{r}
origins_sf <- london_lsoa_pop_weighted_centroids %>%
  st_transform(4326)

origins_coords <- st_coordinates(origins_sf)

origins <- origins_sf %>%
  st_drop_geometry() %>%
  transmute(
    id = as.character(LSOA21CD),
    lon = origins_coords[, 1],
    lat = origins_coords[, 2]
  )

destinations_sf <- cooling_effective_access_points %>%
  st_transform(4326)

destinations_coords <- st_coordinates(destinations_sf)

destinations <- destinations_sf %>%
  st_drop_geometry() %>%
  transmute(
    id = as.character(id),
    lon = destinations_coords[, 1],
    lat = destinations_coords[, 2]
  )
```

```{r}
names(origins)
names(destinations)

head(origins)
head(destinations)

# These should both return TRUE
all(c("id", "lon", "lat") %in% names(origins))
all(c("id", "lon", "lat") %in% names(destinations))
```

Save origins and destinations objects as RDS

```{r, eval=FALSE}
saveRDS(origins, "data/origins_destinations/origins.rds")
saveRDS(destinations, "data/origins_destinations/destinations.rds")
```

**Map (a) Origins and (b) Destinations in one plot**

```{r}
origin_destination_map <- cowplot::plot_grid(
  london_centroids_map,
  access_point_map,
  labels = c("a", "b"),
  label_size = 14,
  label_fontface = "bold",
  label_x = c(0.02, 0.02),
  label_y = c(0.92, 0.92),
  hjust = 0,
  vjust = 1,
  ncol = 2,
  align = "h"
)

origin_destination_map
```

```{r}
# save as PNG
ggsave(
  filename = "figures/Fig5_Origin_Destinations.png",
  plot = origin_destination_map,
  width = 11.69, # keep A4 size width
  height = 5.5,
  units = "in",
  dpi = 300,
  bg = "white"
)
```

## 4. Visualise Road Network

## 5. r5r

### 5.1. GTFS Data Checks

```{r}
# GTFS Bus Data
gtfs_bus <- read_gtfs("data/r5r/itm_london_gtfs.zip")

# GTFS Rail
gtfs_rail <- read_gtfs("data/r5r/gtfs_rail.zip")
```

**Check Bus Open Data Service GTFS**

```{r}
# Route types
# 3 = bus, 1 = subway/tube, 2 = rail, 0 = tram
table(gtfs_bus$routes$route_type)

# Service providers (agency)
unique(gtfs_bus$agency$agency_name)
```

**Check rail GTFS**

```{r}
# Check all agencies (providers)
unique(gtfs_rail$agency$agency_name)
```

Check what TfL Rail Includes

```{r}
gtfs_rail$routes %>%
  left_join(gtfs_rail$agency, by = "agency_id") %>%
  filter(grepl("TfL Rail", agency_name)) %>%
  select(route_short_name, route_long_name, route_type)

# check lizzy line stops 
gtfs_rail$stops %>%
  filter(grepl("Abbey Wood|Shenfield|Paddington|Liverpool Street|Whitechapel|Canary Wharf|Custom House|Woolwich", stop_name, ignore.case = TRUE)) %>%
  select(stop_name, stop_id) %>%
  arrange(stop_name)

# check number of trips TfL Rail operates
gtfs_rail$trips %>%
  left_join(gtfs_rail$routes, by = "route_id") %>%
  left_join(gtfs_rail$agency, by = "agency_id") %>%
  filter(grepl("TfL Rail", agency_name)) %>%
  nrow()
```

**Check GTFS date ranges**

```{r}
# Check date ranges
cat("Bus GTFS\n")
cat("Start date:", as.character(min(gtfs_bus$calendar$start_date)), "\n")
cat("End date:  ", as.character(max(gtfs_bus$calendar$end_date)), "\n")

cat("\nRail GTFS\n")
cat("Start date:", as.character(min(gtfs_rail$calendar$start_date)), "\n")
cat("End date:  ", as.character(max(gtfs_rail$calendar$end_date)), "\n")
```

### 5.2. Build r5r network

#### 5.2.1. r5r network setup

```{r, eval=FALSE}
#install.packages('rJavaEnv')

# check version of Java currently installed (if any) 
#rJavaEnv::java_check_version_rjava()

# install Java 21
#rJavaEnv::java_quick_install(version = 21)
```

**Check Java Environment**

```{r}
library(rJavaEnv)
rJavaEnv::java_check_version_rjava()
```

**Read in Origins and Destinations again:**

```{r}
origins1 <- readRDS("data/origins_destinations/origins.rds")
destinations1 <- readRDS("data/origins_destinations/destinations.rds")
```

```{r}
# check again after loading the saved RDS
nrow(origins1)
nrow(destinations1)
```

#### 5.2.2. r5r network

(this part can be skipped after creating the r5r network folder once)

```{r}
# check package version
packageVersion("r5r")
```

```{r}
# increase Java memory
options(java.parameters = "-Xmx12G")

# load libraries
library(r5r)

#path <- system.file("r5r_path", package = "r5r")
r5r_core <- r5r::build_network(data_path = "data/r5r")

# used setup_r5() previously and got the following: 
# Warning: ! `setup_r5()` is being deprecated in *r5r* v2.3.0 and will be removed in a future release. ℹ Please use `r5r::build_network()` instead.
```

```{r}
# Error message: 

#Error in `r5r::build_network()`:
#✖ High priority GTFS errors found; network build failed.
#ℹ Use `r5r::get_gtfs_errors()` to see them, or review data/r5r/gtfs_errors.csv.
#Run `rlang::last_trace()` to see where the error occurred.
```

```{r}
# check what the error is 
r5r::get_gtfs_errors(data_path)

head(r5r::get_gtfs_errors(data_path), 20)
```

```{r}
# check which GTFS errored
# create a fixed gtfs rail zip --> named : gtfs_rail_fixed
feeds <- c("data/r5r/gtfs_rail.zip", "data/r5r/itm_london_gtfs.zip")  

for (f in feeds) {
  gtfs <- read_gtfs(f)

  if (is.null(gtfs$transfers)) {
    message(basename(f), ": no transfers.txt, nothing to fix")
    next
  }

  valid_stops <- gtfs$stops$stop_id
  bad <- gtfs$transfers[
    !(from_stop_id %in% valid_stops) | !(to_stop_id %in% valid_stops)
  ]
  message(basename(f), ": ", nrow(bad), " broken transfer rows")

  if (nrow(bad) > 0) {
    gtfs$transfers <- gtfs$transfers[
      from_stop_id %in% valid_stops & to_stop_id %in% valid_stops
    ]
    #rewrite a removed error new zip file
    fixed_name <- sub("\\.zip$", "_fixed.zip", f)
    write_gtfs(gtfs, fixed_name)
    message("  wrote ", basename(fixed_name))
  }
}
```

```{r}
# move the old zip out of r5r folder
file.rename("data/r5r/gtfs_rail.zip", "data/raw_data/r5r_raw/gtfs_rail.zip")

# remove any leftovers from the failed build
file.remove(list.files("data/r5r", pattern = "network.dat|mapdb", full.names = TRUE))
```

```{r}
# rebuild
r5r_core <- r5r::build_network(data_path = "data/r5r")
```

```{r}
# error again
#Error in `r5r::build_network()`:
#✖ High priority GTFS errors found; network build failed.
#ℹ Use `r5r::get_gtfs_errors()` to see them, or review data/r5r/gtfs_errors.csv.
#Run `rlang::last_trace()` to see where the error occurred.
```

```{r}
#check again where is the error
list.files("data/r5r")

errors <- r5r::get_gtfs_errors("data/r5r")
errors[errors$priority == "HIGH", ]
```

```{r}
feeds <- c("data/r5r/gtfs_rail_fixed.zip", "data/r5r/itm_london_gtfs.zip")

for (f in feeds) {
  gtfs <- read_gtfs(f)
  valid_stops <- gtfs$stops$stop_id

  bad_trips <- unique(gtfs$stop_times[!(stop_id %in% valid_stops)]$trip_id)
  message(basename(f), ": ", length(bad_trips), " trips reference missing stops")

  if (length(bad_trips) > 0) {
    gtfs <- filter_by_trip_id(gtfs, bad_trips, keep = FALSE)
    fixed_name <- sub("\\.zip$", "_fix2.zip", f)
    write_gtfs(gtfs, fixed_name)
    message("  wrote ", basename(fixed_name))
  }
}
```

```{r}
# move the error2 rail feed to outside gtfs_original file
file.rename("data/r5r/gtfs_rail_fixed.zip", "data/raw_data/r5r_raw/gtfs_rail_fixed.zip")

# clear caches and debris from previous attempts
file.remove(list.files("data/r5r", pattern = "mapdb|gtfs_errors|r5r-log|network.dat", full.names = TRUE))
```

```{r}
# buld r5r network again
r5r_core <- r5r::build_network(data_path = "data/r5r")
```

#### 5.2.3. Calculate Travel Time

+----------------------+------------------------------------------------------------------------------------------+
| Arguements           | Link                                                                                     |
+======================+==========================================================================================+
| Detailed Itineraries | <https://ipeagit.github.io/r5r/reference/detailed_itineraries.html>                      |
|                      |                                                                                          |
|                      | <https://www.rdocumentation.org/packages/r5r/versions/2.2.0/topics/detailed_itineraries> |
+----------------------+------------------------------------------------------------------------------------------+

```{r}
# increase Java memory
options(java.parameters = "-Xmx12G")

# load library again
library(r5r)

# build network again
r5r_core <- r5r::build_network(data_path = "data/r5r")
```

```{r}
# maximum trip durations
max_trip_duration120 <- 120 
```

```{r}
# can change to just a day (no time) and it will return min time of the entire day
# 01/07/2026 no strike no abnormal events
departure_datetime <- as.POSIXct("07-07-2026 07:00:00", 
                                 format = "%d-%m-%Y %H:%M:%S",
                                 tz = "Europe/London")
```

**Public Transit**

```{r}
# travel by (all kinds) public transport available
mode_transit <- ("TRANSIT")

# estimating the travel time
TTM_PT <- travel_time_matrix(r5r_core,
                             origins = origins1,
                             destinations = destinations1,
                             mode = mode_transit,
                             departure_datetime = departure_datetime,
                             max_trip_duration = max_trip_duration120,
                             time_window = 180,
                             max_rides = 3,
                             # percentile (default is 50) 
                             percentiles = 50)
```

```{r}
nrow(TTM_PT)
head(TTM_PT)
```

```{r}
# save RDS
saveRDS(TTM_PT, "data/r5r_TTM/r5r_TTM_PT_raw.rds")
```

**Walk**

```{r}
# travel by walking only 
mode_walk <- c("WALK")

# estimating the travel time
TTM_walk <- travel_time_matrix(r5r_core,
                               origins = origins1,
                               destinations = destinations1,
                               mode = mode_walk,
                               departure_datetime = departure_datetime,
                               max_trip_duration = max_trip_duration120,
                               time_window = 180,
                               walk_speed = 5.15,
                               percentiles = 50)
```

```{r}
nrow(TTM_walk)
head(TTM_walk)
```

```{r}
# save as RDS
saveRDS(TTM_walk, "data/r5r_TTM/r5r_TTM_walk_raw.rds")
```

**Cycle**

```{r}
mode_cycle <- c("BICYCLE")

# estimating the travel time
TTM_cycle <- travel_time_matrix(r5r_core,
                                origins = origins1,
                                destinations = destinations1,
                                mode = mode_cycle,
                                departure_datetime = departure_datetime,
                                max_trip_duration = max_trip_duration120,
                                time_window = 180,
                                percentiles = 50,
                                # Bike speed (default is 12)
                                bike_speed = 12,
                                # Level of Stress (default is 2)
                                max_lts = 2)
```

```{r}
nrow(TTM_cycle)
head(TTM_cycle)
```

```{r}
# save as RDS
saveRDS(TTM_cycle, "data/r5r_TTM/r5r_TTM_cycle_raw.rds")
```

```{r}
nrow(TTM_PT)
nrow(TTM_walk)
nrow(TTM_cycle)
```

```         
Pairs of O-D
PT = 20023385 
Walking = 3497514 
Cycling = 8288685
```

### 5.3. Minimum Travel Time

(returned the same, can skip)

Load the TTM objects

```{r}
TTM_PT <- readRDS("data/r5r_TTM/r5r_TTM_PT_raw.rds")
TTM_walk <- readRDS("data/r5r_TTM/r5r_TTM_walk_raw.rds")
TTM_cycle <- readRDS("data/r5r_TTM/r5r_TTM_cycle_raw.rds")
```

**Select minimum value between pairs**

```{r}
# public transit
TTM_PT_min_pairs <- TTM_PT %>%
  group_by(from_id, to_id) %>% 
  slice_min(travel_time_p50, n = 1, with_ties = FALSE, na_rm = FALSE) %>%
  ungroup()
```

```{r}
# walk
TTM_walk_min_pairs <- TTM_walk %>%
  group_by(from_id, to_id) %>% 
  slice_min(travel_time_p50, n = 1, with_ties = FALSE, na_rm = FALSE) %>%
  ungroup()
```

```{r}
# cycle
TTM_cycle_min_pairs <- TTM_cycle %>%
  group_by(from_id, to_id) %>% 
  slice_min(travel_time_p50, n = 1, with_ties = FALSE, na_rm = FALSE) %>%
  ungroup()
```

```{r}
# already checked nrows have the same (not changed)
# check
head(TTM_PT_min_pairs)
head(TTM_walk_min_pairs)
head(TTM_cycle_min_pairs)
```

### 5.4. Filter Minimum Travel Time for the nearest cooling effective UGS and for 15 min, 30 min and 60 min

```{r}
TTM_PT_access <- TTM_PT_min_pairs %>%
  group_by(from_id) %>%
  slice_min(travel_time_p50, n = 1, with_ties = FALSE, na_rm = FALSE) %>%
  ungroup() %>%
  rename(
    nearest_access_point_PT = to_id,
    travel_time_min_PT = travel_time_p50
  ) %>%
  mutate(
    access_15_PT = travel_time_min_PT <= 15,
    access_30_PT = travel_time_min_PT <= 30,
    access_60_PT = travel_time_min_PT <= 60
  )
```

```{r}
TTM_walking_access <- TTM_walk_min_pairs %>%
  group_by(from_id) %>%
  slice_min(travel_time_p50, n = 1, with_ties = FALSE, na_rm = FALSE) %>%
  ungroup() %>%
  rename(
    nearest_access_point_walking = to_id,
    travel_time_min_walking = travel_time_p50
  ) %>%
  mutate(
    access_15_walking = travel_time_min_walking <= 15,
    access_30_walking = travel_time_min_walking <= 30,
    access_60_walking = travel_time_min_walking <= 60
  )
```

```{r}
TTM_cycle_access <- TTM_cycle_min_pairs %>%
  group_by(from_id) %>%
  slice_min(travel_time_p50, n = 1, with_ties = FALSE, na_rm = FALSE) %>%
  ungroup() %>%
  rename(
    nearest_access_point_cycling = to_id,
    travel_time_min_cycling = travel_time_p50
  ) %>%
  mutate(
    access_15_cycling = travel_time_min_cycling <= 15,
    access_30_cycling = travel_time_min_cycling <= 30,
    access_60_cycling = travel_time_min_cycling <= 60
  )
```

```{r}
saveRDS(TTM_PT_access, "data/r5r_TTM/TTM_PT_access.rds")
saveRDS(TTM_walking_access, "data/r5r_TTM/TTM_walking_access.rds")
saveRDS(TTM_cycle_access, "data/r5r_TTM/TTM_cycle_access.rds")
```

```{r}
# check 
head(TTM_PT_access)
head(TTM_walking_access)
head(TTM_cycle_access)
```

```{r}
# check
nrow(TTM_PT_access)
nrow(TTM_walking_access)
nrow(TTM_cycle_access)
```

Ok, some LSOAs are missing, 1 missing each for PT and walking amd 3 missing for cycling, maybe because no routing available within our set max_trip_duration ... etc.

Find which LSOAs are missing:

```{r}
# get original lsoas from origins
original_lsoas <- origins %>%
  st_drop_geometry() %>%
  transmute(from_id = id)

# check what missing from PT
missing_PT <- original_lsoas %>%
  anti_join(TTM_PT_access %>% distinct(from_id), by = "from_id")

# check what missing from walking
missing_walk <- original_lsoas %>%
  anti_join(TTM_walking_access %>% distinct(from_id), by = "from_id")

# check what missing from cycling
missing_cycle <- original_lsoas %>%
  anti_join(TTM_cycle_access %>% distinct(from_id), by = "from_id")

missing_PT
missing_walk
missing_cycle
```

Missing LSOA in PT: E01002482

Missing LSOA in Walking: E01002482

Missing LSOAs in Cycling: E01033929, E01032783, E01002482

```{r}
# check if this is missing from beginning of r5r 
original_lsoas %>%
  anti_join(TTM_PT %>% distinct(from_id), by = "from_id")

original_lsoas %>%
  anti_join(TTM_walk %>% distinct(from_id), by = "from_id")

original_lsoas %>%
  anti_join(TTM_cycle %>% distinct(from_id), by = "from_id")
```

ok it is the same, next, make it to retain all 4494 LSOAs and join the three modes together

```{r}
# take all lsoa 4994 first
multimodal_travel_time_lsoa <- original_lsoas %>%
  left_join(
    # add PT TTM
    TTM_PT_access %>%
      select(
        from_id,
        nearest_access_point_PT,
        travel_time_min_PT,
        access_15_PT,
        access_30_PT,
        access_60_PT
      ),
    by = "from_id"
  ) %>%
  # add walking TTM
  left_join(
    TTM_walking_access %>%
      select(
        from_id,
        nearest_access_point_walking,
        travel_time_min_walking,
        access_15_walking,
        access_30_walking,
        access_60_walking
      ),
    by = "from_id"
  ) %>%
  # add cycle TTM
  left_join(
    TTM_cycle_access %>%
      select(
        from_id,
        nearest_access_point_cycling,
        travel_time_min_cycling,
        access_15_cycling,
        access_30_cycling,
        access_60_cycling
      ),
    by = "from_id"
  ) %>%
  # mutate and no TTM LSOAs access times in FALSE
  mutate(
    access_15_PT = coalesce(access_15_PT, FALSE),
    access_30_PT = coalesce(access_30_PT, FALSE),
    access_60_PT = coalesce(access_60_PT, FALSE),

    access_15_walking = coalesce(access_15_walking, FALSE),
    access_30_walking = coalesce(access_30_walking, FALSE),
    access_60_walking = coalesce(access_60_walking, FALSE),

    access_15_cycling = coalesce(access_15_cycling, FALSE),
    access_30_cycling = coalesce(access_30_cycling, FALSE),
    access_60_cycling = coalesce(access_60_cycling, FALSE)
  )
```

```{r}
# check
nrow(multimodal_travel_time_lsoa)
```

```{r}
# check
head(multimodal_travel_time_lsoa)
```

```{r}
# check one missing LSOA
multimodal_travel_time_lsoa %>%
  filter(from_id == "E01002482")
```

```{r}
# save as CSV (just a copy for just in case)
write.csv(multimodal_travel_time_lsoa, "data/multimodal_travel_time_lsoa.csv", row.names = FALSE)
```

### 5.5. Join LSOA and Save file

```{r}
london_lsoa_ttm <- london_lsoa %>%
  left_join(
    multimodal_travel_time_lsoa,
    by = c("LSOA21CD" = "from_id")
  )

nrow(london_lsoa_ttm)
```

```{r}
# save as geopackage cus names too long
st_write(london_lsoa_ttm,"data/multimodal_travel_time_lsoa/multimodal_travel_time_lsoa.gpkg", delete_dsn = TRUE)
```
