---
title: "6_Statistical_Analyses"
---

This markdown contains the code for GLM (Section 5) and BYM2 regression modelling (Section 6), validating models, testing model residuals spatial autocorrelation and BiLISA clustering analysis (Section 7). Data loading, preparation, wrangling and exploratory data analysis are performed in Section 1-3 prior the statistical analyses in the subsequent sections.

The first draft code script is under the code_draft folder within this project's directory.

Spatial tools: <https://doc.esri.com/en/arcgis-pro/latest/tool-reference/spatial-statistics/an-overview-of-the-modeling-spatial-relationships-toolset.html>

```{r}
# remove objects from R environment and clear memory
rm(list = ls())    
gc()
```

## Load Packages

```{r}
library(sf)
library(readr)
library(dplyr)
library(knitr)
library(tidyr)
library(tibble)
library(tidyverse)
library(tmap)
library(cowplot)
library(plotly)
library(corrplot)
library(broom)
library(stringr)
library(spdep)
library(car)
library(janitor)
library(MASS)
library(INLA)
library(performance)
```

## Data Loading

Clean names, select initial needed columns and rename columns for consistency across datasets

```{r}
# greater london boundary 
london_boundary <- st_read("data/gla/London_GLA_Boundary.shp")

# london lsoa boundaries
london_lsoa <- st_read("data/lsoa_london/lsoa_london.shp")
```

```{r}
# travel time 
multimodal_accessibility_lsoa <- st_read("data/multimodal_accessibility_lsoa.gpkg")%>%
  clean_names()%>%
  dplyr::select(
    lsoa21cd,
    lsoa21nm,
    access_score_15_pt,
    access_score_60_pt,
    access_score_15_walk,
    access_score_60_walk,
    access_score_15_cycle,
    access_score_60_cycle,
    travel_time_min_pt,
    travel_time_min_walk,
    travel_time_min_cycle,
    geom
  )%>%
  rename(
    geometry = geom
  )
```

```{r}
# SUHII as lsoa heat exposure
suhii_lsoa <- st_read("data/London_LSOA_LST_SUHI/London_LSOA_LST_SUHI.shp")%>% 
  st_transform(.,27700)%>%
  clean_names()%>%
  dplyr::select(
    lsoa21cd,
    lsoa21nm,
    # mean LST per LSOA
    mean,
    # london mean LST
    london_mea,
    # SUHII equivalent
    lst_differ,
    geometry
  )%>%
  # rename column names which were cut short from GEE export
  rename(
    mean_lst_lsoa = mean,
    london_mean = london_mea,
    suhii = lst_differ
  )
```

```{r}
# IMD score
imd_score <- read_csv("data/File_7_IoD2025.csv")%>%
  clean_names()%>%
  dplyr::select(
    lsoa_code_2021,
    lsoa_name_2021,
    index_of_multiple_deprivation_imd_score
    )%>%
  rename(
    lsoa21cd = lsoa_code_2021,
    lsoa21nm = lsoa_name_2021,
    imd_score = index_of_multiple_deprivation_imd_score
  )
```

```{r}
#  census data

# TS006 population density
census_pop_density <- read_csv("data/census2021/census2021-ts006-lsoa.csv")%>%
  clean_names()%>%
  dplyr::select(
    geography,
    geography_code,
    population_density_persons_per_square_kilometre_measures_value
  )%>%
  rename(
    lsoa21cd = geography_code,
    lsoa21nm = geography,
    pop_density = population_density_persons_per_square_kilometre_measures_value
  )

# TS007a age by 5 year band
census_age_5year_band <- read_csv("data/census2021/census2021-ts007a-lsoa.csv")%>%
  clean_names()%>%
  dplyr::select(
    geography,
    geography_code,
    age_total,
    age_aged_4_years_and_under,
    # for filter for age 65 or over 
    age_aged_65_to_69_years,
    age_aged_70_to_74_years,
    age_aged_75_to_79_years,
    age_aged_80_to_84_years,
    age_aged_85_years_and_over
  )%>%
  rename(
    lsoa21cd = geography_code,
    lsoa21nm = geography,
    age_4_and_under = age_aged_4_years_and_under,
    age_65_to_69 = age_aged_65_to_69_years,
    age_70_to_74 = age_aged_70_to_74_years,
    age_75_to_79 = age_aged_75_to_79_years,
    age_80_to_84 = age_aged_80_to_84_years,
    age_85_and_over = age_aged_85_years_and_over
  )

# TS021 Ethnicity
# definition of ethnic minority groups
# https://www.ethnicity-facts-figures.service.gov.uk/style-guide/writing-about-ethnicity/ 
# https://service-manual.ons.gov.uk/content/language/ethnicity-and-race
census_ethnicity <- read_csv("data/census2021/census2021-ts021-lsoa.csv")%>%
  clean_names()%>%
  dplyr::select(
    geography,
    geography_code,
    ethnic_group_total_all_usual_residents,
    # Use “Ethnic minorities” to refer to all ethnic groups except the White British group
    ethnic_group_white_english_welsh_scottish_northern_irish_or_british
  )%>%
  rename(
    lsoa21cd = geography_code,
    lsoa21nm = geography,
    ethnic_group_total_residents = ethnic_group_total_all_usual_residents,
    white_british_group = ethnic_group_white_english_welsh_scottish_northern_irish_or_british
  )
```

## 1. Data Cleaning

### 1.1. Wrangle data variables for regression modelling

```{r}
# take lsoa codes for filtering for london lsoas only
london_lsoa_codes <- london_lsoa %>%
  clean_names() %>%
  pull(lsoa21cd)
```

```{r}
# create percentage variables for age 4 and under and age 65 and over
census_age_cleaned <- census_age_5year_band %>%
  mutate(
    # create % of age 4 and under
    # calculate percentage
    age_4_and_under_pct = ifelse(
      age_total > 0,
      (age_4_and_under / age_total) * 100,
      NA
    ),
    # create % of age 65 and over
    age_65_and_over = rowSums(
      across(
        # sum up 65 to 85 and over
        c(
          age_65_to_69,
          age_70_to_74,
          age_75_to_79,
          age_80_to_84,
          age_85_and_over
        )
      ),
      na.rm = TRUE
    ),
    # calculate percentage
    age_65_and_over_pct = ifelse(
      age_total > 0,
      (age_65_and_over / age_total)* 100,
      NA
    )
  )%>% 
  # keep the ones that are within london lsoas only
  filter(lsoa21cd %in% london_lsoa_codes)
```

```{r}
# create ethnic minority group variable
census_ethnicity_cleaned <- census_ethnicity %>%
   mutate(
     # everything else but white british group - defined by ONS
    ethnic_minority =
      ethnic_group_total_residents - white_british_group,
    # calculate percentage
    ethnic_minority_pct = ifelse(
      ethnic_group_total_residents > 0,
      (ethnic_minority / ethnic_group_total_residents) * 100,
      NA
    )
  )%>% 
  # keep the ones that are within london lsoas only
  filter(lsoa21cd %in% london_lsoa_codes)
```

```{r}
# imd score variable 
# filter for london lsoas only
imd_score_cleaned <- imd_score %>% 
  filter(lsoa21cd %in% london_lsoa_codes)
```

### 1.2. Join Data

```{r}
# join data
london_join <- multimodal_accessibility_lsoa %>%
  # SUHII 
  left_join(
    suhii_lsoa %>%
      # drop geom cus multimodal travel time already have it
      st_drop_geometry() %>%
      # select needed columns
      dplyr::select(lsoa21cd, suhii),
    # join by lsoa code
    by = "lsoa21cd"
  ) %>%
  
  # IMD 
  left_join(
    imd_score_cleaned %>%
      dplyr::select(lsoa21cd, imd_score),
    by = "lsoa21cd"
  ) %>%
  
  # population density
  left_join(
    census_pop_density %>%
      dplyr::select(lsoa21cd, pop_density),
    by = "lsoa21cd"
  ) %>%
  
  # age
  left_join(
    census_age_cleaned %>%
      dplyr::select(lsoa21cd, age_4_and_under_pct, age_65_and_over_pct),
    by = "lsoa21cd"
  ) %>%
  
  # ethnicity
  left_join(
    census_ethnicity_cleaned %>%
      dplyr::select(lsoa21cd, ethnic_minority_pct),
    by = "lsoa21cd"
  )
```

**Add an additional Inner and Outer London control + join to data**

```{r}
# create inner london control variable
# following: https://www.london.gov.uk/programmes-strategies/planning/london-plan/the-london-plan-2021-online/annex-2-inner-and-outer-london-boroughs
inner_london_boroughs <- c(
  "City of London", "Camden", "Greenwich", "Hackney",
  "Hammersmith and Fulham", "Islington", "Kensington and Chelsea",
  "Lambeth", "Lewisham", "Newham", "Southwark", "Tower Hamlets",
  "Wandsworth", "Westminster"
)

london_join <- london_join %>%
  mutate(
    #remove the letters numbers behind borough names to keep it a clean borough name only
    borough = str_remove(lsoa21nm, "\\s\\d+[A-Z]$"),
    # inner and outer london categories
    inner_outer_london = if_else(borough %in% inner_london_boroughs, "Inner", "Outer"),
    inner_outer_london = factor(inner_outer_london, levels = c("Outer", "Inner"))
  )
```

```{r}
#check fully joined data
names(london_join)
```

## 2. Check Data for Regression

```{r}
# check data types
check_data <- london_join %>%
  st_drop_geometry() %>%
  summarise(across(everything(), ~class(.)[1])) %>%
  pivot_longer(everything(), names_to = "variable", values_to = "type")

#view(check_data)
```

```{r}
# descriptive stats
london_join %>%
  st_drop_geometry() %>%
  dplyr::select(where(is.numeric)) %>%
  summarise(across(everything(),
                   list(mean = ~mean(., na.rm = TRUE),
                        sd = ~sd(., na.rm = TRUE),
                        min = ~min(., na.rm = TRUE),
                        max = ~max(., na.rm = TRUE)))) %>%
  pivot_longer(everything(),
               names_to = c("variable", "stat"),
               names_pattern = "(.*)_(mean|sd|min|max)") %>%
  pivot_wider(names_from = stat, values_from = value) %>%
  kable(digits = 2)
```

## 3. Exploratory Data Analysis

**Plot Variables Distribution**

```{r}
# select only numeric columns
numeric_cols <- london_join %>%
  sf::st_drop_geometry() %>%
  dplyr::select(where(is.numeric))

# convert to long format for faceting
numeric_long <- numeric_cols %>%
  pivot_longer(cols = everything(), names_to = "variable", values_to = "value")

# plot faceted histograms
variables_dist_hist <- ggplot(numeric_long, aes(x = value)) +
  geom_histogram(bins = 30, fill = "steelblue", color = "white") +
  facet_wrap(~ variable, scales = "free", ncol = 3) +
  theme_minimal() +
  labs(title = "Histograms of Numerical Variables")

variables_dist_hist
```

**Plot Heatmap Correlation (Spearman)**

```{r}
# select only the numeric variables 
corr_matrix_spearman <- london_join %>%
  st_drop_geometry() %>%
  dplyr::select(
    access_score_15_pt,
    access_score_60_pt,
    access_score_15_walk,
    access_score_60_walk,
    access_score_15_cycle,
    access_score_60_cycle,
    suhii,
    imd_score,
    age_4_and_under_pct,
    age_65_and_over_pct,
    ethnic_minority_pct,
    pop_density
  ) %>%
  cor(
    use = "pairwise.complete.obs",
    method = "spearman"
  )

heatmap <- corrplot(
  corr_matrix_spearman,
  method = "color",
  type = "full",
  col = rev(COL2("RdBu", 200)),
  tl.col = "black",
  tl.srt = 45,
  addCoef.col = "black",
  number.cex = 0.6
)
```

```{r}
# save corrplot as png

dir.create("figures", showWarnings = FALSE)

png(
  filename = "figures/Fig8_Spearman_Heatmap.png",
  width = 8.5,
  height = 5.5,
  units = "in",
  res = 300,
  bg = "white"
)

corrplot(
  corr_matrix_spearman,
  method = "color",
  type = "full",
  col = rev(COL2("RdBu", 200)),
  tl.col = "black",
  tl.srt = 45,
  addCoef.col = "black",
  number.cex = 0.6
)

dev.off()
```

```{r}
# the 6 dependent variables
dependents <- c(
  "access_score_15_pt",
  "access_score_60_pt",
  "access_score_15_walk",
  "access_score_60_walk",
  "access_score_15_cycle",
  "access_score_60_cycle"
)

# check distribution of the dependent variables
distribution_check <- do.call(
  rbind,
  lapply(dependents, function(v) {

    y <- london_join[[v]]

    data.frame(
      dependents = v,
      n = sum(!is.na(y)),
      mean = mean(y, na.rm = TRUE),
      variance = var(y, na.rm = TRUE),
      variance_mean_ratio = var(y, na.rm = TRUE) /
        mean(y, na.rm = TRUE),
      zeros = sum(y == 0, na.rm = TRUE),
      zero_percent = mean(y == 0, na.rm = TRUE) * 100,
      maximum = max(y, na.rm = TRUE)
    )
  })
)

distribution_check
```

### 3.1. Transform Variables

**Log transform population density and standardise other variables**

As above examination of data, highly skewed population density needs to be log transformed and all the variables that are in percentages will be standardised.

```{r}
model_variables <- london_join %>%
  mutate(
    # log transforms pop density variable
    log_pop_density = log(pop_density), 

    # standardise variables by z-scores
    imd_z = as.numeric(scale(imd_score)),
    suhii_z = as.numeric(scale(suhii)),
    log_pop_density_z = as.numeric(scale(log_pop_density)),
    age_4_under_z = as.numeric(scale(age_4_and_under_pct)),
    age_65_over_z = as.numeric(scale(age_65_and_over_pct)),
    ethnic_minority_z  = as.numeric(scale(ethnic_minority_pct))
  )
```

## 4. Model Choices

To model geographic variation in count data, use poisson or negative binomial modelling likelihoods.

## 5. GLM

### 5.1. Negative Binomial

```{r}
# independent variables
independents <- c(
  "imd_z * suhii_z",
  "log_pop_density_z",
  "age_4_under_z",
  "age_65_over_z",
  "ethnic_minority_z",
  "inner_outer_london"
)

# dependent variables 
dependents <- c(
  #"access_score_15_pt",
  "access_score_60_pt",
  #"access_score_15_walk",
  "access_score_60_walk",
  "access_score_15_cycle",
  "access_score_60_cycle"
)

# fit negative binomial GLM for each dependent 
fit_nb <- function(dependents, data, independents) {
  f <- reformulate(independents, response = dependents)
  glm.nb(f, data = data)
}

# loop over dependent variables 
models_nb <- lapply(dependents, fit_nb, data = model_variables, independents = independents)
names(models_nb) <- dependents
```

```{r}
# print summary stats of all models
for (nm in names(models_nb)) {
  cat("\n=====================================================\n")
  cat("Negative Binomial Model:", nm, "\n")
  cat("=====================================================\n")
  print(summary(models_nb[[nm]]))
}
```

### 5.2. Poisson

```{r}
# independent variables
independents <- c(
  "imd_z * suhii_z",
  "log_pop_density_z",
  "age_4_under_z",
  "age_65_over_z",
  "ethnic_minority_z",
  "inner_outer_london"
)

# dependent variables
dependents <- c(
  "access_score_15_pt",
  #"access_score_60_pt",
  "access_score_15_walk"
  #"access_score_60_walk",
  #"access_score_15_cycle",
  #"access_score_60_cycle"
)

# fit poisson GLM for each dependent 
fit_poisson <- function(dependents, data, independents) {
  f <- reformulate(independents, response = dependents)
  glm(f, data = model_variables, family = poisson(link = "log"))
}

# loop over dependent variables 
models_poisson <- lapply(dependents, fit_poisson, data = model_variables, independents = independents)
names(models_poisson) <- dependents
```

```{r}
# print summary stats of all models
for (nm in names(models_poisson)) {
  cat("\n=====================================================\n")
  cat("Poisson Model:", nm, "\n")
  cat("=====================================================\n")
  print(summary(models_poisson[[nm]]))
}
```

```{r}
# check Poisson dispersion
# https://www.rdocumentation.org/packages/insurancerating/versions/0.7.5/topics/check_overdispersion 
dispersion_results <- lapply(
  models_poisson,
  performance::check_overdispersion
)

dispersion_results
```

Poisson did not show overdispersion.

AIC values as below table (bold text refers to lower AIC values. As well as dispersion checked as above, dispersion values close to 1.

| Dependent variable    | AIC - Negative Binomial | AIC - Poisson |
|-----------------------|-------------------------|---------------|
| access_score_15_pt    | 13934                   | **13932**     |
| access_score_60_pt    | **49596**               | 90101         |
| access_score_15_walk  | 16249                   | **16247**     |
| access_score_60_walk  | **32617**               | 33312         |
| access_score_15_cycle | **21932**               | 22076         |
| access_score_60_cycle | **42929**               | 55092         |

### 5.3. Residuals with Moran's I

**Build spatial weights with LSOA boundary**

```{r}
#calculate the centroids of all LSOA in London
coordsLSOA <- london_lsoa%>%
  st_centroid()%>%
  st_geometry()

#plot(coordsLSOA)
```

```{r}
# create spatial weights  

# queen contiguity
LLSOA_nb <- london_lsoa %>%
  poly2nb(., queen=T)

# k nearest neighbours
knn_LSOA <-coordsLSOA %>%
  knearneigh(., k=4)

LLSOA_knn <- knn_LSOA %>%
  knn2nb()
```

```{r, eval=FALSE}
plot(LLSOA_nb, st_geometry(coordsLSOA), col="red")
```

```{r, eval=FALSE}
plot(LLSOA_knn, st_geometry(coordsLSOA), col="blue")
```

```{r}
#create a spatial weights matrix object from these weights

# queen contiguity spatial weight matrix
LLSOA.queens_weight <- LLSOA_nb %>%
  nb2listw(., style="W")

# k=4 nearest neighbours spatial weight matrix
LLSOA.knn_4_weight <- LLSOA_knn %>%
  nb2listw(., style="W")
```

#### 5.3.1. Models VIF and Residuals

```{r}
# list the models
glm_models <- list(
  "access_score_15_pt" = models_poisson$access_score_15_pt,
  "access_score_60_pt" = models_nb$access_score_60_pt,
  "access_score_15_walk" = models_poisson$access_score_15_walk,
  "access_score_60_walk" = models_nb$access_score_60_walk,
  "access_score_15_cycle" = models_nb$access_score_15_cycle,
  "access_score_60_cycle" = models_nb$access_score_60_cycle
)
```

**VIF for each model**

```{r}
# loop and calculate VIF for each model
glm_vif <- lapply(
  glm_models,
  car::vif
)

glm_vif
```

**Loop for each GLM model for Pearson residuals, Queens contiguity Moran's I and KNN Moran's I**

```{r}
# create function to calculate Pearson residuals
# and test spatial autocorrelation using Moran's I
check_glm_residuals <- function(model, model_name) {
  
  # residuals with Pearson 
  model_resids <- residuals(
    model,
    type = "pearson"
  )
  
  # Moran's I using Queen contiguity spatial weights as defined above
  moran_queen <- spdep::moran.test(
    model_resids,
    LLSOA.queens_weight
  ) %>%
    broom::tidy() %>%
    mutate(
      model = model_name,
      weights = "Queen"
    )
  
  # Moran's I using 4-nearest-neighbour spatial weights as defined above
  moran_knn <- spdep::moran.test(
    model_resids,
    LLSOA.knn_4_weight
  ) %>%
    broom::tidy() %>%
    mutate(
      model = model_name,
      weights = "KNN = 4"
    )
  
  # combine both results
  bind_rows(
    moran_queen,
    moran_knn
  )
}


# apply residual Moran's I tests to all six GLM models
glm_moran_results <- bind_rows(
  lapply(
    names(glm_models),
    function(model_name) {
      check_glm_residuals(
        glm_models[[model_name]],
        model_name
      )
    }
  )
)
```

#### 5.3.2. Summary of Models' residuals

```{r}
# GLM Moran's I summary from above results
# pivot it for easier read

glm_morans_table <- glm_moran_results %>%
  dplyr::select(
    model,
    weights,
    estimate1,
    p.value
  ) %>%
  
  # rename
  dplyr::mutate(
    model = dplyr::recode(
      model,
      "access_score_15_pt" = "15-minute PT",
      "access_score_60_pt" = "60-minute PT",
      "access_score_15_walk" = "15-minute Walking",
      "access_score_60_walk" = "60-minute Walking",
      "access_score_15_cycle" = "15-minute Cycling",
      "access_score_60_cycle" = "60-minute Cycling"
    )
  ) %>%
  
  # Queen and KNN results into separate columns
  tidyr::pivot_wider(
    names_from = weights,
    values_from = c(estimate1, p.value)
  ) %>%
  
  # rename columns for reporting
  dplyr::rename(
    Model = model,
    `Queens Moran's I` = estimate1_Queen,
    `Queens p-value` = p.value_Queen,
    `KNN = 4 Moran's I` = `estimate1_KNN = 4`,
    `KNN = 4 p-value` = `p.value_KNN = 4`
  )


# format Moran's I and p-values
glm_morans_table_report <- glm_morans_table %>%
   dplyr::mutate(
    `Queens Moran's I` =
      sprintf("%.3f", `Queens Moran's I`),
  
    `Queens p-value` =
      format.pval(
        `Queens p-value`,
        digits = 3,
        eps = 0.001
      ),
  
    `KNN = 4 Moran's I` =
      sprintf("%.3f", `KNN = 4 Moran's I`),
  
    `KNN = 4 p-value` =
      format.pval(
        `KNN = 4 p-value`,
        digits = 3,
        eps = 0.001
       )
    )
  
  
# show table
knitr::kable(
glm_morans_table_report,
align = c("l", "c", "c", "c", "c"),
caption = "Moran's I tests of GLM residual spatial autocorrelation"
)
```

Strong positive spatial autocorrelation for queens case and nearest neighbours by Moran's I for all models, so next use spatial models

## 6. Global Spatial Regression model (BYM2)

Given that the data type is count data, traditional spatial error models may not be suitable here ( <https://www.econstor.eu/bitstream/10419/168040/1/89655662X.pdf>).

Bayesian spatial multilevel mixed-effects model incoporating BMY2 with `INLA` package

```{r}
# Source: INLA package: https://www.r-inla.org/download/index.html

# testing version tends to work better than the stable version
#install.packages("INLA", repos=c(getOption("repos"), INLA="https://inla.r-inla-download.org/R/stable"), dep=TRUE)
#install.packages("INLA", repos=c(getOption("repos"), INLA="https://inla.r-inla-download.org/R/testing"), dep=TRUE)
```

```{r}
# load INLA package
library(INLA)

# check INLA package version
packageVersion("INLA")
```

```{r}
# uncomment and add this if crash:
# inla.setOption(inla.mode="experimental") 

# check INLA version details
INLA::inla.version()
```

```{r}
# make a spatial model variables object
spatial_model_variables <- model_variables %>%
  dplyr::select(
    lsoa21cd,
    # dependent
    access_score_15_pt,
    access_score_60_pt,
    access_score_15_walk,
    access_score_60_walk,
    access_score_15_cycle,
    access_score_60_cycle,
    # independent
    imd_z,
    suhii_z,
    log_pop_density_z,
    age_4_under_z,
    age_65_over_z,
    ethnic_minority_z,
    inner_outer_london,
    geometry
  ) %>%
  arrange(lsoa21cd)
```

```{r}
# check if it is sf objects
class(spatial_model_variables)
```

### 6.1. Spatial Neighbourhood structure

```{r}
# code adapted from:  
# https://www.r-inla.org/examples/AlvinBong/alvinbong-example.html, a tutorial that uses count-based areal data
# and also:
# https://www.sciencedirect.com/science/article/pii/S1353829225001765?via%3Dihub#sec2 --> https://github.com/mbcalles/Area-Level-Deprivation-Traffic-Injury

# 1. Create a neighborhood list based on polygon contiguity - queens
nb <- spdep::poly2nb(
  spatial_model_variables,
  queen = TRUE
)

# 2. Convert the neighbor list to INLA adjacency format
spdep::nb2INLA("data/map.graph", nb)

# 3. Set the location of the adjacency file for INLA
map.adj <- file.path(getwd(), "data", "map.graph")

# 4. Load adjacency matrix as a graph for INLA spatial models
g <- inla.read.graph(filename = map.adj)
```

### 6.2. Inla call/ fit - Generalised Linear Models (Poisson)

```{r}
# create index for spatial effects
spatial_model_variables$area_id <- 1:nrow(spatial_model_variables)
```

```{r}
# dependent variables for BYM2 
bym2_dependents <- c(
  "15_pt"    = "access_score_15_pt",
  "60_pt"    = "access_score_60_pt",
  "15_walk"  = "access_score_15_walk",
  "60_walk"  = "access_score_60_walk",
  "15_cycle" = "access_score_15_cycle",
  "60_cycle" = "access_score_60_cycle"
)


# create function for BYM2 modelling looping for each dependent
create_bym2_formula <- function(dependent) {
  
  as.formula(
    paste0(
      dependent,
      " ~ ",
      "imd_z * suhii_z + ",
      "log_pop_density_z + ",
      "age_4_under_z + ",
      "age_65_over_z + ",
      "ethnic_minority_z + ",
      "factor(inner_outer_london) + ",
      "f(area_id, ",
      "model = 'bym2', ",
      "graph = g, ",
      "constr = TRUE, ",
      "scale.model = TRUE)"
    )
  )
}
```

Having tested through all depenedent variables with negative bionomial and Poisson modelling likelihood, the final likelihood of Poisson is selected, the below will model BYM2 by Poisson likelihood:

```{r}
# loop BYM2 model one by one
for (model_name in names(bym2_dependents)) {
  
  # select dependent variable
  dependent <- bym2_dependents[[model_name]]
  
  # create model with formula as defined above
  model_formula <- create_bym2_formula(dependent)
  
  # fit BYM2 Poisson model
  model_bym2 <- INLA::inla(
    formula = model_formula,
    family = "poisson",
    data = spatial_model_variables,
    control.predictor = list(
      compute = TRUE
    ),
    # also compute DIC, WAIC, CPO
    control.compute = list(
      dic = TRUE,
      waic = TRUE,
      cpo = TRUE,
      config = TRUE,
      return.marginals.predictor = TRUE
    ),
    verbose = FALSE
  )
  
  # save as RDS
  saveRDS(
    model_bym2,
    paste0(
      "data/BYM2/model_",
      model_name,
      "_bym2.rds"
    )
  )
  
  # remove model before fitting the next model
  rm(model_bym2)
  gc()
}
```

**Reload the RDS objects to continue the analysis**:

```{r}
# reload
model_15_pt_bym2 <- readRDS("data/BYM2/model_15_pt_bym2.rds")
model_60_pt_bym2 <- readRDS("data/BYM2/model_60_pt_bym2.rds")
model_15_walk_bym2 <- readRDS("data/BYM2/model_15_walk_bym2.rds")
model_60_walk_bym2 <- readRDS("data/BYM2/model_60_walk_bym2.rds")
model_15_cycle_bym2 <- readRDS("data/BYM2/model_15_cycle_bym2.rds")
model_60_cycle_bym2 <- readRDS("data/BYM2/model_60_cycle_bym2.rds")
```

```{r}
# name models for cleaner interpretation
bym2_models <- list(
  "Model 1" = model_15_pt_bym2,
  "Model 2" = model_60_pt_bym2,
  "Model 3" = model_15_walk_bym2,
  "Model 4" = model_60_walk_bym2,
  "Model 5" = model_15_cycle_bym2,
  "Model 6" = model_60_cycle_bym2
)

# name variables 
variable_labels <- c(
  "(Intercept)" = "Intercept",
  "imd_z" = "IMD (z)",
  "suhii_z" = "SUHII (z)",
  "imd_z:suhii_z" = "IMD × SUHII",
  "log_pop_density_z" = "Log pop. density (z)",
  "age_4_under_z" = "Age 4 or under % (z)",                 
  "age_65_over_z" = "Age 65 or over % (z)",
  "ethnic_minority_z" = "Ethnic minority % (z)",
  "factor(inner_outer_london)Inner" = "Inner, Outer London"
)


# Extract fixed effects, exponentiate to rate ratios, format as one cell
# just like this: https://www.sciencedirect.com/science/article/pii/S2405673123000144
fixed_effects_long <- lapply(names(bym2_models), function(model_name) {
  bym2_models[[model_name]]$summary.fixed %>%
    as.data.frame() %>%
    rownames_to_column("term") %>%
    transmute(
      model = model_name,
      variable = variable_labels[term],
      cell = sprintf(
        "%.3f (%.3f, %.3f)",
        # mean IRR and the 95% CI
        exp(mean), exp(`0.025quant`), exp(`0.975quant`)
      )
    )
}) %>% bind_rows()

# One column per model, one row per variable
fixed_effects_table <- fixed_effects_long %>%
  pivot_wider(names_from = model, values_from = cell) %>%
  arrange(match(variable, variable_labels))

knitr::kable(
  fixed_effects_table,
  caption = "BYM2 Fixed Effects Mean Incidence Rate Ratios and 95% Credible Intervals (CI)"
)
```

### 6.3. Validation Checks

#### 6.3.1. Diagnostic

<https://becarioprecario.bitbucket.io/inla-gitbook/ch-INLA.html#sec:modelassess>

```{r}
# List the 6 models
six_models <- list(
  "15-minute PT" = model_15_pt_bym2,
  "60-minute PT" = model_60_pt_bym2,
  "15-minute Walking" = model_15_walk_bym2,
  "60-minute Walking" = model_60_walk_bym2,
  "15-minute Cycling" = model_15_cycle_bym2,
  "60-minute Cycling" = model_60_cycle_bym2
)
```

**Adjusted PIT**

```{r}
# Adjusted PIT histograms
png(
  filename = "figures/FigC1_Adjusted_PIT_Histograms.png",
  width = 2400,
  height = 1600,
  res = 300
)

par(mfrow = c(2, 3), mar = c(3, 3, 2, 1))

for (nm in names(six_models)) {

  m <- six_models[[nm]]

  # use valid CPO/PIT observations only
  valid <- is.finite(m$cpo$pit) &
           is.finite(m$cpo$cpo) &
           m$cpo$cpo > 0 &
           m$cpo$failure == 0

  # calculate adjusted PIT for discrete count data
  adjusted_pit <- m$cpo$pit[valid] -
                  0.5 * m$cpo$cpo[valid]

  # histogram information without plotting
  hist(
     adjusted_pit,
     breaks = seq(0, 1, by = 0.05),
     probability = TRUE,
     main = nm,
     xlab = "Adjusted PIT",
     ylab = "Density",
     xlim = c(0, 1),
     # keep y axis consistent for all plots
     ylim = c(0, 3),
     yaxt = "n",  
     col = "steelblue"
   )
 
 
   axis(
     side = 2,
     at = seq(0, 3, by = 0.5),
     labels = sprintf("%.1f", seq(0, 3, by = 0.5))
   )
 
   # uniform PIT distribution line
   abline(h = 1, lty = 2)
}

dev.off()

par(mfrow = c(1, 1))
```

**DIC, WAIC, CPO, MLIK**

```{r}
# summary of DIC, WAIC, CPO and MLIK values
model_summary <- lapply(names(six_models), function(model_name) {

  m <- six_models[[model_name]]
  # use valid CPOs
  valid_cpo <- is.finite(m$cpo$cpo) & m$cpo$cpo > 0
  data.frame(
    Model = model_name,
    DIC = m$dic$dic,
    WAIC = m$waic$waic,
    Negative_sum_log_CPO = -sum(log(m$cpo$cpo[valid_cpo])),
    Log_Marginal_Likelihood =
      m$mlik[1, 1]
  )

}) %>%
  dplyr::bind_rows()

model_summary
```

**CPO Failures**

Only a very tiny change after recomputing, skip to next.

```{r, eval=FALSE}
cpo_failure_counts <- sapply(
  six_models,
  function(model) sum(model$cpo$failure > 0, na.rm = TRUE)
)

cpo_failure_counts
```

```{r, eval=FALSE}
# recompute
model_15_pt_bym2_cpo <- INLA::inla.cpo(model_15_pt_bym2)
model_15_walk_bym2_cpo <- INLA::inla.cpo(model_15_walk_bym2)
model_15_cycle_bym2_cpo <- INLA::inla.cpo(model_15_cycle_bym2)

sum(log(model_15_pt_bym2_cpo$cpo$cpo), na.rm = TRUE)
sum(log(model_15_walk_bym2_cpo$cpo$cpo), na.rm = TRUE)
sum(log(model_15_cycle_bym2_cpo$cpo$cpo), na.rm = TRUE)
```

**Queens Contiguity Adjacency Graph checks**

```{r}
# check if grraph nodes matches LSOAs

cat("Number of graph nodes:", g$n, "\n")
cat("Number of LSOAs:", nrow(spatial_model_variables), "\n")

cat(
  "Graph and data sizes match:",
  g$n == nrow(spatial_model_variables),
  "\n"
)

cat(
  "area_id is correctly ordered:",
  identical(
    spatial_model_variables$area_id,
    seq_len(nrow(spatial_model_variables))
  ),
  "\n"
)

cat(
  "Number of duplicated LSOA codes:",
  anyDuplicated(spatial_model_variables$lsoa21cd),
  "\n"
)
```

#### 6.3.2. Residuals

Rebuild the centroids, queens and knn

```{r}
# centroids 
coordsLSOA2 <- spatial_model_variables %>% 
  st_centroid() %>% 
  st_geometry()

#plot(coordsLSOA2)
```

```{r}
# queens
LLSOA_nb2 <- spatial_model_variables %>%
  poly2nb(., queen = TRUE)

# nearest neighbours
knn_LSOA2 <-coordsLSOA2 %>%
  knearneigh(., k=4)

LLSOA_knn2 <- knn_LSOA2 %>%
  knn2nb()
```

```{r, eval=FALSE}
#plot 
plot(LLSOA_nb2, st_geometry(coordsLSOA2), col="red")
```

```{r, eval=FALSE}
#plot 
plot(LLSOA_knn, st_geometry(coordsLSOA2), col="blue")
```

```{r}
#create a spatial weights matrix object from these weights

LLSOA.queens_weight2 <- LLSOA_nb2 %>%
  nb2listw(., style="W")

LLSOA.knn_4_weight2 <- LLSOA_knn2 %>%
  nb2listw(., style="W")
```

**Pearson Residuals - i = (Y_i − λ̂\_i) / sqrt(λ̂\_i)**

```{r}
# link each BYM2 model to its corresponding dependent variable
bym2_residual_models <- list(
  "15_pt" = list(
    model = model_15_pt_bym2,
    outcome = "access_score_15_pt"
  ),
  "60_pt" = list(
    model = model_60_pt_bym2,
    outcome = "access_score_60_pt"
  ),
  "15_walk" = list(
    model = model_15_walk_bym2,
    outcome = "access_score_15_walk"
  ),
  "60_walk" = list(
    model = model_60_walk_bym2,
    outcome = "access_score_60_walk"
  ),
  "15_cycle" = list(
    model = model_15_cycle_bym2,
    outcome = "access_score_15_cycle"
  ),
  "60_cycle" = list(
    model = model_60_cycle_bym2,
    outcome = "access_score_60_cycle"
  )
)
```

```{r}
# function for BYM2 Pearson residuals
# and test residual spatial autocorrelation by Moran's I
check_bym2_residuals <- function(model_info, model_name) {
  
  model <- model_info$model
  outcome <- model_info$outcome
  
  # fitted mean count for each LSOA
  fitted_values <- model$summary.fitted.values$mean
  
  # Pearson residuals
  # calculate as the model are INLA models
  # pearson residuals --> observed minus fitted and divided by square root of fitted
  model_resids <- (spatial_model_variables[[outcome]] - fitted_values) / sqrt(fitted_values)
  
  # Moran's I using Queen contiguity 
  moran_queen <- spdep::moran.test(
    model_resids,
    LLSOA.queens_weight2
  ) %>%
    broom::tidy() %>%
    mutate(
      model = model_name,
      weights = "Queen"
    )
  
  # Moran's I using KNN=4
  moran_knn <- spdep::moran.test(
    model_resids,
    LLSOA.knn_4_weight2
  ) %>%
    broom::tidy() %>%
    mutate(
      model = model_name,
      weights = "KNN = 4"
    )
  
  # combine
  bind_rows(
    moran_queen,
    moran_knn
  )
}
```

```{r}
# loop for all six BYM2 models
bym2_moran_results <- bind_rows(
  lapply(
    names(bym2_residual_models),
    function(model_name) {
      
      check_bym2_residuals(
        bym2_residual_models[[model_name]],
        model_name
      )
    }
  )
)
```

These results showed significantly reduced compared to the GLM models, but still some spatial autocorrelation remains –\> write in discussion

```{r}
# BYM2 Moran's I summary from above results
# pivot it for easier read
bym2_morans_table <- bym2_moran_results %>%
  dplyr::select(
    model,
    weights,
    estimate1,
    p.value
  ) %>%
  
  # rename values
  dplyr::mutate(
    model = dplyr::recode(
      model,
      "15_pt" = "15-minute PT",
      "60_pt" = "60-minute PT",
      "15_walk" = "15-minute Walking",
      "60_walk" = "60-minute Walking",
      "15_cycle" = "15-minute Cycling",
      "60_cycle" = "60-minute Cycling"
    )
  ) %>%
  
  # Queen and KNN results into separate columns
  tidyr::pivot_wider(
    names_from = weights,
    values_from = c(estimate1, p.value)
  ) %>%
  
  # rename columns
  dplyr::rename(
    Model = model,
    `Queens Moran's I` = estimate1_Queen,
    `Queens p-value` = p.value_Queen,
    `KNN = 4 Moran's I` = `estimate1_KNN = 4`,
    `KNN = 4 p-value` = `p.value_KNN = 4`
  )


# format Moran's I and p-values
bym2_morans_table_report <- bym2_morans_table %>%
  dplyr::mutate(
    `Queens Moran's I` =
      sprintf("%.3f", `Queens Moran's I`),

    `Queens p-value` =
      format.pval(
        `Queens p-value`,
        digits = 3,
        eps = 0.001
      ),

    `KNN = 4 Moran's I` =
      sprintf("%.3f", `KNN = 4 Moran's I`),

    `KNN = 4 p-value` =
      format.pval(
        `KNN = 4 p-value`,
        digits = 3,
        eps = 0.001
      )
  )


# show table
knitr::kable(
  bym2_morans_table_report,
  align = c("l", "c", "c", "c", "c"),
  caption = "Moran's I tests of BYM2 residual spatial autocorrelation"
)
```

## 7. BiLISA

Well known method to address issues of spatial nonstationarity, allowing geographic variations in the relationships between dependent and explanatory variables

use `spdep` package

```{r}
bilisa_variables <- spatial_model_variables %>%
  arrange(lsoa21cd)

# Create Queen contiguity neighbours
bilisa_nb <- bilisa_variables %>%
  poly2nb(.,queen = T)

bilisa_weights <-  bilisa_nb %>%
  nb2listw(.,style = "W",)
```

### 7.1. BiLISA mapping

code adpated from: <https://r-spatial.github.io/spdep/reference/localmoran_bv.html>

```{r}
# function to run BiLISA and create maps
run_bilisa_map <- function(
    data,
    access_var,
    context_var,
    context_label,
    map_title,
    listw = bilisa_weights,
    boundary = london_boundary,
    nsim = 199,
    seed = 42,
    show_legend = TRUE
) {

  # standardise accessibility variables
  access_z <- as.numeric(
    scale(data[[access_var]])
  )

  # use the already made SUHII or IMD z-score
  context_z <- data[[context_var]]

  # check that variables and weights have matching lengths
  stopifnot(
    length(access_z) == length(context_z),
    length(access_z) == length(listw$neighbours)
  )

  # calculate BiLISA
  lisa_result <- spdep::localmoran_bv(
    x = access_z,
    y = context_z,
    listw = listw,
    nsim = nsim,
    scale = FALSE,
    alternative = "two.sided",
    iseed = seed,
    zero.policy = TRUE
  )

  # significant clusters at p < 0.05
  lisa_cluster <- spdep::hotspot(
    lisa_result,
    Prname = "Pr(folded) Sim",
    cutoff = 0.05,
    quadrant.type = "pysal",
    p.adjust = "none"
  )

  # set clearer name labels for different clusters 
  cluster_labels <- c(
    "Low-High" =
      paste0("Low access–High ", context_label),

    "High-High" =
      paste0("High access–High ", context_label),

    "Low-Low" =
      paste0("Low access–Low ", context_label),

    "High-Low" =
      paste0("High access–Low ", context_label)
  )

  # set order of cluster to be presented
  cluster_levels <- unname(
    cluster_labels[
      c(
        "Low-High",
        "High-High",
        "Low-Low",
        "High-Low"
      )
    ]
  )

  # set cluster colours accordingly
  cluster_colours <- stats::setNames(
    c(
      "#B2182B",  # Low-High
      "#FDDBC7",  # High-High
      "#D1E5F0",  # Low-Low
      "#2166AC"   # High-Low
    ),
    cluster_levels
  )

  # add BiLISA results 
  result_data <- data %>%
    dplyr::mutate(
      # BiLISA stats
      bilisa_i = as.numeric(lisa_result[, 1]),
      # p value 
      bilisa_p = as.numeric(
        lisa_result[, "Pr(folded) Sim"]
      ),
      # cluster type
      bilisa_cluster = as.character(
        lisa_cluster
      ),
      # replace with names defined above
      bilisa_cluster_plot = unname(
        cluster_labels[bilisa_cluster]
      ),
      # use the order defined above
      bilisa_cluster_plot = factor(
        bilisa_cluster_plot,
        levels = cluster_levels
      )
    )

  # keep significant clusters only
  significant_data <- result_data %>%
    dplyr::filter(
      !is.na(bilisa_cluster_plot)
    )

  # legend settings
  legend_setting <- if (show_legend) {
    tm_legend(
      title = "BiLISA cluster",
      position = tm_pos_out("right", "center"),
      frame = FALSE
    )
  } else {
    tm_legend_hide()
  }

  # map
  map_object <-
    tm_shape(result_data) +
    tm_fill(
      fill = "white"
    ) +
    # fill in clusters colour by cluster type
    tm_shape(significant_data) +
    tm_fill(
      fill = "bilisa_cluster_plot",
      fill.scale = tm_scale_categorical(
        values = cluster_colours
      ),
      fill.legend = legend_setting
    ) +
    # add Greater London boundary
    tm_shape(boundary) +
    tm_borders(
      col = "grey30",
      lwd = 1
    ) +
    tm_title(
      map_title
    ) +

    tm_layout(
      frame = FALSE,
      bg.color = "white",
      outer.margins = 0.02
    )

  # return results / outputs
  list(
    result = lisa_result,
    clusters = lisa_cluster,
    data = result_data,
    map = map_object
  )
}
```

```{r}
# accessibility variables for BiLISA analyses
bilisa_access_variables <- c(
  "15_pt" = "access_score_15_pt",
  "60_pt" = "access_score_60_pt",
  "15_walk" = "access_score_15_walk",
  "60_walk" = "access_score_60_walk",
  "15_cycle" = "access_score_15_cycle",
  "60_cycle" = "access_score_60_cycle"
)

# run each accessibility variable once with SUHII and once with IMD
bilisa_context_variables <- c(
  "suhii" = "suhii_z",
  "imd" = "imd_z"
)

# names for mapping
bilisa_context_labels <- c(
  "suhii" = "SUHII",
  "imd" = "IMD"
)


# loop through all accessibility and contextual-variable combinations
for (access_name in names(bilisa_access_variables)) {

  for (context_name in names(bilisa_context_variables)) {

    # run BiLISA analysis and create map
    bilisa_result <- run_bilisa_map(
      data = bilisa_variables,
      access_var = bilisa_access_variables[[access_name]],
      context_var = bilisa_context_variables[[context_name]],
      context_label = bilisa_context_labels[[context_name]],
      map_title = NULL,
      show_legend = FALSE
    )

    # save result using the original object name
    assign(
      paste0(
        "bilisa_",
        access_name,
        "_",
        context_name
      ),
      bilisa_result
    )

    # save map using the original object name
    assign(
      paste0(
        "bilisa_",
        access_name,
        "_",
        context_name,
        "_map"
      ),
      bilisa_result$map
    )
  }
}
```

**Cowplot Combine Maps**

```{r}
# convert into objects for cowplot to combine next
bilisa_grobs <- list(
  tmap::tmap_grob(bilisa_15_pt_suhii_map),
  tmap::tmap_grob(bilisa_15_pt_imd_map),

  tmap::tmap_grob(bilisa_15_walk_suhii_map),
  tmap::tmap_grob(bilisa_15_walk_imd_map),

  tmap::tmap_grob(bilisa_15_cycle_suhii_map),
  tmap::tmap_grob(bilisa_15_cycle_imd_map)
)

# cowplot combine maps
bilisa_15_maps_combined <- cowplot::plot_grid(
  plotlist = bilisa_grobs,
  ncol = 2,
  labels = c("a", "b", "c", "d", "e", "f"),
  label_size = 12,
  align = "hv"
)

#bilisa_15_maps_combined
```

```{r}
# shared legend for plotting purposes

# legend names
legend_labels <- c(
  "Low access–High SUHII/IMD",
  "High access–High SUHII/IMD",
  "Low access–Low SUHII/IMD",
  "High access–Low SUHII/IMD"
)

# match each legend label with colours used in maps
legend_colours <- c(
  "Low access–High SUHII/IMD"  = "#B2182B",
  "High access–High SUHII/IMD" = "#FDDBC7",
  "Low access–Low SUHII/IMD"   = "#D1E5F0",
  "High access–Low SUHII/IMD"  = "#2166AC"
)

# make a shared legend
legend_data <- data.frame(
  cluster = factor(
    legend_labels,
    levels = legend_labels
  ),
  x = 1,
  y = 1:4
)
# plot the legend
legend_plot <- ggplot(
  legend_data,
  aes(x = x, y = y, fill = cluster)
) +
  geom_point(
    shape = 22,
    size = 8,
    colour = NA
  ) +
  scale_fill_manual(
    values = legend_colours,
    breaks = legend_labels,
    drop = FALSE,
    name = "BiLISA clusters"
  ) +
  guides(
    fill = guide_legend(
      override.aes = list(
        shape = 22,
        size = 8,
        colour = NA
      )
    )
  ) +
  theme_void() +
  theme(
    legend.position = "right",
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 10),
    legend.key.height = grid::unit(0.65, "cm"),
    legend.key.width = grid::unit(0.65, "cm"),
    legend.background = element_blank()
  )

# share legend on right
shared_legend <- cowplot::get_legend(
  legend_plot,
  legend = "right"
)
```

```{r}
# Combine the 6 maps with the shared legend
bilisa_15_body <- cowplot::plot_grid(
  bilisa_15_maps_combined,
  shared_legend,
  ncol = 2,
  rel_widths = c(0.82, 0.18)
)

# Create main title
bilisa_15_title <- cowplot::ggdraw() +
  cowplot::draw_label(
    "15-minute BiLISA Clusters of\nLow Accessibility-High SUHII or IMD",
    fontface = "bold",
    size = 14,
    hjust = 0.5,
    x = 0.5
  )

# Place the title above the maps
bilisa_15_final <- cowplot::plot_grid(
  bilisa_15_title,
  bilisa_15_body,
  ncol = 1,
  rel_heights = c(0.06, 0.94)
)

bilisa_15_final
```

```{r}
# save as png
ggsave(
  filename = "figures/FigD1_BiLISA_15_min_LH.png",
  plot = bilisa_15_final,
  width = 12.2,
  height = 9.5,
  units = "in",
  dpi = 300,
  bg = "white"
)
```

```{r}
# 60 min BiLISA maps
bilisa_grobs2 <- list(
  tmap::tmap_grob(bilisa_60_pt_suhii_map),
  tmap::tmap_grob(bilisa_60_pt_imd_map),

  tmap::tmap_grob(bilisa_60_walk_suhii_map),
  tmap::tmap_grob(bilisa_60_walk_imd_map),

  tmap::tmap_grob(bilisa_60_cycle_suhii_map),
  tmap::tmap_grob(bilisa_60_cycle_imd_map)
)

# cowplot combine maps
bilisa_60_maps_combined <- cowplot::plot_grid(
  plotlist = bilisa_grobs2,
  ncol = 2,
  labels = c("a", "b", "c", "d", "e", "f"),
  label_size = 12,
  align = "hv"
)
#bilisa_60_maps_combined
```

```{r}
# Combine the 6 maps with the shared legend
bilisa_60_body <- cowplot::plot_grid(
  bilisa_60_maps_combined,
  shared_legend,
  ncol = 2,
  rel_widths = c(0.82, 0.18)
)

# Create main title
bilisa_60_title <- cowplot::ggdraw() +
  cowplot::draw_label(
    "60-minute BiLISA Clusters of\nLow Accessibility-High SUHII or IMD",
    fontface = "bold",
    size = 14,
    hjust = 0.5,
    x = 0.5
  )

# Place the title above the maps
bilisa_60_final <- cowplot::plot_grid(
  bilisa_60_title,
  bilisa_60_body,
  ncol = 1,
  rel_heights = c(0.06, 0.94)
)

bilisa_60_final
```

```{r}
ggsave(
  filename = "figures/FigD2_BiLISA_60_min_LH.png",
  plot = bilisa_60_final,
  width = 12.2,
  height = 9.5,
  units = "in",
  dpi = 300,
  bg = "white"
)
```

### 7.2. Distribution of BiLISA Cluster

```{r}
# Put the 12 existing BiLISA objects into list
bilisa_results <- list(
  "15-minute Public Transit and SUHII" = bilisa_15_pt_suhii,
  "15-minute Public Transit and IMD" = bilisa_15_pt_imd,

  "15-minute Walking and SUHII" = bilisa_15_walk_suhii,
  "15-minute Walking and IMD" = bilisa_15_walk_imd,

  "15-minute Cycling and SUHII" = bilisa_15_cycle_suhii,
  "15-minute Cycling and IMD" = bilisa_15_cycle_imd,

  "60-minute Public Transit and SUHII" = bilisa_60_pt_suhii,
  "60-minute Public Transit and IMD" = bilisa_60_pt_imd,

  "60-minute Walking and SUHII" = bilisa_60_walk_suhii,
  "60-minute Walking and IMD" = bilisa_60_walk_imd,

  "60-minute Cycling and SUHII" = bilisa_60_cycle_suhii,
  "60-minute Cycling and IMD"   = bilisa_60_cycle_imd
)


# take cluster classifications
bilisa_long <- bind_rows(
  lapply(
    names(bilisa_results),
    function(analysis_name) {
      cluster_values <- as.character(
        bilisa_results[[analysis_name]]$clusters
      )
      # change NA as not significant 
      cluster_values[is.na(cluster_values)] <- "Not significant"
      # create a tibble
      tibble(
        Analysis = analysis_name,
        Cluster = cluster_values
      )
    }
  )
)
```

```{r}
# bilisa results table for appendix
analysis_order <- names(bilisa_results)

cluster_order <- c(
  "High-High",
  "High-Low",
  "Low-High",
  "Low-Low",
  "Not significant"
)

# Count LSOAs in each cluster type
bilisa_counts <- bilisa_long %>%
  dplyr::count(
    Analysis,
    Cluster,
    name = "n"
  ) %>%
  tidyr::complete(
    Analysis = analysis_order,
    Cluster = cluster_order,
    fill = list(n = 0)
  ) %>%
  # calculate percentage of LSOA in each cluster
  dplyr::group_by(Analysis) %>%
  dplyr::mutate(
    total_lsoas = sum(n),
    percentage = (n / total_lsoas) * 100,
    table_value = sprintf(
      "%s (%.1f%%)",
      scales::comma(n),
      percentage
    )
  ) %>%
  dplyr::ungroup()

# convert into 12-row table
bilisa_table <- bilisa_counts %>%
  # keep rows and columns in order
  dplyr::mutate(
    Analysis = factor(
      Analysis,
      levels = analysis_order
    ),
    Cluster = factor(
      Cluster,
      levels = cluster_order
    )
  ) %>%
  dplyr::arrange(
    Analysis,
    Cluster
  ) %>%
  # select only needed values to show in table
  dplyr::select(
    Analysis,
    Cluster,
    table_value
  ) %>%
  # make each cluster a new separate column
  tidyr::pivot_wider(
    names_from = Cluster,
    values_from = table_value
  ) %>%
  dplyr::select(
    Analysis,
    `High-High`,
    `High-Low`,
    `Low-High`,
    `Low-Low`,
    `Not significant`
  ) %>%
  dplyr::mutate(
    Analysis = as.character(Analysis)
  )

bilisa_table
```

### 7.3. Low High BiLISA

```{r}
# Function to compare the Low access–High SUHII and
# Low access–High IMD clusters for the same transport mode
create_priority_two_combined <- function(suhii_object, imd_object) {
  # SUHII clusters for each LSOA
  suhii_object$data %>%
    dplyr::select(
      lsoa21cd,
      suhii_cluster = bilisa_cluster
    ) %>%
    # join IMD BiLISA clusters by LSOA code
    dplyr::left_join(
      imd_object$data %>%
        sf::st_drop_geometry() %>%
        dplyr::select(
          lsoa21cd,
          imd_cluster = bilisa_cluster
        ),
      by = "lsoa21cd"
    ) %>%

    # identify whether each LSOA belongs to one of or both Low access–High SUHII / IMD
    dplyr::mutate(
      two_combined_class = dplyr::case_when(

        suhii_cluster == "Low-High" &
          imd_cluster == "Low-High" ~
          "High SUHII and IMD",

        suhii_cluster == "Low-High" ~
          "High SUHII",

        imd_cluster == "Low-High" ~
          "High IMD",

        TRUE ~ "Not Significant"
      ),

      # Set the legend order
      two_combined_class = factor(
        two_combined_class,
        levels = c(
          "High SUHII and IMD",
          "High SUHII",
          "High IMD",
          "Not Significant"
        )
      )
    )
}
```

```{r}
# set colours for L-H combined maps
two_combined_colours <- c(
  "High SUHII and IMD" = "#5F2022",
  "High SUHII"         = "#AD423B",
  "High IMD"           = "#DC9179",
  "Not Significant"    = "#FBF8F2"
)


# function to create combined map
create_two_combined_map <- function(data) {
  # clusters
  tm_shape(data) +
    tm_fill(
      fill = "two_combined_class",
      fill.scale = tm_scale_categorical(
        values = two_combined_colours
      ),
      fill.legend = tm_legend_hide()
    ) +
    # greater london boundary
    tm_shape(london_boundary) +
    tm_borders(
      col = "grey30",
      lwd = 1
    ) +
    tm_layout(
      frame = FALSE,
      bg.color = "white",
      outer.margins = 0
    )
}

# accessibility for combined Low-High
combined_access_names <- c(
  "15_pt",
  "15_walk",
  "15_cycle",
  "60_pt",
  "60_walk",
  "60_cycle"
)


# create combined SUHII / IMD clusters for each accessibility 
for (access_name in combined_access_names) {
  # take SUHII BiLISA 
  suhii_object <- get(
    paste0(
      "bilisa_",
      access_name,
      "_suhii"
    )
  )
  # take IMD BiLISA
  imd_object <- get(
    paste0(
      "bilisa_",
      access_name,
      "_imd"
    )
  )
  # combine Low-High SUHII and IMD 
  combined_result <- create_priority_two_combined(
    suhii_object = suhii_object,
    imd_object = imd_object
  )

  # combined results with the accessibility name
  assign(
    paste0(
      "two_combined_",
      access_name
    ),
    combined_result
  )

  # create combined map
  combined_map <- create_two_combined_map(
    combined_result
  )

  # object name
  assign(
    paste0(
      "two_combined_",
      access_name,
      "_map"
    ),
    combined_map
  )
}
```

```{r}
# 15 minute BiLISA L-H maps
two_combined_15_grobs <- list(
  tmap::tmap_grob(two_combined_15_pt_map),
  tmap::tmap_grob(two_combined_15_walk_map),
  tmap::tmap_grob(two_combined_15_cycle_map)
)


# Combine maps
two_combined_15_maps <- cowplot::plot_grid(
  plotlist = two_combined_15_grobs,
  ncol = 1,
  labels = c("a", "b", "c"),
  label_size = 12,
  align = "v",
  axis = "lr"
)


# make a shared legend
two_combined_legend_data <- data.frame(
  two_combined_class = factor(
    names(two_combined_colours),
    levels = names(two_combined_colours)
  ),
  x = 1,
  y = seq_along(two_combined_colours)
)
# plot the legend
two_combined_legend_plot <- ggplot(
  two_combined_legend_data,
  aes(x = x, y = y, fill = two_combined_class)
) +
  geom_point(
    shape = 22,
    size = 8,
    colour = NA
  ) +
  scale_fill_manual(
    values = two_combined_colours,
    breaks = names(two_combined_colours),
    drop = FALSE,
    name = "BiLISA Clusters"
  ) +
  guides(
    fill = guide_legend(
      override.aes = list(
        shape = 22,
        size = 8,
        colour = NA
      )
    )
  ) +
  theme_void() +
  theme(
    legend.position = "right",
    legend.title = element_text(size = 11),
    legend.text = element_text(size = 9),
    legend.key.height = grid::unit(0.65, "cm"),
    legend.key.width = grid::unit(0.65, "cm")
  )

# legend on right
shared_two_combined_legend <- cowplot::get_legend(
  two_combined_legend_plot,
  legend = "right"
)
```

```{r}
# combine maps and add shared legend
two_combined_15_final_body <- cowplot::plot_grid(
  two_combined_15_maps,
  shared_two_combined_legend,
  ncol = 2,
  rel_widths = c(0.76, 0.24)
)


# Add overall figure title
two_combined_15_final <- cowplot::ggdraw() +
  cowplot::draw_label(
    paste0(
      "15-minute BiLISA Clusters of\n",
      "Low Accessibility–High SUHII and/or High IMD"
    ),
    x = 0.5,
    y = 0.98,
    hjust = 0.5,
    fontface = "bold",
    size = 14
  ) +
  cowplot::draw_plot(
    two_combined_15_final_body,
    x = 0,
    y = 0,
    width = 1,
    height = 0.93
  )


# Display final figure
two_combined_15_final
```

```{r}
# save as PNG
ggsave(
  filename = "figures/Fig13_BiLISA_15_min_LH_Combined.png",
  plot = two_combined_15_final,
  width = 9,
  height = 11,
  units = "in",
  dpi = 300,
  bg = "white"
)
```

```{r}
# 60 minute BiLISA L-H maps
two_combined_60_grobs <- list(
  tmap::tmap_grob(two_combined_60_pt_map),
  tmap::tmap_grob(two_combined_60_walk_map),
  tmap::tmap_grob(two_combined_60_cycle_map)
)

# combine maps
two_combined_60_maps <- cowplot::plot_grid(
  plotlist = two_combined_60_grobs,
  ncol = 1,
  labels = c("a", "b", "c"),
  label_size = 12,
  align = "v",
  axis = "lr"
)


# share legend  
two_combined_60_final_body <- cowplot::plot_grid(
  two_combined_60_maps,
  shared_two_combined_legend,
  ncol = 2,
  rel_widths = c(0.76, 0.24)
)


# add main title
two_combined_60_final <- cowplot::ggdraw() +
  cowplot::draw_label(
    paste0(
      "60-minute BiLISA Clusters of\n",
      "Low Accessibility–High SUHII and/or High IMD"
    ),
    x = 0.5,
    y = 0.98,
    hjust = 0.5,
    fontface = "bold",
    size = 14
  ) +
  cowplot::draw_plot(
    two_combined_60_final_body,
    x = 0,
    y = 0,
    width = 1,
    height = 0.93
  )


two_combined_60_final
```

```{r}
# save as PNG
ggsave(
  filename = "figures/Fig14_BiLISA_60_min_LH_Combined.png",
  plot = two_combined_60_final,
  width = 9,
  height = 11,
  units = "in",
  dpi = 300,
  bg = "white"
)
```

### 7.4. Summary of overlap L-H clusters of High SUHII and IMD

```{r}
overlap_objects <- list(
  "15-minute Public Transit" = two_combined_15_pt,
  "15-minute Walking" = two_combined_15_walk,
  "15-minute Cycling" = two_combined_15_cycle,
  "60-minute Public Transit" = two_combined_60_pt,
  "60-minute Walking" = two_combined_60_walk,
  "60-minute Cycling" = two_combined_60_cycle
)

# calculate number and percentage of London LSOAs that overlap SUHII and IMD in L-H
overlap_summary <- dplyr::bind_rows(
  
  lapply(names(overlap_objects), function(x) {
    
    data <- overlap_objects[[x]] %>%
      sf::st_drop_geometry()
    
    # table of one row per variable
    tibble::tibble(
      Accessibility = x,
      
      # the overlap of SUHII and IMD
      overlap_n = sum(
        data$two_combined_class == "High SUHII and IMD",
        na.rm = TRUE
      ),
      # number of LSOAs
      total_lsoas = nrow(data),
      
      # calculate the percentage
      overlap_percentage = 
        overlap_n / total_lsoas * 100
    )
    
  })
)

overlap_summary
```
