---
title: "6_Statistical_Analyses"
---

```{r}
# remove objects from R environment and clear memory
rm(list = ls())    
gc()
```

## Load Packages

```{r}
library(sf)
library(sp)
library(readr)
library(dplyr)
library(knitr)
library(tidyverse)
library(tmap)
library(cowplot)
library(plotly)
library(corrplot)
library(broom)
library(mapview)
library(stringr)
library(spdep)
library(car)
library(fs)
library(janitor)
library(GGally)
library(MASS)
library(spatialreg)
library(INLA)
library(INLAtools)
library(performance)
```

## Data Loading

Clean names, select initial needed columns and rename columns for consistency across datasets

**Greater London Boundary:**

```{r}
london_boundary <- st_read("data/gla/London_GLA_Boundary.shp")
```

**London only LSOAs**

```{r}
london_lsoa <- st_read("data/lsoa_london/lsoa_london.shp")
```

**Data obtained from previous pre-processing**

```{r}
# travel time 
multimodal_accessibility_lsoa <- st_read("data/multimodal_accessibility_lsoa.gpkg")%>%
  clean_names()%>%
  dplyr::select(
    lsoa21cd,
    lsoa21nm,
    access_score_15_pt,
    access_score_60_pt,
    access_score_15_walk,
    access_score_60_walk,
    access_score_15_cycle,
    access_score_60_cycle,
    travel_time_min_pt,
    travel_time_min_walk,
    travel_time_min_cycle,
    geom
  )%>%
  rename(
    geometry = geom
  )
```

```{r}
# SUHII as lsoa heat exposure
suhii_lsoa <- st_read("data/London_LSOA_LST_SUHI/London_LSOA_LST_SUHI.shp")%>% 
  st_transform(.,27700)%>%
  clean_names()%>%
  dplyr::select(
    lsoa21cd,
    lsoa21nm,
    mean,
    london_mea,
    lst_differ,
    geometry
  )%>%
  rename(
    mean_lst_lsoa = mean,
    london_mean = london_mea,
    suhii = lst_differ
  )
```

**IMD 2025 data**

```{r}
# IMD score
imd_score <- read_csv("data/IMD_2025.csv")%>%
  clean_names()%>%
  dplyr::select(
    lsoa_code_2021,
    lsoa_name_2021,
    index_of_multiple_deprivation_imd_score
    )%>%
  rename(
    lsoa21cd = lsoa_code_2021,
    lsoa21nm = lsoa_name_2021,
    imd_score = index_of_multiple_deprivation_imd_score
  )
```

**Census Data**

```{r}
# TS006 population density
census_pop_density <- read_csv("data/census2021/census2021-ts006-lsoa.csv")%>%
  clean_names()%>%
  dplyr::select(
    geography,
    geography_code,
    population_density_persons_per_square_kilometre_measures_value
  )%>%
  rename(
    lsoa21cd = geography_code,
    lsoa21nm = geography,
    pop_density = population_density_persons_per_square_kilometre_measures_value
  )
```

```{r}
# TS007a age by 5 year band
census_age_5year_band <- read_csv("data/census2021/census2021-ts007a-lsoa.csv")%>%
  clean_names()%>%
  dplyr::select(
    geography,
    geography_code,
    age_total,
    age_aged_4_years_and_under,
    # for filter for age 65 or over 
    age_aged_65_to_69_years,
    age_aged_70_to_74_years,
    age_aged_75_to_79_years,
    age_aged_80_to_84_years,
    age_aged_85_years_and_over
  )%>%
  rename(
    lsoa21cd = geography_code,
    lsoa21nm = geography,
    age_4_and_under = age_aged_4_years_and_under,
    age_65_to_69 = age_aged_65_to_69_years,
    age_70_to_74 = age_aged_70_to_74_years,
    age_75_to_79 = age_aged_75_to_79_years,
    age_80_to_84 = age_aged_80_to_84_years,
    age_85_and_over = age_aged_85_years_and_over
  )
```

```{r, eval=FALSE}
# TS038 disability
census_disability <- read_csv("data/census2021/census2021-ts038-lsoa.csv")%>%
  clean_names()%>%
  select(
    geography,
    geography_code,
    disability_total_all_usual_residents,
    disability_disabled_under_the_equality_act
  )%>%
  rename(
    lsoa21cd = geography_code,
    lsoa21nm = geography,
    # total number of residents 
    disability_total_residents = disability_total_all_usual_residents,
    # disabled under the Equality Act 2010
    disabled = disability_disabled_under_the_equality_act
  )
```

```{r, eval=FALSE}
# TS037 general health
census_health <- read_csv("data/census2021/census2021-ts037-lsoa.csv")%>% 
  clean_names()%>%
  select(
    geography,
    geography_code,
    general_health_total_all_usual_residents,
    general_health_bad_health,
    general_health_very_bad_health
  )%>%
  rename(
    lsoa21cd = geography_code,
    lsoa21nm = geography,
    general_health_total_residents = general_health_total_all_usual_residents,
    general_health_bad = general_health_bad_health,
    general_health_very_bad = general_health_very_bad_health
  )
```

```{r}
# TS021 Ethnicity
# definition of ethnic minority groups
# https://www.ethnicity-facts-figures.service.gov.uk/style-guide/writing-about-ethnicity/ 
# https://service-manual.ons.gov.uk/content/language/ethnicity-and-race
census_ethnicity <- read_csv("data/census2021/census2021-ts021-lsoa.csv")%>%
  clean_names()%>%
  dplyr::select(
    geography,
    geography_code,
    ethnic_group_total_all_usual_residents,
    # Use “Ethnic minorities” to refer to all ethnic groups except the White British group
    ethnic_group_white_english_welsh_scottish_northern_irish_or_british
  )%>%
  rename(
    lsoa21cd = geography_code,
    lsoa21nm = geography,
    ethnic_group_total_residents = ethnic_group_total_all_usual_residents,
    white_british_group = ethnic_group_white_english_welsh_scottish_northern_irish_or_british
  )
```

## 1. Data Cleaning

### Check Loaded + Initial Cleaned Data Column Names

```{r}
names(london_lsoa)
```

```{r}
# check 3 modes 
names(multimodal_accessibility_lsoa)
```

```{r}
# check names in SUHII export
names(suhii_lsoa)
```

```{r}
# check IMD names
names(imd_score)
```

```{r}
# check census data
names(census_pop_density)
```

```{r}
names(census_age_5year_band)
```

```{r}
names(census_ethnicity)
```

### Wrangle data variables for regression modelling

```{r}
# filter london lsoas only
london_lsoa_codes <- london_lsoa %>%
  clean_names() %>%
  pull(lsoa21cd)
```

```{r}
census_age_cleaned <- census_age_5year_band %>%
  mutate(
    age_4_and_under_pct = ifelse(
      age_total > 0,
      (age_4_and_under / age_total) * 100,
      NA
    ),
    age_65_and_over = rowSums(
      across(
        c(
          age_65_to_69,
          age_70_to_74,
          age_75_to_79,
          age_80_to_84,
          age_85_and_over
        )
      ),
      na.rm = TRUE
    ),
    age_65_and_over_pct = ifelse(
      age_total > 0,
      (age_65_and_over / age_total)* 100,
      NA
    )
  )%>% 
  filter(lsoa21cd %in% london_lsoa_codes)
```

```{r, eval=FALSE}
census_disability_cleaned <- census_disability %>%
  mutate(
    disability_pct = ifelse(
      disability_total_residents > 0,
      (disabled / disability_total_residents)* 100,
      NA
    )
  )%>% 
  filter(lsoa21cd %in% london_lsoa_codes)
```

```{r, eval=FALSE}
census_health_cleaned <- census_health %>%
  mutate(
    health_bad_verybad = rowSums(
      across(
        c(
          general_health_bad,
          general_health_very_bad
        )
      ),
      na.rm = TRUE
    ),
    health_bad_verybad_pct = ifelse(
      general_health_total_residents > 0,
      (health_bad_verybad / general_health_total_residents)* 100,
      NA
    )
  )%>% 
  filter(lsoa21cd %in% london_lsoa_codes)
```

```{r}
census_ethnicity_cleaned <- census_ethnicity %>%
   mutate(
    ethnic_minority =
      ethnic_group_total_residents - white_british_group,
    
    ethnic_minority_pct = ifelse(
      ethnic_group_total_residents > 0,
      (ethnic_minority / ethnic_group_total_residents) * 100,
      NA
    )
  )%>% 
  filter(lsoa21cd %in% london_lsoa_codes)
```

```{r}
imd_score_cleaned <- imd_score %>% filter(lsoa21cd %in% london_lsoa_codes)
```

### Join Data

```{r}
london_join <- multimodal_accessibility_lsoa %>%
  # SUHII 
  left_join(
    suhii_lsoa %>%
      # drop geom cus multimodal travel time already have it
      st_drop_geometry() %>%
      # select needed columns
      dplyr::select(lsoa21cd, suhii),
    # join by lsoa code
    by = "lsoa21cd"
  ) %>%
  
  # IMD 
  left_join(
    imd_score_cleaned %>%
      dplyr::select(lsoa21cd, imd_score),
    by = "lsoa21cd"
  ) %>%
  
  # population density
  left_join(
    census_pop_density %>%
      dplyr::select(lsoa21cd, pop_density),
    by = "lsoa21cd"
  ) %>%
  
  # age
  left_join(
    census_age_cleaned %>%
      dplyr::select(lsoa21cd, age_4_and_under_pct, age_65_and_over_pct),
    by = "lsoa21cd"
  ) %>%
  
  # disability
  #left_join(
    #census_disability_cleaned %>%
      #select(lsoa21cd, disability_pct),
    #by = "lsoa21cd"
  #) %>%
  
  # health
  #left_join(
    #census_health_cleaned %>%
      #select(lsoa21cd, health_bad_verybad_pct),
    #by = "lsoa21cd"
  #) %>%
  
  # ethnicity
  left_join(
    census_ethnicity_cleaned %>%
      dplyr::select(lsoa21cd, ethnic_minority_pct),
    by = "lsoa21cd"
  )
```

```{r, eval=FALSE}
# check joined table
london_join %>% 
  st_drop_geometry() %>% 
  View()
```

```{r}
names(london_join)
```

```{r}
# create inner london control variable
# following: https://www.london.gov.uk/programmes-strategies/planning/london-plan/the-london-plan-2021-online/annex-2-inner-and-outer-london-boroughs
inner_london_boroughs <- c(
  "City of London", "Camden", "Greenwich", "Hackney",
  "Hammersmith and Fulham", "Islington", "Kensington and Chelsea",
  "Lambeth", "Lewisham", "Newham", "Southwark", "Tower Hamlets",
  "Wandsworth", "Westminster"
)

london_join <- london_join %>%
  mutate(
    #remove the letters numbers behind borough names to keep it a clean borough name only
    borough = str_remove(lsoa21nm, "\\s\\d+[A-Z]$"),
    # inner and outer london categories
    inner_outer_london = if_else(borough %in% inner_london_boroughs, "Inner", "Outer"),
    inner_outer_london = factor(inner_outer_london, levels = c("Outer", "Inner"))
  )
```

```{r}
#check
names(london_join)
```

## 2. Regression Data Preparation

```{r}
# check data types
check_data <- london_join %>%
  st_drop_geometry() %>%
  summarise(across(everything(), ~class(.)[1])) %>%
  pivot_longer(everything(), names_to = "variable", values_to = "type")

view(check_data)
```

```{r}
# descriptive stats
london_join %>%
  st_drop_geometry() %>%
  dplyr::select(where(is.numeric)) %>%
  summarise(across(everything(),
                   list(mean = ~mean(., na.rm = TRUE),
                        sd = ~sd(., na.rm = TRUE),
                        min = ~min(., na.rm = TRUE),
                        max = ~max(., na.rm = TRUE)))) %>%
  pivot_longer(everything(),
               names_to = c("variable", "stat"),
               names_pattern = "(.*)_(mean|sd|min|max)") %>%
  pivot_wider(names_from = stat, values_from = value) %>%
  kable(digits = 2)
```

## 3. Exploratory Data Analysis

**Plot Variables Distribution**

```{r}
# Select only numeric columns
numeric_cols <- london_join %>%
  sf::st_drop_geometry() %>%
  dplyr::select(where(is.numeric))

# Convert to long format for faceting
numeric_long <- numeric_cols %>%
  pivot_longer(cols = everything(), names_to = "variable", values_to = "value")

# Plot faceted histograms
variables_dist_hist <- ggplot(numeric_long, aes(x = value)) +
  geom_histogram(bins = 30, fill = "steelblue", color = "white") +
  facet_wrap(~ variable, scales = "free", ncol = 3) +
  theme_minimal() +
  labs(title = "Histograms of Numerical Variables")

variables_dist_hist
```

```{r, eval=FALSE}
# save as PNG
ggsave(
  filename = "figures/Fig_Appendix_Variables_Distribution_Hist.png",
  plot = variables_dist_hist,
  width = 7.5,
  height = 10,
  units = "in",
  dpi = 300,
  bg = "white"
)
```

**Plot Heatmap Correlation (Spearman)**

```{r}
# Select only the numeric variables (drop codes, names, geometry)
corr_matrix_spearman <- london_join %>%
  st_drop_geometry() %>%
  dplyr::select(
    access_score_15_pt,
    access_score_60_pt,
    access_score_15_walk,
    access_score_60_walk,
    access_score_15_cycle,
    access_score_60_cycle,
    imd_score,
    suhii,
    pop_density,
    age_4_and_under_pct,
    age_65_and_over_pct,
    ethnic_minority_pct
  ) %>%
  cor(
    use = "pairwise.complete.obs",
    method = "spearman"
  )

heatmap <- corrplot(
  corr_matrix_spearman,
  method = "color",
  type = "full",
  col = rev(COL2("RdBu", 200)),
  tl.col = "black",
  tl.srt = 45,
  addCoef.col = "black",
  number.cex = 0.6
)
```

```{r, eval=FALSE}
# save corrplot as png

dir.create("figures", showWarnings = FALSE)

png(
  filename = "figures/Fig_Appendix_Spearman_Heatmap.png",
  width = 8.5,
  height = 5.5,
  units = "in",
  res = 300,
  bg = "white"
)

corrplot(
  corr_matrix_spearman,
  method = "color",
  type = "full",
  col = rev(COL2("RdBu", 200)),
  tl.col = "black",
  tl.srt = 45,
  addCoef.col = "black",
  number.cex = 0.6
)

dev.off()
```

```{r}
outcomes <- c(
  "access_score_15_pt",
  "access_score_60_pt",
  "access_score_15_walk",
  "access_score_60_walk",
  "access_score_15_cycle",
  "access_score_60_cycle"
)

distribution_check <- do.call(
  rbind,
  lapply(outcomes, function(v) {

    y <- london_join[[v]]

    data.frame(
      outcome = v,
      n = sum(!is.na(y)),
      mean = mean(y, na.rm = TRUE),
      variance = var(y, na.rm = TRUE),
      variance_mean_ratio = var(y, na.rm = TRUE) /
        mean(y, na.rm = TRUE),
      zeros = sum(y == 0, na.rm = TRUE),
      zero_percent = mean(y == 0, na.rm = TRUE) * 100,
      maximum = max(y, na.rm = TRUE)
    )
  })
)

distribution_check
```

### Transform Variables

**Log transform population density and standardise other variables for clean interpretation**

This is due to the above examination of data, population density is highly skewed (so, log transform) and standardise all the variables that are in percentages.

```{r}
model_variables <- london_join %>%
  mutate(
    # apply log transforms
    log_pop_density = log(pop_density), 

    # standardise --> z-scores
    imd_z = as.numeric(scale(imd_score)),
    suhii_z = as.numeric(scale(suhii)),
    log_pop_density_z = as.numeric(scale(log_pop_density)),
    age_4_under_z = as.numeric(scale(age_4_and_under_pct)),
    age_65_over_z = as.numeric(scale(age_65_and_over_pct)),
    ethnic_minority_z  = as.numeric(scale(ethnic_minority_pct))
  )
```

```{r}
names(model_variables)
```

## 4. Model Choices

<https://rpubs.com/Geesaale/1263990> –\> count regression: poisson or negative binomial, depending on the nature of spatial dependence of data, spatial error more suitable (spatial error models is suitable when the spatial autocorrelation is due to an underlying spatial process or unmeasured factors).

## 5. Negative Binomial GLM

Loop all models:

```{r}
library(MASS)

# The predictors are the same for every model
predictors <- c(
  "imd_z * suhii_z",
  "log_pop_density_z",
  "age_4_under_z",
  "age_65_over_z",
  "ethnic_minority_z",
  "inner_outer_london"
)

# The six outcomes to loop over
outcomes <- c(
  #"access_score_15_pt",
  "access_score_60_pt",
  #"access_score_15_walk",
  "access_score_60_walk",
  "access_score_15_cycle",
  "access_score_60_cycle"
)

# Function to fit one negative binomial GLM given an outcome name
fit_nb <- function(outcome, data, predictors) {
  f <- reformulate(predictors, response = outcome)
  glm.nb(f, data = data)
}

# Loop over all outcomes, returning a named list of fitted models
models_nb <- lapply(outcomes, fit_nb, data = model_variables, predictors = predictors)
names(models_nb) <- outcomes
```

```{r}
# Print a labelled summary for each of the six models
for (nm in names(models_nb)) {
  cat("\n=====================================================\n")
  cat("MODEL:", nm, "\n")
  cat("=====================================================\n")
  print(summary(models_nb[[nm]]))
}
```

## 6. Poisson Regression

```{r}
# The predictors are the same for every model
predictors <- c(
  "imd_z * suhii_z",
  "log_pop_density_z",
  "age_4_under_z",
  "age_65_over_z",
  "ethnic_minority_z",
  "inner_outer_london"
)

# The six outcomes to loop over
outcomes <- c(
  "access_score_15_pt",
  #"access_score_60_pt",
  "access_score_15_walk"
  #"access_score_60_walk",
  #"access_score_15_cycle",
  #"access_score_60_cycle"
)

# Function to fit one Poisson GLM given an outcome name
fit_poisson <- function(outcome, data, predictors) {
  f <- reformulate(predictors, response = outcome)
  glm(f, data = model_variables, family = poisson(link = "log"))
}

# Loop over all outcomes, returning a named list of fitted models
models_poisson <- lapply(outcomes, fit_poisson, data = model_variables, predictors = predictors)
names(models_poisson) <- outcomes
```

```{r}
for (nm in names(models_poisson)) {
  cat("\n=====================================================\n")
  cat("MODEL:", nm, "\n")
  cat("=====================================================\n")
  print(summary(models_poisson[[nm]]))
}
```

```{r}
# calculate Poisson dispersion
check_dispersion <- function(model) {

  pearson_chisq <- sum(
    residuals(model, type = "pearson")^2,
    na.rm = TRUE
  )

  residual_df <- df.residual(model)

  dispersion_ratio <- pearson_chisq / residual_df

  # approximate test of overdispersion
  p_value <- pchisq(
    pearson_chisq,
    df = residual_df,
    lower.tail = FALSE
  )

  data.frame(
    pearson_chisq = pearson_chisq,
    residual_df = residual_df,
    dispersion_ratio = dispersion_ratio,
    p_value = p_value
  )
}

# apply to both Poisson models
dispersion_results <- do.call(
  rbind,
  lapply(models_poisson, check_dispersion)
)

# add model names
dispersion_results$model <- rownames(dispersion_results)
rownames(dispersion_results) <- NULL

# reorder columns
dispersion_results <- dispersion_results[
  c(
    "model",
    "dispersion_ratio",
    "pearson_chisq",
    "residual_df",
    "p_value"
  )
]

dispersion_results
```

Iteration limit for access_score_15_pt and access_score_15_walk, tried on Poisson across all models. Checked AIC values as below table (bold text refers to lower AIC values), their AIC values are also higher in the the negative binomial regression. As well as dispersion checked as above, dispersion values close to 1.

| Dependent variable    | AIC - Negative Binomial | AIC - Poisson |
|-----------------------|-------------------------|---------------|
| access_score_15_pt    | 13934                   | **13932**     |
| access_score_60_pt    | **49596**               | 90101         |
| access_score_15_walk  | 16249                   | **16247**     |
| access_score_60_walk  | **32617**               | 33312         |
| access_score_15_cycle | **21932**               | 22076         |
| access_score_60_cycle | **42929**               | 55092         |

## 7. Residuals with Moran's I

**Build spatial weights with LSOA boundary**

```{r}
#calculate the centroids of all LSOA in London
coordsLSOA <- london_lsoa%>%
  st_centroid()%>%
  st_geometry()

plot(coordsLSOA)
```

```{r}
#Now we need to generate a spatial weights matrix 
#(remember from the lecture a couple of weeks ago). 
#We'll start with a simple binary matrix of queen's case neighbours

LLSOA_nb <- london_lsoa %>%
  poly2nb(., queen=T)

#or nearest neighbours
knn_LSOA <-coordsLSOA %>%
  knearneigh(., k=4)

LLSOA_knn <- knn_LSOA %>%
  knn2nb()
```

```{r, eval=FALSE}
#plot them
plot(LLSOA_nb, st_geometry(coordsLSOA), col="red")
```

```{r, eval=FALSE}
plot(LLSOA_knn, st_geometry(coordsLSOA), col="blue")
```

```{r}
#create a spatial weights matrix object from these weights

LLSOA.queens_weight <- LLSOA_nb %>%
  nb2listw(., style="W")

LLSOA.knn_4_weight <- LLSOA_knn %>%
  nb2listw(., style="W")
```

### 7.1. access_score_15_pt

```{r}
# name object first
model_15_pt <- models_poisson$access_score_15_pt
# VIF
vif(model_15_pt)
```

Ok, VIF values are all below 5 or 10, no problem with multicolinearity, next look at residuals (auto)correlation

```{r}
# residuals with pearson
model_15_pt_resids <- residuals(model_15_pt, "pearson")

model_profile <- model_variables %>%
  mutate(model_15_pt_resids)

#now plot the residuals
tmap_mode("view")
qtm(model_profile, fill = "model_15_pt_resids")
```

```{r}
# queens case
model_15_pt_resids_queen <- model_profile %>%
  st_drop_geometry()%>%
  dplyr::select(model_15_pt_resids)%>%
  pull()%>%
  moran.test(., LLSOA.queens_weight)%>%
  tidy()

model_15_pt_resids_queen
```

```{r}
# nearest neighbour
model_15_pt_resids_nearest_neighbour <- model_profile %>%
  st_drop_geometry()%>%
  dplyr::select(model_15_pt_resids)%>%
  pull()%>%
  moran.test(., LLSOA.knn_4_weight)%>%
  tidy()

model_15_pt_resids_nearest_neighbour
```

### 7.2. access_score_60_pt

```{r}
# name object first
model_60_pt <- models_nb$access_score_60_pt

# VIF
vif(model_60_pt)
```

**Residuals and Map**

```{r}
# residuals with pearson
model_60_pt_resids <- residuals(model_60_pt, "pearson")

model_profile <- model_variables %>%
  mutate(model_60_pt_resids)

#now plot the residuals
tmap_mode("view")
qtm(model_profile, fill = "model_60_pt_resids")
```

**Moran's I - Queens Case**

```{r}
# queens case
model_60_pt_resids_queen <- model_profile %>%
  st_drop_geometry()%>%
  dplyr::select(model_60_pt_resids)%>%
  pull()%>%
  moran.test(., LLSOA.queens_weight)%>%
  tidy()

model_60_pt_resids_queen
```

**Moran's I - 4 Nearest Neighbours**

```{r}
# nearest neighbour
model_60_pt_resids_nearest_neighbour <- model_profile %>%
  st_drop_geometry()%>%
  dplyr::select(model_60_pt_resids)%>%
  pull()%>%
  moran.test(., LLSOA.knn_4_weight)%>%
  tidy()

model_60_pt_resids_nearest_neighbour
```

### 7.3. access_score_15_walk (Poisson)

```{r}
# name object first
model_15_walk <- models_poisson$access_score_15_walk
# VIF
vif(model_15_walk)
```

**Residuals and Map**

```{r}
# residuals with pearson
model_15_walk_resids <- residuals(model_15_walk, "pearson")

model_profile <- model_variables %>%
  mutate(model_15_walk_resids)

#now plot the residuals
tmap_mode("view")
qtm(model_profile, fill = "model_15_walk_resids")
```

**Moran's I - Queens Case**

```{r}
# queens case
model_15_walk_resids_queen <- model_profile %>%
  st_drop_geometry()%>%
  dplyr::select(model_15_walk_resids)%>%
  pull()%>%
  moran.test(., LLSOA.queens_weight)%>%
  tidy()

model_15_walk_resids_queen
```

**Moran's I - 4 Nearest Neighbours**

```{r}
# nearest neighbour
model_15_walk_resids_nearest_neighbour <- model_profile %>%
  st_drop_geometry()%>%
  dplyr::select(model_15_walk_resids)%>%
  pull()%>%
  moran.test(., LLSOA.knn_4_weight)%>%
  tidy()

model_15_walk_resids_nearest_neighbour
```

### 7.4. access_score_60_walk

```{r}
# name object first
model_60_walk <- models_nb$access_score_60_walk
# VIF
vif(model_60_walk)
```

**Residuals and Map**

```{r}
# residuals with pearson 
model_60_walk_resids <- residuals(model_60_walk, "pearson")  

model_profile <- model_variables %>%   
  mutate(model_60_walk_resids)  

#now plot the residuals 
tmap_mode("view") 
qtm(model_profile, fill = "model_60_walk_resids")
```

**Moran's I - Queens Case**

```{r}
# queens case 
model_60_walk_resids_queen <- model_profile %>%   
  st_drop_geometry()%>%   
  dplyr::select(model_60_walk_resids)%>%   
  pull()%>%   
  moran.test(., LLSOA.queens_weight)%>%   
  tidy()  

model_60_walk_resids_queen
```

**Moran's I - 4 Nearest Neighbours**

```{r}
# nearest neighbour 
model_60_walk_resids_nearest_neighbour <- model_profile %>%   
  st_drop_geometry()%>%   
  dplyr::select(model_60_walk_resids)%>%   
  pull()%>%   
  moran.test(., LLSOA.knn_4_weight)%>%   
  tidy()  

model_60_walk_resids_nearest_neighbour
```

### 7.5. access_score_15_cycle

```{r}
# name object first
model_15_cycle <- models_nb$access_score_15_cycle
# VIF
vif(model_15_cycle)
```

**Residuals and Map**

```{r}
# residuals with pearson  
model_15_cycle_resids <- residuals(model_15_cycle, "pearson")    

model_profile <- model_variables %>%      
  mutate(model_15_cycle_resids)    

#now plot the residuals  
tmap_mode("view")  
qtm(model_profile, fill = "model_15_cycle_resids")
```

**Moran's I - Queens Case**

```{r}
# queens case  
model_15_cycle_resids_queen <- model_profile %>%      
  st_drop_geometry()%>%      
  dplyr::select(model_15_cycle_resids)%>%      
  pull()%>%      
  moran.test(., LLSOA.queens_weight)%>%      
  tidy()    

model_15_cycle_resids_queen
```

**Moran's I - 4 Nearest Neighbours**

```{r}
# nearest neighbour  
model_15_cycle_resids_nearest_neighbour <- model_profile %>%      
  st_drop_geometry()%>%      
  dplyr::select(model_15_cycle_resids)%>%      
  pull()%>%      
  moran.test(., LLSOA.knn_4_weight)%>%      
  tidy()    

model_15_cycle_resids_nearest_neighbour
```

### 7.6. access_score_60_cycle

```{r}
# name object first
model_60_cycle <- models_nb$access_score_60_cycle
# VIF
vif(model_60_cycle)
```

**Residuals and Map**

```{r}
# residuals with pearson   
model_60_cycle_resids <- residuals(model_60_cycle, "pearson")      

model_profile <- model_variables %>%         
  mutate(model_60_cycle_resids)      

#now plot the residuals   
tmap_mode("view")   
qtm(model_profile, fill = "model_60_cycle_resids")
```

**Moran's I - Queens Case**

```{r}
# queens case   
model_60_cycle_resids_queen <- model_profile %>%         
  st_drop_geometry()%>%         
  dplyr::select(model_60_cycle_resids)%>%         
  pull()%>%         
  moran.test(., LLSOA.queens_weight)%>%         
  tidy()      

model_60_cycle_resids_queen
```

**Moran's I - 4 Nearest Neighbours**

```{r}
# nearest neighbour   
model_60_cycle_resids_nearest_neighbour <- model_profile %>%         
  st_drop_geometry()%>%         
  dplyr::select(model_60_cycle_resids)%>%         
  pull()%>%         
  moran.test(., LLSOA.knn_4_weight)%>%         
  tidy()      

model_60_cycle_resids_nearest_neighbour
```

### 7.7. Conclusion of Models' residuals

| Model          | Family            | Moran's I (Queens) | Moran's I (knn) |
|----------------|-------------------|--------------------|-----------------|
| model_15_pt    | Poisson           | 0.5119248          | 0.5322355       |
| model_60_pt    | Negative Binomial | 0.6729391          | 0.6872948       |
| model_15_walk  | Poisson           | 0.6667913          | 0.6918659       |
| model_60_walk  | Negative Binomial | 0.9055903          | 0.9090586       |
| model_15_cycle | Negative Binomial | 0.7601062          | 0.7765267       |
| model_60_cycle | Negative Binomial | 0.9209963          | 0.9120826       |

Strong positive spatial autocorrelation for queens case and nearest neighbours by Moran's I for all models, so next use spatial models

## 8. Global Spatial Regression model (BYM2)

Given that the data type is **count data**, traditional **spatial error** models **may not be suitable** here as a direct transfer of the spatial model types for continuous data is not possible (Reference: <https://www.econstor.eu/bitstream/10419/168040/1/89655662X.pdf>).

Bayesian spatial multilevel mixed-effects model incoporating BMY2 with `INLA` package

```{r}
# https://www.r-inla.org/download/index.html

install.packages("INLA", repos=c(getOption("repos"), INLA="https://inla.r-inla-download.org/R/stable"), dep=TRUE)

#install.packages("INLA", repos=c(getOption("repos"), INLA="https://inla.r-inla-download.org/R/testing"), dep=TRUE)
```

```{r}
# check 
packageVersion("INLA")
```

```{r}
library(INLA)
# inla.setOption(inla.mode="experimental") # if crash add this on
INLA::inla.version()
```

```{r}
# make a spatial model variables to be clear
spatial_model_variables <- model_variables %>%
  dplyr::select(
    lsoa21cd,
    # dependent
    access_score_15_pt,
    access_score_60_pt,
    access_score_15_walk,
    access_score_60_walk,
    access_score_15_cycle,
    access_score_60_cycle,
    # independent
    imd_z,
    suhii_z,
    log_pop_density_z,
    age_4_under_z,
    age_65_over_z,
    ethnic_minority_z,
    inner_outer_london,
    geometry
  ) %>%
  arrange(lsoa21cd)
```

```{r}
# check if it is sf objects
class(spatial_model_variables)
```

### 8.1. Spatial Neighbourhood structure

```{r}
# from here onwards follow code from: 
# https://www.r-inla.org/examples/AlvinBong/alvinbong-example.html, a tutorial that uses count-based areal data
# and also looked at 
# https://www.sciencedirect.com/science/article/pii/S1353829225001765?via%3Dihub#sec2 --> https://github.com/mbcalles/Area-Level-Deprivation-Traffic-Injury

# library(INLA)

# 1. Create a neighborhood list based on polygon contiguity - queens
nb <- spdep::poly2nb(
  spatial_model_variables,
  queen = TRUE
)

# 2. Convert the neighbor list to INLA adjacency format
spdep::nb2INLA("map.graph", nb)

# 3. Set the location of the adjacency file for INLA
map.adj <- file.path(getwd(), "map.graph")

# 4. Load adjacency matrix as a graph for INLA spatial models
g <- inla.read.graph(filename = map.adj)
```

### 8.2. Inla call/ fit - Generalised Linear Models (Poisson)

```{r}
# index for spatial effects
spatial_model_variables$area_id <- 1:nrow(spatial_model_variables)
```

```{r}
# specify model formula
formula_15_pt <- access_score_15_pt ~
  imd_z * suhii_z +
  log_pop_density_z +
  age_4_under_z +
  age_65_over_z +
  ethnic_minority_z +
  factor(inner_outer_london) +
  f(
    area_id,
    model = "bym2",
    graph = g,
    constr = TRUE,
    scale.model = TRUE
  )


formula_60_pt <- access_score_60_pt ~
  imd_z * suhii_z +
  log_pop_density_z +
  age_4_under_z +
  age_65_over_z +
  ethnic_minority_z +
  factor(inner_outer_london) +
  f(
    area_id,
    model = "bym2",
    graph = g,
    constr = TRUE,
    scale.model = TRUE
  )


formula_15_walk <- access_score_15_walk ~
  imd_z * suhii_z +
  log_pop_density_z +
  age_4_under_z +
  age_65_over_z +
  ethnic_minority_z +
  factor(inner_outer_london) +
  f(
    area_id,
    model = "bym2",
    graph = g,
    constr = TRUE,
    scale.model = TRUE
  )


formula_60_walk <- access_score_60_walk ~
  imd_z * suhii_z +
  log_pop_density_z +
  age_4_under_z +
  age_65_over_z +
  ethnic_minority_z +
  factor(inner_outer_london) +
  f(
    area_id,
    model = "bym2",
    graph = g,
    constr = TRUE,
    scale.model = TRUE
  )


formula_15_cycle <- access_score_15_cycle ~
  imd_z * suhii_z +
  log_pop_density_z +
  age_4_under_z +
  age_65_over_z +
  ethnic_minority_z +
  factor(inner_outer_london) +
  f(
    area_id,
    model = "bym2",
    graph = g,
    constr = TRUE,
    scale.model = TRUE
  )


formula_60_cycle <- access_score_60_cycle ~
  imd_z * suhii_z +
  log_pop_density_z +
  age_4_under_z +
  age_65_over_z +
  ethnic_minority_z +
  factor(inner_outer_london) +
  f(
    area_id,
    model = "bym2",
    graph = g,
    constr = TRUE,
    scale.model = TRUE
  )
```

**15 pt**

```{r}
# run model
model_15_pt_bym2 <- INLA::inla(
  formula = formula_15_pt,
  family = "poisson",
  data = spatial_model_variables,
  control.predictor = list(compute = TRUE),
  # control.compute = list(return.marginals.predictor = TRUE), 
  # https://becarioprecario.bitbucket.io/inla-gitbook/ch-spatial.html#sec:lattice 
  control.compute = list(dic = TRUE, waic = TRUE, cpo = TRUE, config = TRUE, 
                         return.marginals.predictor=TRUE),
  verbose = FALSE
)

#summary(model_15_pt_bym2)
```

**60 pt**

```{r}
# run model
model_60_pt_bym2 <- INLA::inla(
  formula = formula_60_pt,
  family = "poisson",
  data = spatial_model_variables,
  control.predictor = list(compute = TRUE),
  #control.compute = list(return.marginals.predictor = TRUE), 
  control.compute = list(dic = TRUE, waic = TRUE, cpo = TRUE, config = TRUE, 
                         return.marginals.predictor=TRUE),
  verbose = FALSE
)

#summary(model_60_pt_bym2)
```

**15 walk**

```{r}
# run model
model_15_walk_bym2 <- INLA::inla(
  formula = formula_15_walk,
  family = "poisson",
  data = spatial_model_variables,
  control.predictor = list(compute = TRUE),
  #control.compute = list(return.marginals.predictor = TRUE), 
  control.compute = list(dic = TRUE, waic = TRUE, cpo = TRUE, config = TRUE, 
                         return.marginals.predictor=TRUE), 
  verbose = FALSE
)

#summary(model_15_walk_bym2)
```

**60 walk**

```{r}
# run model
model_60_walk_bym2 <- INLA::inla(
  formula = formula_60_walk,
  family = "poisson",
  data = spatial_model_variables, 
  control.predictor = list(compute = TRUE),
  #control.compute = list(return.marginals.predictor = TRUE), 
  control.compute = list(dic = TRUE, waic = TRUE, cpo = TRUE, config = TRUE, 
                         return.marginals.predictor=TRUE),
  verbose = FALSE
)

#summary(model_60_walk_bym2)
```

**15 cycle**

```{r}
# run model
model_15_cycle_bym2 <- INLA::inla(
  formula = formula_15_cycle,
  family = "nbinomial",
  data = spatial_model_variables, 
  control.predictor = list(compute = TRUE),
  #control.compute = list(return.marginals.predictor = TRUE), 
  control.compute = list(dic = TRUE, waic = TRUE, cpo = TRUE, config = TRUE, 
                         return.marginals.predictor=TRUE),
  verbose = FALSE
)

summary(model_15_cycle_bym2)
```

**60 cycle**

```{r}
# run model
model_60_cycle_bym2 <- INLA::inla(
  formula = formula_60_cycle,
  family = "poisson",
  data = spatial_model_variables,
  control.predictor = list(compute = TRUE),
  #control.compute = list(return.marginals.predictor = TRUE), 
  control.compute = list(dic = TRUE, waic = TRUE, cpo = TRUE, config = TRUE, 
                         return.marginals.predictor=TRUE), 
  verbose = FALSE
)

#summary(model_60_cycle_bym2)
```

```{r}
# 1. Named list with self-explanatory labels
bym2_models <- list(
  "Model 1" = model_15_pt_bym2,
  "Model 2" = model_60_pt_bym2,
  "Model 3" = model_15_walk_bym2,
  "Model 4" = model_60_walk_bym2,
  "Model 5" = model_15_cycle_bym2,
  "Model 6" = model_60_cycle_bym2
)

# 2. Readable variable names (order here = row order in the table)
variable_labels <- c(
  "(Intercept)" = "Intercept",
  "imd_z" = "IMD (z)",
  "suhii_z" = "SUHII (z)",
  "imd_z:suhii_z" = "IMD × SUHII",
  "log_pop_density_z" = "Log pop. density (z)",
  "age_4_under_z" = "Age 4 or under % (z)",                 
  "age_65_over_z" = "Age 65 or over % (z)",
  "ethnic_minority_z" = "Ethnic minority % (z)",
  "factor(inner_outer_london)Inner" = "Inner, Outer London"
)

# 3. Extract fixed effects, exponentiate to rate ratios, format as one cell
# just like this: https://www.sciencedirect.com/science/article/pii/S2405673123000144?utm_source=chatgpt.com
fixed_effects_long <- lapply(names(bym2_models), function(model_name) {
  bym2_models[[model_name]]$summary.fixed %>%
    as.data.frame() %>%
    rownames_to_column("term") %>%
    transmute(
      model = model_name,
      variable = variable_labels[term],
      cell = sprintf(
        "%.3f (%.3f, %.3f)",
        exp(`0.5quant`), exp(`0.025quant`), exp(`0.975quant`)
      )
    )
}) %>% bind_rows()

# 4. One column per model, one row per variable
fixed_effects_table <- fixed_effects_long %>%
  pivot_wider(names_from = model, values_from = cell) %>%
  arrange(match(variable, variable_labels))

knitr::kable(
  fixed_effects_table,
  caption = "BYM2 fixed effects: Rate Ratios and 95% Credible Intervals (CI)"
)
```

### 8.3. Validation Checks

#### 8.3.1. Predictive Diagnostic Methods - CPO and PIT for all models

Predictive measures can be used both to validate and to compare models and using marginal... DIC and CPO and PIT (Referencing from here: <https://www.math.ntnu.no/preprint/statistics/2008/S2-2008.pdf>)

```{r}
# List the current six models
six_models <- list(
  "15_pt"    = model_15_pt_bym2,
  "60_pt"    = model_60_pt_bym2,
  "15_walk"  = model_15_walk_bym2,
  "60_walk"  = model_60_walk_bym2,
  "15_cycle" = model_15_cycle_bym2,
  "60_cycle" = model_60_cycle_bym2
)

# CPO/PIT summary table 
cpo_pit_summary <- lapply(names(six_models), function(model_name) {
  m <- six_models[[model_name]]
  data.frame(
    model = model_name,
    # model-level predictive comparison score - CPO
    sum_log_cpo = sum(log(m$cpo$cpo), na.rm = TRUE),
    # PIT
    mean_pit = mean(m$cpo$pit, na.rm = TRUE)
  )
}) %>% bind_rows()

cpo_pit_summary

# PIT histograms
par(mfrow = c(2, 3), mar = c(3, 3, 2, 1))
for (nm in names(six_models)) {
  hist(six_models[[nm]]$cpo$pit, breaks = 20,
       main = paste0(nm, " (Poisson)"), xlab = "", ylab = "", col = "steelblue")
}
par(mfrow = c(1, 1))
```

```{r}
# check if grraph nodes matches LSOAs

# Number of graph nodes
cat("Number of graph nodes:", g$n, "\n")
cat("Number of LSOAs:", nrow(spatial_model_variables), "\n")

cat(
  "Graph and data sizes match:",
  g$n == nrow(spatial_model_variables),
  "\n"
)

cat(
  "area_id is correctly ordered:",
  identical(
    spatial_model_variables$area_id,
    seq_len(nrow(spatial_model_variables))
  ),
  "\n"
)

cat(
  "Number of duplicated LSOA codes:",
  anyDuplicated(spatial_model_variables$lsoa21cd),
  "\n"
)
```

#### 8.3.2. CPO failures

```{r}
bym2_models <- list(
  "15-minute Public Transit" = model_15_pt_bym2,
  "60-minute Public Transit" = model_60_pt_bym2,
  "15-minute Walking" = model_15_walk_bym2,
  "60-minute Walking" = model_60_walk_bym2,
  "15-minute Cycling" = model_15_cycle_bym2,
  "60-minute Cycling" = model_60_cycle_bym2
)

cpo_failure_counts <- sapply(
  bym2_models,
  function(model) sum(model$cpo$failure > 0, na.rm = TRUE)
)

cpo_failure_counts
```

```{r, eval=FALSE}
# recompute
model_15_pt_bym2_cpo <- INLA::inla.cpo(
  model_15_pt_bym2
)

model_15_walk_bym2_cpo <- INLA::inla.cpo(
  model_15_walk_bym2
)

model_15_cycle_bym2_cpo <- INLA::inla.cpo(
  model_15_cycle_bym2
)

sum(log(model_15_pt_bym2_cpo$cpo$cpo), na.rm = TRUE)
sum(log(model_15_walk_bym2_cpo$cpo$cpo), na.rm = TRUE)
sum(log(model_15_cycle_bym2_cpo$cpo$cpo), na.rm = TRUE)
```

#### 8.3.3. Residuals

Rebuild the centroids, queens, knn first

```{r}
# centroids 
coordsLSOA2 <- spatial_model_variables %>% 
  st_centroid() %>% 
  st_geometry()

plot(coordsLSOA2)
```

```{r}
# queens
LLSOA_nb2 <- spatial_model_variables %>%
  poly2nb(., queen = TRUE)

#or nearest neighbours
knn_LSOA2 <-coordsLSOA2 %>%
  knearneigh(., k=4)

LLSOA_knn2 <- knn_LSOA2 %>%
  knn2nb()
```

```{r}
#plot them
plot(LLSOA_nb2, st_geometry(coordsLSOA2), col="red")
```

```{r}
plot(LLSOA_knn, st_geometry(coordsLSOA2), col="blue")
```

```{r}
#create a spatial weights matrix object from these weights

LLSOA.queens_weight2 <- LLSOA_nb2 %>%
  nb2listw(., style="W")

LLSOA.knn_4_weight2 <- LLSOA_knn2 %>%
  nb2listw(., style="W")
```

**Pearson Residuals - i = (Y_i − λ̂\_i) / sqrt(λ̂\_i)**

**15 pt**

```{r}
model_15_pt_bym2_resids <- (
  # observed minus 
  spatial_model_variables$access_score_15_pt -
    # fitted value
    model_15_pt_bym2$summary.fitted.values$mean
  # divided by square root of fitted values
)/sqrt(model_15_pt_bym2$summary.fitted.values$mean)

model_profile <- spatial_model_variables %>%
  mutate(model_15_pt_bym2_resids)

#now plot the residuals
tmap_mode("view")
qtm(model_profile, fill = "model_15_pt_bym2_resids")
```

```{r}
# queens case
model_15_pt_bym2_resids_queen <- model_profile %>%
  st_drop_geometry()%>%
  dplyr::select(model_15_pt_bym2_resids)%>%
  pull()%>%
  moran.test(., LLSOA.queens_weight2)%>%
  tidy()

model_15_pt_bym2_resids_queen
```

```{r}
# nearest neighbour
model_15_pt_bym2_resids_nearest_neighbour <- model_profile %>%
  st_drop_geometry()%>%
  dplyr::select(model_15_pt_bym2_resids)%>%
  pull()%>%
  moran.test(., LLSOA.knn_4_weight2)%>%
  tidy()

model_15_pt_bym2_resids_nearest_neighbour
```

**60 pt**

```{r}
model_60_pt_bym2_resids <- (
  spatial_model_variables$access_score_60_pt -
    model_60_pt_bym2$summary.fitted.values$mean
)/sqrt(model_60_pt_bym2$summary.fitted.values$mean)

model_profile <- spatial_model_variables %>%
  mutate(model_60_pt_bym2_resids)

#now plot the residuals
tmap_mode("view")
qtm(model_profile, fill = "model_60_pt_bym2_resids")
```

```{r}
# queens case
model_60_pt_bym2_resids_queen <- model_profile %>%
  st_drop_geometry()%>%
  dplyr::select(model_60_pt_bym2_resids)%>%
  pull()%>%
  moran.test(., LLSOA.queens_weight2)%>%
  tidy()

model_60_pt_bym2_resids_queen
```

```{r}
# nearest neighbour
model_60_pt_bym2_resids_nearest_neighbour <- model_profile %>%
  st_drop_geometry()%>%
  dplyr::select(model_60_pt_bym2_resids)%>%
  pull()%>%
  moran.test(., LLSOA.knn_4_weight2)%>%
  tidy()

model_60_pt_bym2_resids_nearest_neighbour
```

**15 walk**

```{r}
model_15_walk_bym2_resids <- (
  spatial_model_variables$access_score_15_walk -
    model_15_walk_bym2$summary.fitted.values$mean
)/sqrt(model_15_walk_bym2$summary.fitted.values$mean)

model_profile <- spatial_model_variables %>%
  mutate(model_15_walk_bym2_resids)

#now plot the residuals
tmap_mode("view")
qtm(model_profile, fill = "model_15_walk_bym2_resids")
```

```{r}
# queens case
model_15_walk_bym2_resids_queen <- model_profile %>%
  st_drop_geometry()%>%
  dplyr::select(model_15_walk_bym2_resids)%>%
  pull()%>%
  moran.test(., LLSOA.queens_weight2)%>%
  tidy()

model_15_walk_bym2_resids_queen
```

```{r}
# nearest neighbour
model_15_walk_bym2_resids_nearest_neighbour <- model_profile %>%
  st_drop_geometry()%>%
  dplyr::select(model_15_walk_bym2_resids)%>%
  pull()%>%
  moran.test(., LLSOA.knn_4_weight2)%>%
  tidy()

model_15_walk_bym2_resids_nearest_neighbour
```

**60 walk**

```{r}
model_60_walk_bym2_resids <- (
  spatial_model_variables$access_score_60_walk -
    model_60_walk_bym2$summary.fitted.values$mean
)/sqrt(model_60_walk_bym2$summary.fitted.values$mean)

model_profile <- spatial_model_variables %>%
  mutate(model_60_walk_bym2_resids)

#now plot the residuals
tmap_mode("view")
qtm(model_profile, fill = "model_60_walk_bym2_resids")
```

```{r}
# queens case
model_60_walk_bym2_resids_queen <- model_profile %>%
  st_drop_geometry()%>%
  dplyr::select(model_60_walk_bym2_resids)%>%
  pull()%>%
  moran.test(., LLSOA.queens_weight2)%>%
  tidy()

model_60_walk_bym2_resids_queen
```

```{r}
# nearest neighbour
model_60_walk_bym2_resids_nearest_neighbour <- model_profile %>%
  st_drop_geometry()%>%
  dplyr::select(model_60_walk_bym2_resids)%>%
  pull()%>%
  moran.test(., LLSOA.knn_4_weight2)%>%
  tidy()

model_60_walk_bym2_resids_nearest_neighbour
```

**15 cycle**

```{r}
model_15_cycle_bym2_resids <- (
  spatial_model_variables$access_score_15_cycle -
    model_15_cycle_bym2$summary.fitted.values$mean
)/sqrt(model_15_cycle_bym2$summary.fitted.values$mean)

model_profile <- spatial_model_variables %>%
  mutate(model_15_cycle_bym2_resids)

#now plot the residuals
tmap_mode("view")
qtm(model_profile, fill = "model_15_cycle_bym2_resids")
```

```{r}
# queens case
model_15_cycle_bym2_resids_queen <- model_profile %>%
  st_drop_geometry()%>%
  dplyr::select(model_15_cycle_bym2_resids)%>%
  pull()%>%
  moran.test(., LLSOA.queens_weight2)%>%
  tidy()

model_15_cycle_bym2_resids_queen
```

```{r}
# nearest neighbour
model_15_cycle_bym2_resids_nearest_neighbour <- model_profile %>%
  st_drop_geometry()%>%
  dplyr::select(model_15_cycle_bym2_resids)%>%
  pull()%>%
  moran.test(., LLSOA.knn_4_weight2)%>%
  tidy()

model_15_cycle_bym2_resids_nearest_neighbour
```

**60 cycle**

```{r}
model_60_cycle_bym2_resids <- (
  spatial_model_variables$access_score_60_cycle -
    model_60_cycle_bym2$summary.fitted.values$mean
)/sqrt(model_60_cycle_bym2$summary.fitted.values$mean)

model_profile <- spatial_model_variables %>%
  mutate(model_60_cycle_bym2_resids)

#now plot the residuals
tmap_mode("view")
qtm(model_profile, fill = "model_60_cycle_bym2_resids")
```

```{r}
# queens case
model_60_cycle_bym2_resids_queen <- model_profile %>%
  st_drop_geometry()%>%
  dplyr::select(model_60_cycle_bym2_resids)%>%
  pull()%>%
  moran.test(., LLSOA.queens_weight2)%>%
  tidy()

model_60_cycle_bym2_resids_queen
```

```{r}
# nearest neighbour
model_60_cycle_bym2_resids_nearest_neighbour <- model_profile %>%
  st_drop_geometry()%>%
  dplyr::select(model_60_cycle_bym2_resids)%>%
  pull()%>%
  moran.test(., LLSOA.knn_4_weight2)%>%
  tidy()

model_60_cycle_bym2_resids_nearest_neighbour
```

## 9. BiLISA - Bivariate LISA mapping

Well known method to address issues of spatial nonstationarity, allowing geographic variations in the relationships between dependent and explanatory variables –\> use `sepdep` package

Each of the six dependent variables will be mapped with IMD and SUHII respectively:

```{r}
bilisa_variables <- spatial_model_variables %>%
  arrange(lsoa21cd)

# Create Queen-contiguity neighbours
bilisa_nb <- bilisa_variables %>%
  poly2nb(.,queen = T)

bilisa_weights <-  bilisa_nb %>%
  nb2listw(.,style = "W",)
```

### 9.1. Map SUHII

```{r}
# https://r-spatial.github.io/spdep/reference/localmoran_bv.html

bilisa_15_pt_suhii <- localmoran_bv(
  x = bilisa_variables$access_score_15_pt,
  y = bilisa_variables$suhii_z,
  listw = bilisa_weights,
  nsim = 199,
  scale = TRUE,
  alternative = "two.sided",
  iseed = 42,
  zero.policy = TRUE
)

head(bilisa_15_pt_suhii)
```

```{r}
bilisa_cluster <- spdep::hotspot(
  bilisa_15_pt_suhii,
  Prname = "Pr(folded) Sim",
  cutoff = 0.05,
  quadrant.type = "pysal",
  p.adjust = "none"
)
```

```{r}
bilisa_variables$bilisa_cluster <- as.character(
  bilisa_cluster
)

bilisa_variables$bilisa_cluster[
  is.na(bilisa_variables$bilisa_cluster)
] <- "Not significant"
```

```{r}
table(bilisa_variables$bilisa_cluster)
```

```{r}
tmap_mode("plot")

tm_shape(bilisa_variables) +
  tm_polygons(
    fill = "bilisa_cluster",
    fill.scale = tm_scale_categorical(),
    fill.legend = tm_legend(
      title = "BiLISA Cluster"
    ),
    col = NA
  ) +
  tm_title(
    "15-minute Public Transit Accessibility and SUHII"
  )
```

```{r}
# reorder clusters and make better labels
bilisa_variables <- bilisa_variables %>%
  mutate(
    bilisa_cluster_plot = case_when(
      bilisa_cluster == "Low-High" ~ "Low access–High SUHII",
      bilisa_cluster == "High-High" ~ "High access–High SUHII",
      bilisa_cluster == "Low-Low" ~ "Low access–Low SUHII",
      bilisa_cluster == "High-Low" ~ "High access–Low SUHII",
      TRUE ~ NA_character_
    )
  )

# set legend order
bilisa_variables$bilisa_cluster_plot <- factor(
  bilisa_variables$bilisa_cluster_plot,
  levels = c(
    "Low access–High SUHII",
    "High access–High SUHII",
    "Low access–Low SUHII",
    "High access–Low SUHII"
  )
)


cluster_cols <- c(
  "Low access–High SUHII"  = "#D73027",  # red
  "High access–High SUHII" = "#F4A582",  # light red / salmon
  "Low access–Low SUHII"   = "#91BFDB",  # light blue
  "High access–Low SUHII"  = "#4575B4"   # blue
)
```

```{r}
tmap_mode("plot")

bilisa_15_pt_suhii_suhii_map <-
  # Base: all LSOAs white
  tm_shape(bilisa_variables) +
  tm_fill(
    fill = "white"
  ) +

  # Significant clusters
  tm_shape(
    dplyr::filter(
      bilisa_variables,
      !is.na(bilisa_cluster_plot)
    )
  ) +
  tm_fill(
    fill = "bilisa_cluster_plot",
    fill.scale = tm_scale_categorical(
      values = cluster_cols
    ),
    fill.legend = tm_legend(
      title = "BiLISA cluster",
      position = tm_pos_out("right", "center"),
      frame = FALSE
    )
  ) +

  # London boundary must come last
  tm_shape(london_boundary) +
  tm_borders(
    col = "grey30",
    lwd = 1
  ) +

  tm_title(
    "15-minute public transport accessibility\nand neighbouring SUHII"
  ) +

  tm_layout(
    frame = FALSE,
    bg.color = "white",
    outer.margins = 0.02
  )

bilisa_15_pt_suhii_suhii_map 
```

**Loop over maps:**

```{r}
run_bilisa_map <- function(
    data,
    access_var,
    context_var,
    context_label,
    map_title,
    listw = bilisa_weights,
    boundary = london_boundary,
    nsim = 199,
    seed = 42,
    show_legend = TRUE
) {

  # Standardise only the accessibility variable
  access_z <- as.numeric(
    scale(data[[access_var]])
  )

  # SUHII or IMD is already a z-score
  context_z <- data[[context_var]]

  # Check that both variables and weights have matching lengths
  stopifnot(
    length(access_z) == length(context_z),
    length(access_z) == length(listw$neighbours)
  )

  # Bivariate Local Moran's I
  lisa_result <- spdep::localmoran_bv(
    x = access_z,
    y = context_z,
    listw = listw,
    nsim = nsim,
    scale = FALSE,
    alternative = "two.sided",
    iseed = seed,
    zero.policy = TRUE
  )

  # Significant cluster classification
  lisa_cluster <- spdep::hotspot(
    lisa_result,
    Prname = "Pr(folded) Sim",
    cutoff = 0.05,
    quadrant.type = "pysal",
    p.adjust = "none"
  )

  # Full, readable cluster labels
  cluster_labels <- c(
    "Low-High" =
      paste0("Low access–High ", context_label),

    "High-High" =
      paste0("High access–High ", context_label),

    "Low-Low" =
      paste0("Low access–Low ", context_label),

    "High-Low" =
      paste0("High access–Low ", context_label)
  )

  cluster_levels <- unname(
    cluster_labels[
      c(
        "Low-High",
        "High-High",
        "Low-Low",
        "High-Low"
      )
    ]
  )

  cluster_colours <- stats::setNames(
    c(
      "#D73027",  # Low-High
      "#F4A582",  # High-High
      "#91BFDB",  # Low-Low
      "#4575B4"   # High-Low
    ),
    cluster_levels
  )

  # Create a separate result dataset
  result_data <- data %>%
    dplyr::mutate(
      bilisa_i = as.numeric(lisa_result[, 1]),

      bilisa_p = as.numeric(
        lisa_result[, "Pr(folded) Sim"]
      ),

      bilisa_cluster = as.character(
        lisa_cluster
      ),

      bilisa_cluster_plot = unname(
        cluster_labels[bilisa_cluster]
      ),

      bilisa_cluster_plot = factor(
        bilisa_cluster_plot,
        levels = cluster_levels
      )
    )

  # Significant areas only
  significant_data <- result_data %>%
    dplyr::filter(
      !is.na(bilisa_cluster_plot)
    )

  # Show or hide the legend
  legend_setting <- if (show_legend) {
    tm_legend(
      title = "BiLISA cluster",
      position = tm_pos_out("right", "center"),
      frame = FALSE
    )
  } else {
    tm_legend_hide()
  }

  # Create the map
  map_object <-
    tm_shape(result_data) +
    tm_fill(
      fill = "white"
    ) +

    tm_shape(significant_data) +
    tm_fill(
      fill = "bilisa_cluster_plot",

      fill.scale = tm_scale_categorical(
        values = cluster_colours
      ),

      fill.legend = legend_setting
    ) +

    # London boundary last
    tm_shape(boundary) +
    tm_borders(
      col = "grey30",
      lwd = 1
    ) +

    tm_title(
      map_title
    ) +

    tm_layout(
      frame = FALSE,
      bg.color = "white",
      outer.margins = 0.02
    )

  # Return everything
  list(
    result = lisa_result,
    clusters = lisa_cluster,
    data = result_data,
    map = map_object
  )
}
```

```{r}
bilisa_15_pt_suhii <- run_bilisa_map(
  data = bilisa_variables,
  access_var = "access_score_15_pt",
  context_var = "suhii_z",
  context_label = "SUHII",
  map_title =
    "15-minute Public Transit Accessibility and SUHII",
  show_legend = FALSE
)

# assign as object
bilisa_15_pt_suhii_map <- bilisa_15_pt_suhii$map

# plot
bilisa_15_pt_suhii_map
```

```{r}
bilisa_60_pt_suhii <- run_bilisa_map(
  data = bilisa_variables,
  access_var = "access_score_60_pt",
  context_var = "suhii_z",
  context_label = "SUHII",
  map_title =
    "60-minute Public Transit Accessibility and SUHII",
  show_legend = FALSE
)

# assign as object
bilisa_60_pt_suhii_map <- bilisa_60_pt_suhii$map

# plot
bilisa_60_pt_suhii_map
```

```{r}
bilisa_15_walk_suhii <- run_bilisa_map(
  data = bilisa_variables,
  access_var = "access_score_15_walk",
  context_var = "suhii_z",
  context_label = "SUHII",
  map_title =
    "15-minute Walking Accessibility and SUHII",
  show_legend = FALSE
)

# assign as object
bilisa_15_walk_suhii_map <- bilisa_15_walk_suhii$map

# plot
bilisa_15_walk_suhii_map
```

```{r}
bilisa_60_walk_suhii <- run_bilisa_map(
  data = bilisa_variables,
  access_var = "access_score_60_walk",
  context_var = "suhii_z",
  context_label = "SUHII",
  map_title =
    "60-minute Walking Accessibility and SUHII",
  show_legend = FALSE
)

# assign as object
bilisa_60_walk_suhii_map <- bilisa_60_walk_suhii$map

# plot
bilisa_60_walk_suhii_map
```

```{r}
bilisa_15_cycle_suhii <- run_bilisa_map(
  data = bilisa_variables,
  access_var = "access_score_15_cycle",
  context_var = "suhii_z",
  context_label = "SUHII",
  map_title =
    "15-minute Cycling Accessibility and SUHII",
  show_legend = FALSE
)

# assign as object
bilisa_15_cycle_suhii_map <- bilisa_15_cycle_suhii$map

# plot
bilisa_15_cycle_suhii_map
```

```{r}
bilisa_60_cycle_suhii <- run_bilisa_map(
  data = bilisa_variables,
  access_var = "access_score_60_cycle",
  context_var = "suhii_z",
  context_label = "SUHII",
  map_title =
    "60-minute Cycling Accessibility and SUHII",
  show_legend = FALSE
)

# assign as object
bilisa_60_cycle_suhii_map <- bilisa_60_cycle_suhii$map

# plot
bilisa_60_cycle_suhii_map
```

```{r}
bilisa_15_pt_imd <- run_bilisa_map(
  data = bilisa_variables,
  access_var = "access_score_15_pt",
  context_var = "imd_z",
  context_label = "IMD",
  map_title =
    "15-minute Public Transit Accessibility and IMD",
  show_legend = FALSE
)

# assign as object
bilisa_15_pt_imd_map <- bilisa_15_pt_imd$map

# plot
bilisa_15_pt_imd_map
```

```{r}
bilisa_60_pt_imd <- run_bilisa_map(
  data = bilisa_variables,
  access_var = "access_score_60_pt",
  context_var = "imd_z",
  context_label = "IMD",
  map_title =
    "60-minute Public Transit Accessibility and IMD",
  show_legend = FALSE
)

# assign as object
bilisa_60_pt_imd_map <- bilisa_60_pt_imd$map

# plot
bilisa_60_pt_imd_map
```

```{r}
bilisa_15_walk_imd <- run_bilisa_map(
  data = bilisa_variables,
  access_var = "access_score_15_walk",
  context_var = "imd_z",
  context_label = "IMD",
  map_title = "15-minute Walking Accessibility and IMD",
  show_legend = FALSE
)

# assign as object
bilisa_15_walk_imd_map <- bilisa_15_walk_imd$map

# plot
bilisa_15_walk_imd_map
```

```{r}
bilisa_60_walk_imd <- run_bilisa_map(
  data = bilisa_variables,
  access_var = "access_score_60_walk",
  context_var = "imd_z",
  context_label = "IMD",
  map_title =
    "60-minute Walking Accessibility and IMD",
  show_legend = FALSE
)

# assign as object
bilisa_60_walk_imd_map <- bilisa_60_walk_imd$map

# plot
bilisa_60_walk_imd_map
```

```{r}
bilisa_15_cycle_imd <- run_bilisa_map(
  data = bilisa_variables,
  access_var = "access_score_15_cycle",
  context_var = "imd_z",
  context_label = "IMD",
  map_title =
    "15-minute Cycle Accessibility and IMD",
  show_legend = FALSE
)

# assign as object
bilisa_15_cycle_imd_map <- bilisa_15_cycle_imd$map

# plot
bilisa_15_cycle_imd_map
```

```{r}
bilisa_60_cycle_imd <- run_bilisa_map(
  data = bilisa_variables,
  access_var = "access_score_60_cycle",
  context_var = "imd_z",
  context_label = "IMD",
  map_title =
    "60-minute Cycle Accessibility and IMD",
  show_legend = FALSE
)

# assign as object
bilisa_60_cycle_imd_map <- bilisa_60_cycle_imd$map

# plot
bilisa_60_cycle_imd_map
```

**Cowplot Combine Maps**

```{r}
bilisa_grobs <- list(
  tmap::tmap_grob(bilisa_15_pt_suhii_map),
  tmap::tmap_grob(bilisa_15_pt_imd_map),

  tmap::tmap_grob(bilisa_15_walk_suhii_map),
  tmap::tmap_grob(bilisa_15_walk_imd_map),

  tmap::tmap_grob(bilisa_15_cycle_suhii_map),
  tmap::tmap_grob(bilisa_15_cycle_imd_map)
)

bilisa_15_maps_combined <- cowplot::plot_grid(
  plotlist = bilisa_grobs,
  ncol = 2,
  labels = c("a", "b", "c", "d", "e", "f"),
  label_size = 12,
  align = "hv"
)

bilisa_15_maps_combined
```

```{r}
# set up a shared legend for plotting purposes

legend_labels <- c(
  "Low access–High SUHII/IMD",
  "High access–High SUHII/IMD",
  "Low access–Low SUHII/IMD",
  "High access–Low SUHII/IMD"
)

legend_colours <- c(
  "Low access–High SUHII/IMD"  = "#D73027",
  "High access–High SUHII/IMD" = "#F4A582",
  "Low access–Low SUHII/IMD"   = "#91BFDB",
  "High access–Low SUHII/IMD"  = "#4575B4"
)

legend_data <- data.frame(
  cluster = factor(
    legend_labels,
    levels = legend_labels
  ),
  x = 1,
  y = 1:4
)

legend_plot <- ggplot(
  legend_data,
  aes(x = x, y = y, fill = cluster)
) +
  geom_point(
    shape = 22,
    size = 8,
    colour = NA
  ) +
  scale_fill_manual(
    values = legend_colours,
    breaks = legend_labels,
    drop = FALSE,
    name = "BiLISA cluster"
  ) +
  guides(
    fill = guide_legend(
      override.aes = list(
        shape = 22,
        size = 8,
        colour = NA
      )
    )
  ) +
  theme_void() +
  theme(
    legend.position = "right",
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 10),
    legend.key.height = grid::unit(0.65, "cm"),
    legend.key.width = grid::unit(0.65, "cm"),
    legend.background = element_blank()
  )

shared_legend <- cowplot::get_legend(
  legend_plot,
  legend = "right"
)
```

```{r}
# 15 min BiLISA maps
bilisa_15_final <- cowplot::plot_grid(
  bilisa_15_maps_combined,
  shared_legend,
  ncol = 2,
  rel_widths = c(0.78, 0.22)
)

bilisa_15_final
```

```{r}
# save as png
ggsave(
  filename = "figures/bilisa_15_maps_combined.png",
  plot = bilisa_15_final,
  width = 12.2,
  height = 8.27,
  units = "in",
  dpi = 300,
  bg = "white"
)
```

```{r}
# 60 min BiLISA maps
bilisa_grobs2 <- list(
  tmap::tmap_grob(bilisa_60_pt_suhii_map),
  tmap::tmap_grob(bilisa_60_pt_imd_map),

  tmap::tmap_grob(bilisa_60_walk_suhii_map),
  tmap::tmap_grob(bilisa_60_walk_imd_map),

  tmap::tmap_grob(bilisa_60_cycle_suhii_map),
  tmap::tmap_grob(bilisa_60_cycle_imd_map)
)

bilisa_60_maps_combined <- cowplot::plot_grid(
  plotlist = bilisa_grobs2,
  ncol = 2,
  labels = c("a", "b", "c", "d", "e", "f"),
  label_size = 12,
  align = "hv"
)

bilisa_60_maps_combined
```

```{r}
bilisa_60_final <- cowplot::plot_grid(
  bilisa_60_maps_combined,
  shared_legend,
  ncol = 2,
  rel_widths = c(0.78, 0.22)
)

bilisa_60_final
```

```{r}
ggsave(
  filename = "figures/bilisa_60_maps_combined.png",
  plot = bilisa_60_final,
  width = 12.2,
  height = 8.27,
  units = "in",
  dpi = 300,
  bg = "white"
)
```
