---
title: "6_Equity_Analysis"
---

```{r}
# remove objects from R environment and clear memory
rm(list = ls())    
gc()
```

## Load Packages

```{r}
library(sf)
library(sp)
library(readr)
library(dplyr)
library(knitr)
library(tidyverse)
library(tmap)
library(cowplot)
library(plotly)
library(corrplot)
library(broom)
library(mapview)
library(stringr)
library(spdep)
library(car)
library(fs)
library(janitor)
library(GGally)
library(MASS)
library(spatialreg)
library(DHARMa)
```

## Data Loading

Clean names, select initial needed columns and rename columns for consistency across datasets

**London only LSOAs**

```{r}
london_lsoa <- st_read("data/lsoa_london/lsoa_london.shp")
```

**Data obtained from previous pre-processing**

```{r}
# travel time 
multimodal_accessibility_lsoa <- st_read("data/multimodal_accessibility_lsoa.gpkg")%>%
  clean_names()%>%
  dplyr::select(
    lsoa21cd,
    lsoa21nm,
    access_score_15_pt,
    access_score_60_pt,
    access_score_15_walk,
    access_score_60_walk,
    access_score_15_cycle,
    access_score_60_cycle,
    travel_time_min_pt,
    travel_time_min_walk,
    travel_time_min_cycle,
    geom
  )%>%
  rename(
    geometry = geom
  )
```

```{r}
# SUHII as lsoa heat exposure
suhii_lsoa <- st_read("data/London_LSOA_LST_SUHI/London_LSOA_LST_SUHI.shp")%>% 
  st_transform(.,27700)%>%
  clean_names()%>%
  dplyr::select(
    lsoa21cd,
    lsoa21nm,
    mean,
    london_mea,
    lst_differ,
    geometry
  )%>%
  rename(
    mean_lst_lsoa = mean,
    london_mean = london_mea,
    suhii = lst_differ
  )
```

**IMD 2025 data**

```{r}
# IMD score
imd_score <- read_csv("data/IMD_2025.csv")%>%
  clean_names()%>%
  dplyr::select(
    lsoa_code_2021,
    lsoa_name_2021,
    index_of_multiple_deprivation_imd_score
    )%>%
  rename(
    lsoa21cd = lsoa_code_2021,
    lsoa21nm = lsoa_name_2021,
    imd_score = index_of_multiple_deprivation_imd_score
  )
```

**Census Data**

```{r}
# TS006 population density
census_pop_density <- read_csv("data/census2021/census2021-ts006-lsoa.csv")%>%
  clean_names()%>%
  dplyr::select(
    geography,
    geography_code,
    population_density_persons_per_square_kilometre_measures_value
  )%>%
  rename(
    lsoa21cd = geography_code,
    lsoa21nm = geography,
    pop_density = population_density_persons_per_square_kilometre_measures_value
  )
```

```{r}
# TS007a age by 5 year band
census_age_5year_band <- read_csv("data/census2021/census2021-ts007a-lsoa.csv")%>%
  clean_names()%>%
  dplyr::select(
    geography,
    geography_code,
    age_total,
    age_aged_4_years_and_under,
    # for filter for age 65 or over 
    age_aged_65_to_69_years,
    age_aged_70_to_74_years,
    age_aged_75_to_79_years,
    age_aged_80_to_84_years,
    age_aged_85_years_and_over
  )%>%
  rename(
    lsoa21cd = geography_code,
    lsoa21nm = geography,
    age_4_and_under = age_aged_4_years_and_under,
    age_65_to_69 = age_aged_65_to_69_years,
    age_70_to_74 = age_aged_70_to_74_years,
    age_75_to_79 = age_aged_75_to_79_years,
    age_80_to_84 = age_aged_80_to_84_years,
    age_85_and_over = age_aged_85_years_and_over
  )
```

```{r, eval=FALSE}
# TS038 disability
census_disability <- read_csv("data/census2021/census2021-ts038-lsoa.csv")%>%
  clean_names()%>%
  select(
    geography,
    geography_code,
    disability_total_all_usual_residents,
    disability_disabled_under_the_equality_act
  )%>%
  rename(
    lsoa21cd = geography_code,
    lsoa21nm = geography,
    # total number of residents 
    disability_total_residents = disability_total_all_usual_residents,
    # disabled under the Equality Act 2010
    disabled = disability_disabled_under_the_equality_act
  )
```

```{r, eval=FALSE}
# TS037 general health
census_health <- read_csv("data/census2021/census2021-ts037-lsoa.csv")%>% 
  clean_names()%>%
  select(
    geography,
    geography_code,
    general_health_total_all_usual_residents,
    general_health_bad_health,
    general_health_very_bad_health
  )%>%
  rename(
    lsoa21cd = geography_code,
    lsoa21nm = geography,
    general_health_total_residents = general_health_total_all_usual_residents,
    general_health_bad = general_health_bad_health,
    general_health_very_bad = general_health_very_bad_health
  )
```

```{r}
# TS021 Ethnicity
# definition of ethnic minority groups
# https://www.ethnicity-facts-figures.service.gov.uk/style-guide/writing-about-ethnicity/ 
# https://service-manual.ons.gov.uk/content/language/ethnicity-and-race
census_ethnicity <- read_csv("data/census2021/census2021-ts021-lsoa.csv")%>%
  clean_names()%>%
  dplyr::select(
    geography,
    geography_code,
    ethnic_group_total_all_usual_residents,
    # Use “Ethnic minorities” to refer to all ethnic groups except the White British group
    ethnic_group_white_english_welsh_scottish_northern_irish_or_british
  )%>%
  rename(
    lsoa21cd = geography_code,
    lsoa21nm = geography,
    ethnic_group_total_residents = ethnic_group_total_all_usual_residents,
    white_british_group = ethnic_group_white_english_welsh_scottish_northern_irish_or_british
  )
```

## 1. Data Cleaning

### Check Loaded + Initial Cleaned Data Column Names

```{r}
names(london_lsoa)
```

```{r}
# check 3 modes 
names(multimodal_accessibility_lsoa)
```

```{r}
# check names in SUHII export
names(suhii_lsoa)
```

```{r}
# check IMD names
names(imd_score)
```

```{r}
# check census data
names(census_pop_density)
```

```{r}
names(census_age_5year_band)
```

```{r}
names(census_ethnicity)
```

### Wrangle data variables for regression modelling

```{r}
# filter london lsoas only
london_lsoa_codes <- london_lsoa %>%
  clean_names() %>%
  pull(lsoa21cd)
```

```{r}
census_age_cleaned <- census_age_5year_band %>%
  mutate(
    age_4_and_under_pct = ifelse(
      age_total > 0,
      (age_4_and_under / age_total) * 100,
      NA
    ),
    age_65_and_over = rowSums(
      across(
        c(
          age_65_to_69,
          age_70_to_74,
          age_75_to_79,
          age_80_to_84,
          age_85_and_over
        )
      ),
      na.rm = TRUE
    ),
    age_65_and_over_pct = ifelse(
      age_total > 0,
      (age_65_and_over / age_total)* 100,
      NA
    )
  )%>% 
  filter(lsoa21cd %in% london_lsoa_codes)
```

```{r, eval=FALSE}
census_disability_cleaned <- census_disability %>%
  mutate(
    disability_pct = ifelse(
      disability_total_residents > 0,
      (disabled / disability_total_residents)* 100,
      NA
    )
  )%>% 
  filter(lsoa21cd %in% london_lsoa_codes)
```

```{r, eval=FALSE}
census_health_cleaned <- census_health %>%
  mutate(
    health_bad_verybad = rowSums(
      across(
        c(
          general_health_bad,
          general_health_very_bad
        )
      ),
      na.rm = TRUE
    ),
    health_bad_verybad_pct = ifelse(
      general_health_total_residents > 0,
      (health_bad_verybad / general_health_total_residents)* 100,
      NA
    )
  )%>% 
  filter(lsoa21cd %in% london_lsoa_codes)
```

```{r}
census_ethnicity_cleaned <- census_ethnicity %>%
   mutate(
    ethnic_minority =
      ethnic_group_total_residents - white_british_group,
    
    ethnic_minority_pct = ifelse(
      ethnic_group_total_residents > 0,
      (ethnic_minority / ethnic_group_total_residents) * 100,
      NA
    )
  )%>% 
  filter(lsoa21cd %in% london_lsoa_codes)
```

```{r}
imd_score_cleaned <- imd_score %>% filter(lsoa21cd %in% london_lsoa_codes)
```

### Join Data

```{r}
london_join <- multimodal_accessibility_lsoa %>%
  # SUHII 
  left_join(
    suhii_lsoa %>%
      # drop geom cus multimodal travel time already have it
      st_drop_geometry() %>%
      # select needed columns
      dplyr::select(lsoa21cd, suhii),
    # join by lsoa code
    by = "lsoa21cd"
  ) %>%
  
  # IMD 
  left_join(
    imd_score_cleaned %>%
      dplyr::select(lsoa21cd, imd_score),
    by = "lsoa21cd"
  ) %>%
  
  # population density
  left_join(
    census_pop_density %>%
      dplyr::select(lsoa21cd, pop_density),
    by = "lsoa21cd"
  ) %>%
  
  # age
  left_join(
    census_age_cleaned %>%
      dplyr::select(lsoa21cd, age_4_and_under_pct, age_65_and_over_pct),
    by = "lsoa21cd"
  ) %>%
  
  # disability
  #left_join(
    #census_disability_cleaned %>%
      #select(lsoa21cd, disability_pct),
    #by = "lsoa21cd"
  #) %>%
  
  # health
  #left_join(
    #census_health_cleaned %>%
      #select(lsoa21cd, health_bad_verybad_pct),
    #by = "lsoa21cd"
  #) %>%
  
  # ethnicity
  left_join(
    census_ethnicity_cleaned %>%
      dplyr::select(lsoa21cd, ethnic_minority_pct),
    by = "lsoa21cd"
  )
```

```{r, eval=FALSE}
# check joined table
london_join %>% 
  st_drop_geometry() %>% 
  View()
```

```{r}
names(london_join)
```

```{r}
# create inner london control variable
# following: https://www.london.gov.uk/programmes-strategies/planning/london-plan/the-london-plan-2021-online/annex-2-inner-and-outer-london-boroughs
inner_london_boroughs <- c(
  "City of London", "Camden", "Greenwich", "Hackney",
  "Hammersmith and Fulham", "Islington", "Kensington and Chelsea",
  "Lambeth", "Lewisham", "Newham", "Southwark", "Tower Hamlets",
  "Wandsworth", "Westminster"
)

london_join <- london_join %>%
  mutate(
    #remove the letters numbers behind borough names to keep it a clean borough name only
    borough = str_remove(lsoa21nm, "\\s\\d+[A-Z]$"),
    # inner and outer london categories
    inner_outer_london = if_else(borough %in% inner_london_boroughs, "Inner", "Outer"),
    inner_outer_london = factor(inner_outer_london, levels = c("Outer", "Inner"))
  )
```

```{r}
#check
names(london_join)
```

## 2. Regression Data Preparation

```{r}
# check data types
check_data <- london_join %>%
  st_drop_geometry() %>%
  summarise(across(everything(), ~class(.)[1])) %>%
  pivot_longer(everything(), names_to = "variable", values_to = "type")

view(check_data)
```

```{r}
# descriptive stats
london_join %>%
  st_drop_geometry() %>%
  dplyr::select(where(is.numeric)) %>%
  summarise(across(everything(),
                   list(mean = ~mean(., na.rm = TRUE),
                        sd = ~sd(., na.rm = TRUE),
                        min = ~min(., na.rm = TRUE),
                        max = ~max(., na.rm = TRUE)))) %>%
  pivot_longer(everything(),
               names_to = c("variable", "stat"),
               names_pattern = "(.*)_(mean|sd|min|max)") %>%
  pivot_wider(names_from = stat, values_from = value) %>%
  kable(digits = 2)
```

## 3. Exploratory Data Analysis

**Plot Variables Distribution**

```{r}
# Select only numeric columns
numeric_cols <- london_join %>%
  sf::st_drop_geometry() %>%
  dplyr::select(where(is.numeric))

# Convert to long format for faceting
numeric_long <- numeric_cols %>%
  pivot_longer(cols = everything(), names_to = "variable", values_to = "value")

# Plot faceted histograms
variables_dist_hist <- ggplot(numeric_long, aes(x = value)) +
  geom_histogram(bins = 30, fill = "steelblue", color = "white") +
  facet_wrap(~ variable, scales = "free", ncol = 3) +
  theme_minimal() +
  labs(title = "Histograms of Numerical Variables")

variables_dist_hist
```

```{r, eval=FALSE}
# save as PNG
ggsave(
  filename = "figures/Fig7_Variables_Distribution_Hist.png",
  plot = variables_dist_hist,
  width = 7.5,
  height = 10,
  units = "in",
  dpi = 300,
  bg = "white"
)
```

**Plot Heatmap Correlation (Spearman)**

```{r}
# Select only the numeric variables (drop codes, names, geometry)
corr_matrix_spearman <- london_join %>%
  st_drop_geometry() %>%
  dplyr::select(
    access_score_15_pt,
    access_score_60_pt,
    access_score_15_walk,
    access_score_60_walk,
    access_score_15_cycle,
    access_score_60_cycle,
    imd_score,
    suhii,
    pop_density,
    age_4_and_under_pct,
    age_65_and_over_pct,
    ethnic_minority_pct
  ) %>%
  cor(
    use = "pairwise.complete.obs",
    method = "spearman"
  )

heatmap <- corrplot(
  corr_matrix_spearman,
  method = "color",
  type = "full",
  col = rev(COL2("RdBu", 200)),
  tl.col = "black",
  tl.srt = 45,
  addCoef.col = "black",
  number.cex = 0.6
)
```

```{r, eval=FALSE}
# save corrplot as png

dir.create("figures", showWarnings = FALSE)

png(
  filename = "figures/Fig8_Spearman_Heatmap.png",
  width = 6.5,
  height = 5.5,
  units = "in",
  res = 300,
  bg = "white"
)

corrplot(
  cor_matrix_spearman,
  method = "color",
  type = "full",
  col = rev(COL2("RdBu", 200)),
  tl.col = "black",
  tl.srt = 45,
  addCoef.col = "black",
  number.cex = 0.6
)

dev.off()
```

### Transform Variables

**Log transform population density and standardise other variables for clean interpretation**

This is due to the above examination of data, population density is highly skewed (so, log transform) and standardise all the variables that are in percentages.

```{r}
model_variables <- london_join %>%
  mutate(
    # apply log transforms
    log_pop_density = log(pop_density), 

    # standardise --> z-scores
    imd_z = as.numeric(scale(imd_score)),
    suhii_z = as.numeric(scale(suhii)),
    log_pop_density_z = as.numeric(scale(log_pop_density)),
    age_4_under_z = as.numeric(scale(age_4_and_under_pct)),
    age_65_over_z = as.numeric(scale(age_65_and_over_pct)),
    ethnic_minority_z  = as.numeric(scale(ethnic_minority_pct))
  )
```

```{r}
names(model_variables)
```

## 4. Model Choices

<https://rpubs.com/Geesaale/1263990> –\> count regression: poisson or negative binomial, depending on the nature of spatial dependence of data, spatial error more suitable (spatial error models is suitable when the spatial autocorrelation is due to an underlying spatial process or unmeasured factors).

## 5. Negative Binomial GLM

Loop all models:

```{r}
library(MASS)

# The predictors are the same for every model
predictors <- c(
  "imd_z * suhii_z",
  "log_pop_density_z",
  "age_4_under_z",
  "age_65_over_z",
  "ethnic_minority_z",
  "inner_outer_london"
)

# The six outcomes to loop over
outcomes <- c(
  #"access_score_15_pt",
  "access_score_60_pt",
  #"access_score_15_walk",
  "access_score_60_walk",
  "access_score_15_cycle",
  "access_score_60_cycle"
)

# Function to fit one negative binomial GLM given an outcome name
fit_nb <- function(outcome, data, predictors) {
  f <- reformulate(predictors, response = outcome)
  glm.nb(f, data = data)
}

# Loop over all outcomes, returning a named list of fitted models
models_nb <- lapply(outcomes, fit_nb, data = model_variables, predictors = predictors)
names(models_nb) <- outcomes
```

```{r}
# Print a labelled summary for each of the six models
for (nm in names(models_nb)) {
  cat("\n=====================================================\n")
  cat("MODEL:", nm, "\n")
  cat("=====================================================\n")
  print(summary(models_nb[[nm]]))
}
```

## 6. Poisson Regression

```{r}
# The predictors are the same for every model
predictors <- c(
  "imd_z * suhii_z",
  "log_pop_density_z",
  "age_4_under_z",
  "age_65_over_z",
  "ethnic_minority_z",
  "inner_outer_london"
)

# The six outcomes to loop over
outcomes <- c(
  "access_score_15_pt",
  #"access_score_60_pt",
  "access_score_15_walk"
  #"access_score_60_walk",
  #"access_score_15_cycle",
  #"access_score_60_cycle"
)

# Function to fit one Poisson GLM given an outcome name
fit_poisson <- function(outcome, data, predictors) {
  f <- reformulate(predictors, response = outcome)
  glm(f, data = model_variables, family = poisson(link = "log"))
}

# Loop over all outcomes, returning a named list of fitted models
models_poisson <- lapply(outcomes, fit_poisson, data = model_variables, predictors = predictors)
names(models_poisson) <- outcomes
```

```{r}
for (nm in names(models_poisson)) {
  cat("\n=====================================================\n")
  cat("MODEL:", nm, "\n")
  cat("=====================================================\n")
  print(summary(models_poisson[[nm]]))
}
```

```{r}
# calculate Poisson dispersion
check_dispersion <- function(model) {

  pearson_chisq <- sum(
    residuals(model, type = "pearson")^2,
    na.rm = TRUE
  )

  residual_df <- df.residual(model)

  dispersion_ratio <- pearson_chisq / residual_df

  # approximate test of overdispersion
  p_value <- pchisq(
    pearson_chisq,
    df = residual_df,
    lower.tail = FALSE
  )

  data.frame(
    pearson_chisq = pearson_chisq,
    residual_df = residual_df,
    dispersion_ratio = dispersion_ratio,
    p_value = p_value
  )
}

# apply to both Poisson models
dispersion_results <- do.call(
  rbind,
  lapply(models_poisson, check_dispersion)
)

# add model names
dispersion_results$model <- rownames(dispersion_results)
rownames(dispersion_results) <- NULL

# reorder columns
dispersion_results <- dispersion_results[
  c(
    "model",
    "dispersion_ratio",
    "pearson_chisq",
    "residual_df",
    "p_value"
  )
]

dispersion_results
```

Iteration limit for access_score_15_pt and access_score_15_walk, tried on Poisson across all models. Checked AIC values as below table (bold text refers to lower AIC values), their AIC values are also higher in the the negative binomial regression. As well as dispersion checked as above, dispersion values close to 1.

| Dependent variable    | AIC - Negative Binomial | AIC - Poisson |
|-----------------------|-------------------------|---------------|
| access_score_15_pt    | 13934                   | **13932**     |
| access_score_60_pt    | **49596**               | 90101         |
| access_score_15_walk  | 16249                   | **16247**     |
| access_score_60_walk  | **32617**               | 33312         |
| access_score_15_cycle | **21932**               | 22076         |
| access_score_60_cycle | **42929**               | 55092         |

## 7. Residuals with Moran's I

**Build spatial weights with LSOA boundary**

```{r}
#calculate the centroids of all Wards in London
coordsW <- london_lsoa%>%
  st_centroid()%>%
  st_geometry()

plot(coordsW)
```

```{r}
#Now we need to generate a spatial weights matrix 
#(remember from the lecture a couple of weeks ago). 
#We'll start with a simple binary matrix of queen's case neighbours

LLSOA_nb <- london_lsoa %>%
  poly2nb(., queen=T)

#or nearest neighbours
knn_wards <-coordsW %>%
  knearneigh(., k=4)

LLSOA_knn <- knn_wards %>%
  knn2nb()

#plot them
plot(LLSOA_nb, st_geometry(coordsW), col="red")
```

```{r}
plot(LLSOA_knn, st_geometry(coordsW), col="blue")
```

```{r}
#create a spatial weights matrix object from these weights

LLSOA.queens_weight <- LLSOA_nb %>%
  nb2listw(., style="W")

LLSOA.knn_4_weight <- LLSOA_knn %>%
  nb2listw(., style="W")
```

Residuals, normally Pearson is used for traditional GLMs. But count data, randomised quantile residuals is considered better (reference: <https://pmc.ncbi.nlm.nih.gov/articles/PMC7329451/>)

But here still uses Pearson:

```{r}
# dharma residuals

# install and load DHARMa package
#library(DHARMa)

# residuals
# https://cran.r-project.org/web/packages/DHARMa/vignettes/DHARMa.html#notes-on-particular-data-types 
#model_15_pt_resids <- simulateResiduals(fittedModel = model_15_pt)
```

### 7.1. access_score_15_pt

```{r}
# name object first
model_15_pt <- models_poisson$access_score_15_pt
# VIF
vif(model_15_pt)
```

Ok, VIF values are all below 5 or 10, no problem with multicolinearity, next look at residuals (auto)correlation

```{r}
# residuals with pearson
model_15_pt_resids <- residuals(model_15_pt, "pearson")

model_profile <- model_variables %>%
  mutate(model_15_pt_resids)

#now plot the residuals
tmap_mode("view")
qtm(model_profile, fill = "model_15_pt_resids")
```

```{r}
# queens case
model_15_pt_resids_queen <- model_profile %>%
  st_drop_geometry()%>%
  dplyr::select(model_15_pt_resids)%>%
  pull()%>%
  moran.test(., LLSOA.queens_weight)%>%
  tidy()

model_15_pt_resids_queen
```

```{r}
# nearest neighbour
model_15_pt_resids_nearest_neighbour <- model_profile %>%
  st_drop_geometry()%>%
  dplyr::select(model_15_pt_resids)%>%
  pull()%>%
  moran.test(., LLSOA.knn_4_weight)%>%
  tidy()

model_15_pt_resids_nearest_neighbour
```

### 7.2. access_score_60_pt

```{r}
# name object first
model_60_pt <- models_nb$access_score_60_pt

# VIF
vif(model_60_pt)
```

**Residuals and Map**

```{r}
# residuals with pearson
model_60_pt_resids <- residuals(model_60_pt, "pearson")

model_profile <- model_variables %>%
  mutate(model_60_pt_resids)

#now plot the residuals
tmap_mode("view")
qtm(model_profile, fill = "model_60_pt_resids")
```

**Moran's I - Queens Case**

```{r}
# queens case
model_60_pt_resids_queen <- model_profile %>%
  st_drop_geometry()%>%
  dplyr::select(model_60_pt_resids)%>%
  pull()%>%
  moran.test(., LLSOA.queens_weight)%>%
  tidy()

model_60_pt_resids_queen
```

**Moran's I - 4 Nearest Neighbours**

```{r}
# nearest neighbour
model_60_pt_resids_nearest_neighbour <- model_profile %>%
  st_drop_geometry()%>%
  dplyr::select(model_60_pt_resids)%>%
  pull()%>%
  moran.test(., LLSOA.knn_4_weight)%>%
  tidy()

model_60_pt_resids_nearest_neighbour
```

### 7.3. access_score_15_walk (Poisson)

```{r}
# name object first
model_15_walk <- models_poisson$access_score_15_walk
# VIF
vif(model_15_walk)
```

**Residuals and Map**

```{r}
# residuals with pearson
model_15_walk_resids <- residuals(model_15_walk, "pearson")

model_profile <- model_variables %>%
  mutate(model_15_walk_resids)

#now plot the residuals
tmap_mode("view")
qtm(model_profile, fill = "model_15_walk_resids")
```

**Moran's I - Queens Case**

```{r}
# queens case
model_15_walk_resids_queen <- model_profile %>%
  st_drop_geometry()%>%
  dplyr::select(model_15_walk_resids)%>%
  pull()%>%
  moran.test(., LLSOA.queens_weight)%>%
  tidy()

model_15_walk_resids_queen
```

**Moran's I - 4 Nearest Neighbours**

```{r}
# nearest neighbour
model_15_walk_resids_nearest_neighbour <- model_profile %>%
  st_drop_geometry()%>%
  dplyr::select(model_15_walk_resids)%>%
  pull()%>%
  moran.test(., LLSOA.knn_4_weight)%>%
  tidy()

model_15_walk_resids_nearest_neighbour
```

### 7.4. access_score_60_walk

```{r}
# name object first
model_60_walk <- models_nb$access_score_60_walk
# VIF
vif(model_60_walk)
```

**Residuals and Map**

```{r}
# residuals with pearson 
model_60_walk_resids <- residuals(model_60_walk, "pearson")  

model_profile <- model_variables %>%   
  mutate(model_60_walk_resids)  

#now plot the residuals 
tmap_mode("view") 
qtm(model_profile, fill = "model_60_walk_resids")
```

**Moran's I - Queens Case**

```{r}
# queens case 
model_60_walk_resids_queen <- model_profile %>%   
  st_drop_geometry()%>%   
  dplyr::select(model_60_walk_resids)%>%   
  pull()%>%   
  moran.test(., LLSOA.queens_weight)%>%   
  tidy()  

model_60_walk_resids_queen
```

**Moran's I - 4 Nearest Neighbours**

```{r}
# nearest neighbour 
model_60_walk_resids_nearest_neighbour <- model_profile %>%   
  st_drop_geometry()%>%   
  dplyr::select(model_60_walk_resids)%>%   
  pull()%>%   
  moran.test(., LLSOA.knn_4_weight)%>%   
  tidy()  

model_60_walk_resids_nearest_neighbour
```

### 7.5. access_score_15_cycle

```{r}
# name object first
model_15_cycle <- models_nb$access_score_15_cycle
# VIF
vif(model_15_cycle)
```

**Residuals and Map**

```{r}
# residuals with pearson  
model_15_cycle_resids <- residuals(model_15_cycle, "pearson")    

model_profile <- model_variables %>%      
  mutate(model_15_cycle_resids)    

#now plot the residuals  
tmap_mode("view")  
qtm(model_profile, fill = "model_15_cycle_resids")
```

**Moran's I - Queens Case**

```{r}
# queens case  
model_15_cycle_resids_queen <- model_profile %>%      
  st_drop_geometry()%>%      
  dplyr::select(model_15_cycle_resids)%>%      
  pull()%>%      
  moran.test(., LLSOA.queens_weight)%>%      
  tidy()    

model_15_cycle_resids_queen
```

**Moran's I - 4 Nearest Neighbours**

```{r}
# nearest neighbour  
model_15_cycle_resids_nearest_neighbour <- model_profile %>%      
  st_drop_geometry()%>%      
  dplyr::select(model_15_cycle_resids)%>%      
  pull()%>%      
  moran.test(., LLSOA.knn_4_weight)%>%      
  tidy()    

model_15_cycle_resids_nearest_neighbour
```

### 7.6. access_score_60_cycle

```{r}
# name object first
model_60_cycle <- models_nb$access_score_60_cycle
# VIF
vif(model_60_cycle)
```

**Residuals and Map**

```{r}
# residuals with pearson   
model_60_cycle_resids <- residuals(model_60_cycle, "pearson")      

model_profile <- model_variables %>%         
  mutate(model_60_cycle_resids)      

#now plot the residuals   
tmap_mode("view")   
qtm(model_profile, fill = "model_60_cycle_resids")
```

**Moran's I - Queens Case**

model_15_cycle

```{r}
# queens case   
model_60_cycle_resids_queen <- model_profile %>%         
  st_drop_geometry()%>%         
  dplyr::select(model_60_cycle_resids)%>%         
  pull()%>%         
  moran.test(., LLSOA.queens_weight)%>%         
  tidy()      

model_60_cycle_resids_queen
```

**Moran's I - 4 Nearest Neighbours**

```{r}
# nearest neighbour   
model_60_cycle_resids_nearest_neighbour <- model_profile %>%         
  st_drop_geometry()%>%         
  dplyr::select(model_60_cycle_resids)%>%         
  pull()%>%         
  moran.test(., LLSOA.knn_4_weight)%>%         
  tidy()      

model_60_cycle_resids_nearest_neighbour
```

### 7.7. Conclusion of Models' residuals

| Model          | Family            | Moran's I (Queens) | Moran's I (knn) |
|----------------|-------------------|--------------------|-----------------|
| model_15_pt    | Poisson           | 0.5119248          | 0.5322355       |
| model_60_pt    | Negative Binomial | 0.6729391          | 0.6872948       |
| model_15_walk  | Poisson           | 0.6667913          | 0.6918659       |
| model_60_walk  | Negative Binomial | 0.9055903          | 0.9090586       |
| model_15_cycle | Negative Binomial | 0.7601062          | 0.7765267       |
| model_60_cycle | Negative Binomial | 0.9209963          | 0.9120826       |

Strong positive spatial autocorrelation for queens case and nearest neighbours by Moran's I for all models, so next use spatial models

## 8. Spatial model -  Eigenvector Spatial Filtering (RE-ESF)

Given that the data type is count data, traditional spatial error models may not be suitable here as a direct transfer of the spatial model types for continuous data is not possible (Reference: <https://www.econstor.eu/bitstream/10419/168040/1/89655662X.pdf>).

Random effect - eigenvector spatial filtering (RE-ESP) to account for positive spatial autocorrelation/ dependence (identified in section 6,), can help reduce effect of spatial autocorrelation and reduce model bias and enahncing model fitting (<https://link.springer.com/article/10.1007/s11769-021-1191-y>) . This has been used in accessibility: (<https://www.sciencedirect.com/science/article/pii/S1618866721000807#sec0030>).

Using the `spmoran` package

```{r}
# rewrite just to be clear 
predictors <- c(
  "imd_z * suhii_z",
  "log_pop_density_z",
  "age_4_under_z",
  "age_65_over_z",
  "ethnic_minority_z",
  "inner_outer_london"
)

outcomes <- c(
  "access_score_15_pt",
  "access_score_60_pt",
  "access_score_15_walk",
  "access_score_60_walk",
  "access_score_15_cycle",
  "access_score_60_cycle"
)

# check 
names(model_variables)
```

## 9. GWPR - Geographically Weighted Poisson Regression

well known method to address issues of spatial nonstationarity, allowing geographic variations in the relationships between dependent and explanatory variables.
