---
title: "6_Statistical_Analyses"
---

This markdown file is the first draft script used for the complete analysis of `6_Statistical_Analyses.Rmd`, the final cleaned tidied version is under the project directory.

Spatial tools: <https://doc.esri.com/en/arcgis-pro/latest/tool-reference/spatial-statistics/an-overview-of-the-modeling-spatial-relationships-toolset.html>

```{r}
# remove objects from R environment and clear memory
rm(list = ls())    
gc()
```

## Load Packages

```{r}
library(sf)
library(sp)
library(readr)
library(dplyr)
library(knitr)
library(tidyr)
library(tibble)
library(tidyverse)
library(tmap)
library(cowplot)
library(plotly)
library(corrplot)
library(broom)
library(mapview)
library(stringr)
library(spdep)
library(car)
library(fs)
library(janitor)
library(GGally)
library(MASS)
library(spatialreg)
library(INLA)
library(INLAtools)
library(performance)
```

## Data Loading

Clean names, select initial needed columns and rename columns for consistency across datasets

**Greater London Boundary:**

```{r}
london_boundary <- st_read("data/gla/London_GLA_Boundary.shp")
```

**London only LSOAs**

```{r}
london_lsoa <- st_read("data/lsoa_london/lsoa_london.shp")
```

**Data obtained from previous pre-processing**

```{r}
# travel time 
multimodal_accessibility_lsoa <- st_read("data/multimodal_accessibility_lsoa.gpkg")%>%
  clean_names()%>%
  dplyr::select(
    lsoa21cd,
    lsoa21nm,
    access_score_15_pt,
    access_score_60_pt,
    access_score_15_walk,
    access_score_60_walk,
    access_score_15_cycle,
    access_score_60_cycle,
    travel_time_min_pt,
    travel_time_min_walk,
    travel_time_min_cycle,
    geom
  )%>%
  rename(
    geometry = geom
  )
```

```{r}
# SUHII as lsoa heat exposure
suhii_lsoa <- st_read("data/London_LSOA_LST_SUHI/London_LSOA_LST_SUHI.shp")%>% 
  st_transform(.,27700)%>%
  clean_names()%>%
  dplyr::select(
    lsoa21cd,
    lsoa21nm,
    # mean LST per LSOA
    mean,
    # london mean LST
    london_mea,
    # SUHII equivalent
    lst_differ,
    geometry
  )%>%
  # rename column names which were cut short from GEE export
  rename(
    mean_lst_lsoa = mean,
    london_mean = london_mea,
    suhii = lst_differ
  )
```

**IMD 2025 data**

```{r}
# IMD score
imd_score <- read_csv("data/IMD_2025.csv")%>%
  clean_names()%>%
  dplyr::select(
    lsoa_code_2021,
    lsoa_name_2021,
    index_of_multiple_deprivation_imd_score
    )%>%
  rename(
    lsoa21cd = lsoa_code_2021,
    lsoa21nm = lsoa_name_2021,
    imd_score = index_of_multiple_deprivation_imd_score
  )
```

**Census Data**

```{r}
# TS006 population density
census_pop_density <- read_csv("data/census2021/census2021-ts006-lsoa.csv")%>%
  clean_names()%>%
  dplyr::select(
    geography,
    geography_code,
    population_density_persons_per_square_kilometre_measures_value
  )%>%
  rename(
    lsoa21cd = geography_code,
    lsoa21nm = geography,
    pop_density = population_density_persons_per_square_kilometre_measures_value
  )
```

```{r}
# TS007a age by 5 year band
census_age_5year_band <- read_csv("data/census2021/census2021-ts007a-lsoa.csv")%>%
  clean_names()%>%
  dplyr::select(
    geography,
    geography_code,
    age_total,
    age_aged_4_years_and_under,
    # for filter for age 65 or over 
    age_aged_65_to_69_years,
    age_aged_70_to_74_years,
    age_aged_75_to_79_years,
    age_aged_80_to_84_years,
    age_aged_85_years_and_over
  )%>%
  rename(
    lsoa21cd = geography_code,
    lsoa21nm = geography,
    age_4_and_under = age_aged_4_years_and_under,
    age_65_to_69 = age_aged_65_to_69_years,
    age_70_to_74 = age_aged_70_to_74_years,
    age_75_to_79 = age_aged_75_to_79_years,
    age_80_to_84 = age_aged_80_to_84_years,
    age_85_and_over = age_aged_85_years_and_over
  )
```

```{r, eval=FALSE}
# TS038 disability
census_disability <- read_csv("data/census2021/census2021-ts038-lsoa.csv")%>%
  clean_names()%>%
  select(
    geography,
    geography_code,
    disability_total_all_usual_residents,
    disability_disabled_under_the_equality_act
  )%>%
  rename(
    lsoa21cd = geography_code,
    lsoa21nm = geography,
    # total number of residents 
    disability_total_residents = disability_total_all_usual_residents,
    # disabled under the Equality Act 2010
    disabled = disability_disabled_under_the_equality_act
  )
```

```{r, eval=FALSE}
# TS037 general health
census_health <- read_csv("data/census2021/census2021-ts037-lsoa.csv")%>% 
  clean_names()%>%
  select(
    geography,
    geography_code,
    general_health_total_all_usual_residents,
    general_health_bad_health,
    general_health_very_bad_health
  )%>%
  rename(
    lsoa21cd = geography_code,
    lsoa21nm = geography,
    general_health_total_residents = general_health_total_all_usual_residents,
    general_health_bad = general_health_bad_health,
    general_health_very_bad = general_health_very_bad_health
  )
```

```{r}
# TS021 Ethnicity
# definition of ethnic minority groups
# https://www.ethnicity-facts-figures.service.gov.uk/style-guide/writing-about-ethnicity/ 
# https://service-manual.ons.gov.uk/content/language/ethnicity-and-race
census_ethnicity <- read_csv("data/census2021/census2021-ts021-lsoa.csv")%>%
  clean_names()%>%
  dplyr::select(
    geography,
    geography_code,
    ethnic_group_total_all_usual_residents,
    # Use “Ethnic minorities” to refer to all ethnic groups except the White British group
    ethnic_group_white_english_welsh_scottish_northern_irish_or_british
  )%>%
  rename(
    lsoa21cd = geography_code,
    lsoa21nm = geography,
    ethnic_group_total_residents = ethnic_group_total_all_usual_residents,
    white_british_group = ethnic_group_white_english_welsh_scottish_northern_irish_or_british
  )
```

## 1. Data Cleaning

### Check Loaded + Initial Cleaned Data Column Names

```{r}
names(london_lsoa)
```

```{r}
# check 3 modes 
names(multimodal_accessibility_lsoa)
```

```{r}
# check names in SUHII export
names(suhii_lsoa)
```

```{r}
# check IMD names
names(imd_score)
```

```{r}
# check census data
names(census_pop_density)
```

```{r}
names(census_age_5year_band)
```

```{r}
names(census_ethnicity)
```

### Wrangle data variables for regression modelling

```{r}
# filter for london lsoas only
london_lsoa_codes <- london_lsoa %>%
  clean_names() %>%
  pull(lsoa21cd)
```

```{r}
# create percentage variables for age 4 and under and age 65 and over
census_age_cleaned <- census_age_5year_band %>%
  mutate(
    # create % of age 4 and under
    # calculate percentage
    age_4_and_under_pct = ifelse(
      age_total > 0,
      (age_4_and_under / age_total) * 100,
      NA
    ),
    # create % of age 65 and over
    age_65_and_over = rowSums(
      across(
        # sum up 65 to 85 and over
        c(
          age_65_to_69,
          age_70_to_74,
          age_75_to_79,
          age_80_to_84,
          age_85_and_over
        )
      ),
      na.rm = TRUE
    ),
    # calculate percentage
    age_65_and_over_pct = ifelse(
      age_total > 0,
      (age_65_and_over / age_total)* 100,
      NA
    )
  )%>% 
  # keep the ones that are within london lsoas only
  filter(lsoa21cd %in% london_lsoa_codes)
```

```{r}
# create ethnic minority group variable
census_ethnicity_cleaned <- census_ethnicity %>%
   mutate(
     # everything else but white british group - defined by ONS
    ethnic_minority =
      ethnic_group_total_residents - white_british_group,
    # calculate percentage
    ethnic_minority_pct = ifelse(
      ethnic_group_total_residents > 0,
      (ethnic_minority / ethnic_group_total_residents) * 100,
      NA
    )
  )%>% 
  # keep the ones that are within london lsoas only
  filter(lsoa21cd %in% london_lsoa_codes)
```

```{r}
# imd score variable 
# filter for london lsoas only
imd_score_cleaned <- imd_score %>% 
  filter(lsoa21cd %in% london_lsoa_codes)
```

### Join Data

```{r}
# join data
london_join <- multimodal_accessibility_lsoa %>%
  # SUHII 
  left_join(
    suhii_lsoa %>%
      # drop geom cus multimodal travel time already have it
      st_drop_geometry() %>%
      # select needed columns
      dplyr::select(lsoa21cd, suhii),
    # join by lsoa code
    by = "lsoa21cd"
  ) %>%
  
  # IMD 
  left_join(
    imd_score_cleaned %>%
      dplyr::select(lsoa21cd, imd_score),
    by = "lsoa21cd"
  ) %>%
  
  # population density
  left_join(
    census_pop_density %>%
      dplyr::select(lsoa21cd, pop_density),
    by = "lsoa21cd"
  ) %>%
  
  # age
  left_join(
    census_age_cleaned %>%
      dplyr::select(lsoa21cd, age_4_and_under_pct, age_65_and_over_pct),
    by = "lsoa21cd"
  ) %>%
  
  # ethnicity
  left_join(
    census_ethnicity_cleaned %>%
      dplyr::select(lsoa21cd, ethnic_minority_pct),
    by = "lsoa21cd"
  )
```

**Add an additional Inner and Outer London control + join to data**

```{r}
# create inner london control variable
# following: https://www.london.gov.uk/programmes-strategies/planning/london-plan/the-london-plan-2021-online/annex-2-inner-and-outer-london-boroughs
inner_london_boroughs <- c(
  "City of London", "Camden", "Greenwich", "Hackney",
  "Hammersmith and Fulham", "Islington", "Kensington and Chelsea",
  "Lambeth", "Lewisham", "Newham", "Southwark", "Tower Hamlets",
  "Wandsworth", "Westminster"
)

london_join <- london_join %>%
  mutate(
    #remove the letters numbers behind borough names to keep it a clean borough name only
    borough = str_remove(lsoa21nm, "\\s\\d+[A-Z]$"),
    # inner and outer london categories
    inner_outer_london = if_else(borough %in% inner_london_boroughs, "Inner", "Outer"),
    inner_outer_london = factor(inner_outer_london, levels = c("Outer", "Inner"))
  )
```

```{r}
#check fully joined data
names(london_join)
```

## 2. Check Data for Regression

```{r}
# check data types
check_data <- london_join %>%
  st_drop_geometry() %>%
  summarise(across(everything(), ~class(.)[1])) %>%
  pivot_longer(everything(), names_to = "variable", values_to = "type")

view(check_data)
```

```{r}
# descriptive stats
london_join %>%
  st_drop_geometry() %>%
  dplyr::select(where(is.numeric)) %>%
  summarise(across(everything(),
                   list(mean = ~mean(., na.rm = TRUE),
                        sd = ~sd(., na.rm = TRUE),
                        min = ~min(., na.rm = TRUE),
                        max = ~max(., na.rm = TRUE)))) %>%
  pivot_longer(everything(),
               names_to = c("variable", "stat"),
               names_pattern = "(.*)_(mean|sd|min|max)") %>%
  pivot_wider(names_from = stat, values_from = value) %>%
  kable(digits = 2)
```

## 3. Exploratory Data Analysis

**Plot Variables Distribution**

```{r}
# Select only numeric columns
numeric_cols <- london_join %>%
  sf::st_drop_geometry() %>%
  dplyr::select(where(is.numeric))

# Convert to long format for faceting
numeric_long <- numeric_cols %>%
  pivot_longer(cols = everything(), names_to = "variable", values_to = "value")

# Plot faceted histograms
variables_dist_hist <- ggplot(numeric_long, aes(x = value)) +
  geom_histogram(bins = 30, fill = "steelblue", color = "white") +
  facet_wrap(~ variable, scales = "free", ncol = 3) +
  theme_minimal() +
  labs(title = "Histograms of Numerical Variables")

#variables_dist_hist
```

**Plot Heatmap Correlation (Spearman)**

```{r}
# Select only the numeric variables (drop codes, names, geometry)
corr_matrix_spearman <- london_join %>%
  st_drop_geometry() %>%
  dplyr::select(
    access_score_15_pt,
    access_score_60_pt,
    access_score_15_walk,
    access_score_60_walk,
    access_score_15_cycle,
    access_score_60_cycle,
    suhii,
    imd_score,
    age_4_and_under_pct,
    age_65_and_over_pct,
    ethnic_minority_pct,
    pop_density
  ) %>%
  cor(
    use = "pairwise.complete.obs",
    method = "spearman"
  )

heatmap <- corrplot(
  corr_matrix_spearman,
  method = "color",
  type = "full",
  col = rev(COL2("RdBu", 200)),
  tl.col = "black",
  tl.srt = 45,
  addCoef.col = "black",
  number.cex = 0.6
)
```

```{r}
# save corrplot as png

dir.create("figures", showWarnings = FALSE)

png(
  filename = "figures/Fig8_Spearman_Heatmap.png",
  width = 8.5,
  height = 5.5,
  units = "in",
  res = 300,
  bg = "white"
)

corrplot(
  corr_matrix_spearman,
  method = "color",
  type = "full",
  col = rev(COL2("RdBu", 200)),
  tl.col = "black",
  tl.srt = 45,
  addCoef.col = "black",
  number.cex = 0.6
)

dev.off()
```

```{r}
dependents <- c(
  "access_score_15_pt",
  "access_score_60_pt",
  "access_score_15_walk",
  "access_score_60_walk",
  "access_score_15_cycle",
  "access_score_60_cycle"
)

distribution_check <- do.call(
  rbind,
  lapply(dependents, function(v) {

    y <- london_join[[v]]

    data.frame(
      dependents = v,
      n = sum(!is.na(y)),
      mean = mean(y, na.rm = TRUE),
      variance = var(y, na.rm = TRUE),
      variance_mean_ratio = var(y, na.rm = TRUE) /
        mean(y, na.rm = TRUE),
      zeros = sum(y == 0, na.rm = TRUE),
      zero_percent = mean(y == 0, na.rm = TRUE) * 100,
      maximum = max(y, na.rm = TRUE)
    )
  })
)

distribution_check
```

### 3.1. Transform Variables

**Log transform population density and standardise other variables**

As above examination of data, highly skewed population density needs to be log transformed and all the variables that are in percentages will be standardised.

```{r}
model_variables <- london_join %>%
  mutate(
    # log transforms pop density variable
    log_pop_density = log(pop_density), 

    # standardise variables by z-scores
    imd_z = as.numeric(scale(imd_score)),
    suhii_z = as.numeric(scale(suhii)),
    log_pop_density_z = as.numeric(scale(log_pop_density)),
    age_4_under_z = as.numeric(scale(age_4_and_under_pct)),
    age_65_over_z = as.numeric(scale(age_65_and_over_pct)),
    ethnic_minority_z  = as.numeric(scale(ethnic_minority_pct))
  )
```

## 4. Model Choices

<https://rpubs.com/Geesaale/1263990> –\> count regression: poisson or negative binomial, allow modelling of geographic variation in count data

## 5. Negative Binomial GLM

Loop all models:

```{r}
# independent variables to be the same for every model
independents <- c(
  "imd_z * suhii_z",
  "log_pop_density_z",
  "age_4_under_z",
  "age_65_over_z",
  "ethnic_minority_z",
  "inner_outer_london"
)

# dependent variables 
dependents <- c(
  #"access_score_15_pt",
  "access_score_60_pt",
  #"access_score_15_walk",
  "access_score_60_walk",
  "access_score_15_cycle",
  "access_score_60_cycle"
)

# fit negative binomial GLM for each dependent 
fit_nb <- function(dependents, data, independents) {
  f <- reformulate(independents, response = dependents)
  glm.nb(f, data = data)
}

# loop over dependent variables 
models_nb <- lapply(dependents, fit_nb, data = model_variables, independents = independents)
names(models_nb) <- dependents
```

```{r}
# print summary stats of all models
for (nm in names(models_nb)) {
  cat("\n=====================================================\n")
  cat("MODEL:", nm, "\n")
  cat("=====================================================\n")
  print(summary(models_nb[[nm]]))
}
```

## 6. Poisson Regression

```{r}
# independent variables to be the same for every model
independents <- c(
  "imd_z * suhii_z",
  "log_pop_density_z",
  "age_4_under_z",
  "age_65_over_z",
  "ethnic_minority_z",
  "inner_outer_london"
)

# dependent variables
dependents <- c(
  "access_score_15_pt",
  #"access_score_60_pt",
  "access_score_15_walk"
  #"access_score_60_walk",
  #"access_score_15_cycle",
  #"access_score_60_cycle"
)

# fit poisson GLM for each dependent 
fit_poisson <- function(dependents, data, independents) {
  f <- reformulate(independents, response = dependents)
  glm(f, data = model_variables, family = poisson(link = "log"))
}

# loop over dependent variables 
models_poisson <- lapply(dependents, fit_poisson, data = model_variables, independents = independents)
names(models_poisson) <- dependents
```

```{r}
# print summary stats of all models
for (nm in names(models_poisson)) {
  cat("\n=====================================================\n")
  cat("MODEL:", nm, "\n")
  cat("=====================================================\n")
  print(summary(models_poisson[[nm]]))
}
```

```{r}
# check Poisson dispersion
# https://www.rdocumentation.org/packages/insurancerating/versions/0.7.5/topics/check_overdispersion 
dispersion_results <- lapply(
  models_poisson,
  performance::check_overdispersion
)

dispersion_results
```

AIC values as below table (bold text refers to lower AIC values. As well as dispersion checked as above, dispersion values close to 1.

| Dependent variable    | AIC - Negative Binomial | AIC - Poisson |
|-----------------------|-------------------------|---------------|
| access_score_15_pt    | 13934                   | **13932**     |
| access_score_60_pt    | **49596**               | 90101         |
| access_score_15_walk  | 16249                   | **16247**     |
| access_score_60_walk  | **32617**               | 33312         |
| access_score_15_cycle | **21932**               | 22076         |
| access_score_60_cycle | **42929**               | 55092         |

## 7. Residuals with Moran's I

**Build spatial weights with LSOA boundary**

```{r}
#calculate the centroids of all LSOA in London
coordsLSOA <- london_lsoa%>%
  st_centroid()%>%
  st_geometry()

#plot(coordsLSOA)
```

```{r}
# generate spatial weights matrix 

# queen contiguity
LLSOA_nb <- london_lsoa %>%
  poly2nb(., queen=T)

# k nearest neighbours
knn_LSOA <-coordsLSOA %>%
  knearneigh(., k=4)

LLSOA_knn <- knn_LSOA %>%
  knn2nb()
```

```{r, eval=FALSE}
plot(LLSOA_nb, st_geometry(coordsLSOA), col="red")
```

```{r, eval=FALSE}
plot(LLSOA_knn, st_geometry(coordsLSOA), col="blue")
```

```{r}
#create a spatial weights matrix object from these weights

# queen contiguity spatial weight matrix
LLSOA.queens_weight <- LLSOA_nb %>%
  nb2listw(., style="W")

# k=4 nearest neighbours spatial weight matrix
LLSOA.knn_4_weight <- LLSOA_knn %>%
  nb2listw(., style="W")
```

### 7.1. access_score_15_pt

```{r}
# name object first
model_15_pt <- models_poisson$access_score_15_pt
# VIF
vif(model_15_pt)
```

VIF values are all below 5 or 10, no problem with multicolinearity, next look at residuals (auto)correlation (repeat for all)

**Residuals and Map**

```{r}
# residuals with pearson
model_15_pt_resids <- residuals(model_15_pt, "pearson")

model_profile <- model_variables %>%
  mutate(model_15_pt_resids)

#now plot the residuals
#tmap_mode("view")
#qtm(model_profile, fill = "model_15_pt_resids")
```

**Moran's I - Queens Case**

```{r}
# queens case
model_15_pt_resids_queen <- model_profile %>%
  st_drop_geometry()%>%
  dplyr::select(model_15_pt_resids)%>%
  pull()%>%
  moran.test(., LLSOA.queens_weight)%>%
  tidy()

model_15_pt_resids_queen
```

**Moran's I - 4 Nearest Neighbours**

```{r}
# nearest neighbour
model_15_pt_resids_nearest_neighbour <- model_profile %>%
  st_drop_geometry()%>%
  dplyr::select(model_15_pt_resids)%>%
  pull()%>%
  moran.test(., LLSOA.knn_4_weight)%>%
  tidy()

model_15_pt_resids_nearest_neighbour
```

### 7.2. access_score_60_pt

```{r}
# name object first
model_60_pt <- models_nb$access_score_60_pt

# VIF
vif(model_60_pt)
```

**Residuals and Map**

```{r}
# residuals with pearson
model_60_pt_resids <- residuals(model_60_pt, "pearson")

model_profile <- model_variables %>%
  mutate(model_60_pt_resids)

#now plot the residuals
#tmap_mode("view")
#qtm(model_profile, fill = "model_60_pt_resids")
```

**Moran's I - Queens Case**

```{r}
# queens case
model_60_pt_resids_queen <- model_profile %>%
  st_drop_geometry()%>%
  dplyr::select(model_60_pt_resids)%>%
  pull()%>%
  moran.test(., LLSOA.queens_weight)%>%
  tidy()

model_60_pt_resids_queen
```

**Moran's I - 4 Nearest Neighbours**

```{r}
# nearest neighbour
model_60_pt_resids_nearest_neighbour <- model_profile %>%
  st_drop_geometry()%>%
  dplyr::select(model_60_pt_resids)%>%
  pull()%>%
  moran.test(., LLSOA.knn_4_weight)%>%
  tidy()

model_60_pt_resids_nearest_neighbour
```

### 7.3. access_score_15_walk (Poisson)

```{r}
# name object first
model_15_walk <- models_poisson$access_score_15_walk
# VIF
vif(model_15_walk)
```

**Residuals and Map**

```{r}
# residuals with pearson
model_15_walk_resids <- residuals(model_15_walk, "pearson")

model_profile <- model_variables %>%
  mutate(model_15_walk_resids)

#now plot the residuals
#tmap_mode("view")
#qtm(model_profile, fill = "model_15_walk_resids")
```

**Moran's I - Queens Case**

```{r}
# queens case
model_15_walk_resids_queen <- model_profile %>%
  st_drop_geometry()%>%
  dplyr::select(model_15_walk_resids)%>%
  pull()%>%
  moran.test(., LLSOA.queens_weight)%>%
  tidy()

model_15_walk_resids_queen
```

**Moran's I - 4 Nearest Neighbours**

```{r}
# nearest neighbour
model_15_walk_resids_nearest_neighbour <- model_profile %>%
  st_drop_geometry()%>%
  dplyr::select(model_15_walk_resids)%>%
  pull()%>%
  moran.test(., LLSOA.knn_4_weight)%>%
  tidy()

model_15_walk_resids_nearest_neighbour
```

### 7.4. access_score_60_walk

```{r}
# name object first
model_60_walk <- models_nb$access_score_60_walk
# VIF
vif(model_60_walk)
```

**Residuals and Map**

```{r}
# residuals with pearson 
model_60_walk_resids <- residuals(model_60_walk, "pearson")  

model_profile <- model_variables %>%   
  mutate(model_60_walk_resids)  

#now plot the residuals 
#tmap_mode("view") 
#qtm(model_profile, fill = "model_60_walk_resids")
```

**Moran's I - Queens Case**

```{r}
# queens case 
model_60_walk_resids_queen <- model_profile %>%   
  st_drop_geometry()%>%   
  dplyr::select(model_60_walk_resids)%>%   
  pull()%>%   
  moran.test(., LLSOA.queens_weight)%>%   
  tidy()  

model_60_walk_resids_queen
```

**Moran's I - 4 Nearest Neighbours**

```{r}
# nearest neighbour 
model_60_walk_resids_nearest_neighbour <- model_profile %>%   
  st_drop_geometry()%>%   
  dplyr::select(model_60_walk_resids)%>%   
  pull()%>%   
  moran.test(., LLSOA.knn_4_weight)%>%   
  tidy()  

model_60_walk_resids_nearest_neighbour
```

### 7.5. access_score_15_cycle

```{r}
# name object first
model_15_cycle <- models_nb$access_score_15_cycle
# VIF
vif(model_15_cycle)
```

**Residuals and Map**

```{r}
# residuals with pearson  
model_15_cycle_resids <- residuals(model_15_cycle, "pearson")    

model_profile <- model_variables %>%      
  mutate(model_15_cycle_resids)    

#now plot the residuals  
#tmap_mode("view")  
#qtm(model_profile, fill = "model_15_cycle_resids")
```

**Moran's I - Queens Case**

```{r}
# queens case  
model_15_cycle_resids_queen <- model_profile %>%      
  st_drop_geometry()%>%      
  dplyr::select(model_15_cycle_resids)%>%      
  pull()%>%      
  moran.test(., LLSOA.queens_weight)%>%      
  tidy()    

model_15_cycle_resids_queen
```

**Moran's I - 4 Nearest Neighbours**

```{r}
# nearest neighbour  
model_15_cycle_resids_nearest_neighbour <- model_profile %>%      
  st_drop_geometry()%>%      
  dplyr::select(model_15_cycle_resids)%>%      
  pull()%>%      
  moran.test(., LLSOA.knn_4_weight)%>%      
  tidy()    

model_15_cycle_resids_nearest_neighbour
```

### 7.6. access_score_60_cycle

```{r}
# name object first
model_60_cycle <- models_nb$access_score_60_cycle
# VIF
vif(model_60_cycle)
```

**Residuals and Map**

```{r}
# residuals with pearson   
model_60_cycle_resids <- residuals(model_60_cycle, "pearson")      

model_profile <- model_variables %>%         
  mutate(model_60_cycle_resids)      

#now plot the residuals   
#tmap_mode("view")   
#qtm(model_profile, fill = "model_60_cycle_resids")
```

**Moran's I - Queens Case**

```{r}
# queens case   
model_60_cycle_resids_queen <- model_profile %>%         
  st_drop_geometry()%>%         
  dplyr::select(model_60_cycle_resids)%>%         
  pull()%>%         
  moran.test(., LLSOA.queens_weight)%>%         
  tidy()      

model_60_cycle_resids_queen
```

**Moran's I - 4 Nearest Neighbours**

```{r}
# nearest neighbour   
model_60_cycle_resids_nearest_neighbour <- model_profile %>%         
  st_drop_geometry()%>%         
  dplyr::select(model_60_cycle_resids)%>%         
  pull()%>%         
  moran.test(., LLSOA.knn_4_weight)%>%         
  tidy()      

model_60_cycle_resids_nearest_neighbour
```

### 7.7. Summary of Models' residuals

```{r}
# ============================================================
# GLM Moran's I summary table
# ============================================================

# Function to calculate Moran's I and extract Moran's I + p-value
get_moran <- function(residuals, weights) {
  
  test <- spdep::moran.test(residuals, weights)
  
  tibble::tibble(
    moran_i = unname(test$estimate[["Moran I statistic"]]),
    p_value = test$p.value
  )
}


# Run Moran's I tests for all six GLM residuals
glm_morans_table <- dplyr::bind_rows(
  
  # 15-minute PT
  dplyr::bind_cols(
    tibble::tibble(Model = "15-minute PT"),
    get_moran(
      model_15_pt_resids,
      LLSOA.queens_weight
    ) %>%
      dplyr::rename(
        `Queens Moran's I` = moran_i,
        `Queens p-value` = p_value
      ),
    get_moran(
      model_15_pt_resids,
      LLSOA.knn_4_weight
    ) %>%
      dplyr::rename(
        `KNN = 4 Moran's I` = moran_i,
        `KNN = 4 p-value` = p_value
      )
  ),
  
  # 60-minute PT
  dplyr::bind_cols(
    tibble::tibble(Model = "60-minute PT"),
    get_moran(
      model_60_pt_resids,
      LLSOA.queens_weight
    ) %>%
      dplyr::rename(
        `Queens Moran's I` = moran_i,
        `Queens p-value` = p_value
      ),
    get_moran(
      model_60_pt_resids,
      LLSOA.knn_4_weight
    ) %>%
      dplyr::rename(
        `KNN = 4 Moran's I` = moran_i,
        `KNN = 4 p-value` = p_value
      )
  ),
  
  # 15-minute Walking
  dplyr::bind_cols(
    tibble::tibble(Model = "15-minute Walking"),
    get_moran(
      model_15_walk_resids,
      LLSOA.queens_weight
    ) %>%
      dplyr::rename(
        `Queens Moran's I` = moran_i,
        `Queens p-value` = p_value
      ),
    get_moran(
      model_15_walk_resids,
      LLSOA.knn_4_weight
    ) %>%
      dplyr::rename(
        `KNN = 4 Moran's I` = moran_i,
        `KNN = 4 p-value` = p_value
      )
  ),
  
  # 60-minute Walking
  dplyr::bind_cols(
    tibble::tibble(Model = "60-minute Walking"),
    get_moran(
      model_60_walk_resids,
      LLSOA.queens_weight
    ) %>%
      dplyr::rename(
        `Queens Moran's I` = moran_i,
        `Queens p-value` = p_value
      ),
    get_moran(
      model_60_walk_resids,
      LLSOA.knn_4_weight
    ) %>%
      dplyr::rename(
        `KNN = 4 Moran's I` = moran_i,
        `KNN = 4 p-value` = p_value
      )
  ),
  
  # 15-minute Cycling
  dplyr::bind_cols(
    tibble::tibble(Model = "15-minute Cycling"),
    get_moran(
      model_15_cycle_resids,
      LLSOA.queens_weight
    ) %>%
      dplyr::rename(
        `Queens Moran's I` = moran_i,
        `Queens p-value` = p_value
      ),
    get_moran(
      model_15_cycle_resids,
      LLSOA.knn_4_weight
    ) %>%
      dplyr::rename(
        `KNN = 4 Moran's I` = moran_i,
        `KNN = 4 p-value` = p_value
      )
  ),
  
  # 60-minute Cycling
  dplyr::bind_cols(
    tibble::tibble(Model = "60-minute Cycling"),
    get_moran(
      model_60_cycle_resids,
      LLSOA.queens_weight
    ) %>%
      dplyr::rename(
        `Queens Moran's I` = moran_i,
        `Queens p-value` = p_value
      ),
    get_moran(
      model_60_cycle_resids,
      LLSOA.knn_4_weight
    ) %>%
      dplyr::rename(
        `KNN = 4 Moran's I` = moran_i,
        `KNN = 4 p-value` = p_value
      )
  )
)


# Format values for reporting
glm_morans_table_report <- glm_morans_table %>%
  dplyr::mutate(
    `Queens Moran's I` =
      sprintf("%.3f", `Queens Moran's I`),
    
    `Queens p-value` =
      format.pval(
        `Queens p-value`,
        digits = 3,
        eps = 0.001
      ),
    
    `KNN = 4 Moran's I` =
      sprintf("%.3f", `KNN = 4 Moran's I`),
    
    `KNN = 4 p-value` =
      format.pval(
        `KNN = 4 p-value`,
        digits = 3,
        eps = 0.001
      )
  )


# Display final table
knitr::kable(
  glm_morans_table_report,
  align = c("l", "c", "c", "c", "c"),
  caption = "Moran's I tests of GLM residual spatial autocorrelation"
)
```

| Model          | Family            | Moran's I (Queens) | Moran's I (knn) |
|----------------|-------------------|--------------------|-----------------|
| model_15_pt    | Poisson           | 0.5119248          | 0.5322355       |
| model_60_pt    | Negative Binomial | 0.6729391          | 0.6872948       |
| model_15_walk  | Poisson           | 0.6667913          | 0.6918659       |
| model_60_walk  | Negative Binomial | 0.9055903          | 0.9090586       |
| model_15_cycle | Negative Binomial | 0.7601062          | 0.7765267       |
| model_60_cycle | Negative Binomial | 0.9209963          | 0.9120826       |

Strong positive spatial autocorrelation for queens case and nearest neighbours by Moran's I for all models, so next use spatial models

## 8. Global Spatial Regression model (BYM2)

Given that the data type is **count data**, traditional **spatial error** models **may not be suitable** here ( <https://www.econstor.eu/bitstream/10419/168040/1/89655662X.pdf>).

Bayesian spatial multilevel mixed-effects model incoporating BMY2 with `INLA` package

```{r}
# INLA package: https://www.r-inla.org/download/index.html

# this works ocassionally:
#install.packages("INLA", repos=c(getOption("repos"), INLA="https://inla.r-inla-download.org/R/stable"), dep=TRUE)

# safer to download package with this:
install.packages("INLA", repos=c(getOption("repos"), INLA="https://inla.r-inla-download.org/R/testing"), dep=TRUE)
```

```{r}
# reload packages (if restarted R)
library(ggplot2)
library(tmap)
library(cowplot)
library(sf)
library(spdep)
library(dplyr)
library(tidyverse)
library(broom)
library(INLA)
library(INLAtools)
library(performance)
```

```{r}
# check INLA package version
packageVersion("INLA")
```

```{r}
# uncomment and add this if crash:
# inla.setOption(inla.mode="experimental") 

# check INLA version details
INLA::inla.version()
```

```{r}
# make a spatial model variables object to be clear
spatial_model_variables <- model_variables %>%
  dplyr::select(
    lsoa21cd,
    # dependent
    access_score_15_pt,
    access_score_60_pt,
    access_score_15_walk,
    access_score_60_walk,
    access_score_15_cycle,
    access_score_60_cycle,
    # independent
    imd_z,
    suhii_z,
    log_pop_density_z,
    age_4_under_z,
    age_65_over_z,
    ethnic_minority_z,
    inner_outer_london,
    geometry
  ) %>%
  arrange(lsoa21cd)
```

```{r}
# check if it is sf objects
class(spatial_model_variables)
```

### 8.1. Spatial Neighbourhood structure

```{r}
# from here onwards adapted code from: 
# https://www.r-inla.org/examples/AlvinBong/alvinbong-example.html, a tutorial that uses count-based areal data
# and also:
# https://www.sciencedirect.com/science/article/pii/S1353829225001765?via%3Dihub#sec2 --> https://github.com/mbcalles/Area-Level-Deprivation-Traffic-Injury

library(INLA)

# 1. Create a neighborhood list based on polygon contiguity - queens
nb <- spdep::poly2nb(
  spatial_model_variables,
  queen = TRUE
)

# 2. Convert the neighbor list to INLA adjacency format
spdep::nb2INLA("data/map.graph", nb)

# 3. Set the location of the adjacency file for INLA
map.adj <- file.path(getwd(), "data", "map.graph")

# 4. Load adjacency matrix as a graph for INLA spatial models
g <- inla.read.graph(filename = map.adj)
```

### 8.2. Inla call/ fit - Generalised Linear Models (Poisson)

```{r}
# index for spatial effects
spatial_model_variables$area_id <- 1:nrow(spatial_model_variables)
```

```{r}
# specify model formula
formula_15_pt <- access_score_15_pt ~
  imd_z * suhii_z +
  log_pop_density_z +
  age_4_under_z +
  age_65_over_z +
  ethnic_minority_z +
  factor(inner_outer_london) +
  f(
    area_id,
    model = "bym2",
    graph = g,
    constr = TRUE,
    scale.model = TRUE
  )


formula_60_pt <- access_score_60_pt ~
  imd_z * suhii_z +
  log_pop_density_z +
  age_4_under_z +
  age_65_over_z +
  ethnic_minority_z +
  factor(inner_outer_london) +
  f(
    area_id,
    model = "bym2",
    graph = g,
    constr = TRUE,
    scale.model = TRUE
  )


formula_15_walk <- access_score_15_walk ~
  imd_z * suhii_z +
  log_pop_density_z +
  age_4_under_z +
  age_65_over_z +
  ethnic_minority_z +
  factor(inner_outer_london) +
  f(
    area_id,
    model = "bym2",
    graph = g,
    constr = TRUE,
    scale.model = TRUE
  )


formula_60_walk <- access_score_60_walk ~
  imd_z * suhii_z +
  log_pop_density_z +
  age_4_under_z +
  age_65_over_z +
  ethnic_minority_z +
  factor(inner_outer_london) +
  f(
    area_id,
    model = "bym2",
    graph = g,
    constr = TRUE,
    scale.model = TRUE
  )


formula_15_cycle <- access_score_15_cycle ~
  imd_z * suhii_z +
  log_pop_density_z +
  age_4_under_z +
  age_65_over_z +
  ethnic_minority_z +
  factor(inner_outer_london) +
  f(
    area_id,
    model = "bym2",
    graph = g,
    constr = TRUE,
    scale.model = TRUE
  )


formula_60_cycle <- access_score_60_cycle ~
  imd_z * suhii_z +
  log_pop_density_z +
  age_4_under_z +
  age_65_over_z +
  ethnic_minority_z +
  factor(inner_outer_london) +
  f(
    area_id,
    model = "bym2",
    graph = g,
    constr = TRUE,
    scale.model = TRUE
  )
```

**15 pt**

```{r}
# run model
model_15_pt_bym2 <- INLA::inla(
  formula = formula_15_pt,
  family = "poisson",
  data = spatial_model_variables,
  control.predictor = list(compute = TRUE),
  # control.compute = list(return.marginals.predictor = TRUE), 
  # https://becarioprecario.bitbucket.io/inla-gitbook/ch-spatial.html#sec:lattice 
  control.compute = list(dic = TRUE, waic = TRUE, cpo = TRUE, config = TRUE, 
                         return.marginals.predictor=TRUE),
  verbose = FALSE
)

#summary(model_15_pt_bym2)
```

**60 pt**

```{r}
# run model
model_60_pt_bym2 <- INLA::inla(
  formula = formula_60_pt,
  family = "poisson",
  data = spatial_model_variables,
  control.predictor = list(compute = TRUE),
  #control.compute = list(return.marginals.predictor = TRUE), 
  control.compute = list(dic = TRUE, waic = TRUE, cpo = TRUE, config = TRUE, 
                         return.marginals.predictor=TRUE),
  verbose = FALSE
)

#summary(model_60_pt_bym2)
```

**15 walk**

```{r}
# run model
model_15_walk_bym2 <- INLA::inla(
  formula = formula_15_walk,
  family = "poisson",
  data = spatial_model_variables,
  control.predictor = list(compute = TRUE),
  #control.compute = list(return.marginals.predictor = TRUE), 
  control.compute = list(dic = TRUE, waic = TRUE, cpo = TRUE, config = TRUE, 
                         return.marginals.predictor=TRUE), 
  verbose = FALSE
)

#summary(model_15_walk_bym2)
```

**60 walk**

```{r}
# run model
model_60_walk_bym2 <- INLA::inla(
  formula = formula_60_walk,
  family = "poisson",
  data = spatial_model_variables, 
  control.predictor = list(compute = TRUE),
  #control.compute = list(return.marginals.predictor = TRUE), 
  control.compute = list(dic = TRUE, waic = TRUE, cpo = TRUE, config = TRUE, 
                         return.marginals.predictor=TRUE),
  verbose = FALSE
)

#summary(model_60_walk_bym2)
```

Save as a RDS object:

```{r}
saveRDS(model_15_pt_bym2, "data/BYM2/model_15_pt_bym2.rds")
saveRDS(model_60_pt_bym2, "data/BYM2/model_60_pt_bym2.rds")
saveRDS(model_15_walk_bym2, "data/BYM2//model_15_walk_bym2.rds")
saveRDS(model_60_walk_bym2, "data/BYM2/model_60_walk_bym2.rds")
```

```{r}
# clear the bym2 memory
rm(
  model_15_pt_bym2,
  model_60_pt_bym2,
  model_15_walk_bym2,
  model_60_walk_bym2
)

gc()
```

**15 cycle**

```{r}
# run model
model_15_cycle_bym2 <- INLA::inla(
  formula = formula_15_cycle,
  family = "poisson",
  data = spatial_model_variables, 
  control.predictor = list(compute = TRUE),
  #control.compute = list(return.marginals.predictor = TRUE), 
  control.compute = list(dic = TRUE, waic = TRUE, cpo = TRUE, config = TRUE, 
                         return.marginals.predictor=TRUE),
  verbose = FALSE
)

#summary(model_15_cycle_bym2)

# save as RDS
saveRDS(model_15_cycle_bym2, "data/BYM2/model_15_cycle_bym2.rds")
```

**60 cycle**

```{r}
# run model
model_60_cycle_bym2 <- INLA::inla(
  formula = formula_60_cycle,
  family = "poisson",
  data = spatial_model_variables,
  control.predictor = list(compute = TRUE),
  #control.compute = list(return.marginals.predictor = TRUE), 
  control.compute = list(dic = TRUE, waic = TRUE, cpo = TRUE, config = TRUE, 
                         return.marginals.predictor=TRUE), 
  verbose = FALSE
)

#summary(model_60_cycle_bym2)

# save as RDS
saveRDS(model_60_cycle_bym2,"data/BYM2/model_60_cycle_bym2.rds")
```

```{r}
# reload
model_15_pt_bym2    <- readRDS("data/BYM2/model_15_pt_bym2.rds")
model_60_pt_bym2    <- readRDS("data/BYM2/model_60_pt_bym2.rds")
model_15_walk_bym2  <- readRDS("data/BYM2/model_15_walk_bym2.rds")
model_60_walk_bym2  <- readRDS("data/BYM2/model_60_walk_bym2.rds")
model_15_cycle_bym2 <- readRDS("data/BYM2/model_15_cycle_bym2.rds")
model_60_cycle_bym2 <- readRDS("data/BYM2/model_60_cycle_bym2.rds")
```

```{r}
# name models for clean interpretation
bym2_models <- list(
  "Model 1" = model_15_pt_bym2,
  "Model 2" = model_60_pt_bym2,
  "Model 3" = model_15_walk_bym2,
  "Model 4" = model_60_walk_bym2,
  "Model 5" = model_15_cycle_bym2,
  "Model 6" = model_60_cycle_bym2
)

# name variables 
variable_labels <- c(
  "(Intercept)" = "Intercept",
  "imd_z" = "IMD (z)",
  "suhii_z" = "SUHII (z)",
  "imd_z:suhii_z" = "IMD × SUHII",
  "log_pop_density_z" = "Log pop. density (z)",
  "age_4_under_z" = "Age 4 or under % (z)",                 
  "age_65_over_z" = "Age 65 or over % (z)",
  "ethnic_minority_z" = "Ethnic minority % (z)",
  "factor(inner_outer_london)Inner" = "Inner, Outer London"
)

# Extract fixed effects, exponentiate to rate ratios, format as one cell
# just like this: https://www.sciencedirect.com/science/article/pii/S2405673123000144
fixed_effects_long <- lapply(names(bym2_models), function(model_name) {
  bym2_models[[model_name]]$summary.fixed %>%
    as.data.frame() %>%
    rownames_to_column("term") %>%
    transmute(
      model = model_name,
      variable = variable_labels[term],
      cell = sprintf(
        "%.3f (%.3f, %.3f)",
        # mean IRR and the 95% CI
        exp(mean), exp(`0.025quant`), exp(`0.975quant`)
      )
    )
}) %>% bind_rows()

# One column per model, one row per variable
fixed_effects_table <- fixed_effects_long %>%
  pivot_wider(names_from = model, values_from = cell) %>%
  arrange(match(variable, variable_labels))

knitr::kable(
  fixed_effects_table,
  caption = "BYM2 Fixed Effects Mean Incidence Rate Ratios and 95% Credible Intervals (CI)"
)
```

### 8.3. Validation Checks

#### 8.3.1. Diagnostic

<https://becarioprecario.bitbucket.io/inla-gitbook/ch-INLA.html#sec:modelassess>

```{r}
# List the current six models
six_models <- list(
  "15-minute PT"    = model_15_pt_bym2,
  "60-minute PT"    = model_60_pt_bym2,
  "15-minute Walking"  = model_15_walk_bym2,
  "60-minute Walking"  = model_60_walk_bym2,
  "15-minute Cycling" = model_15_cycle_bym2,
  "60-minute Cycling" = model_60_cycle_bym2
)
```

```{r}
# Adjusted PIT histograms
png(
  filename = "figures/FigC1_Adjusted_PIT_Histograms.png",
  width = 2400,
  height = 1600,
  res = 300
)

par(mfrow = c(2, 3), mar = c(3, 3, 2, 1))

for (nm in names(six_models)) {

  m <- six_models[[nm]]

  # use valid CPO/PIT observations only
  valid <- is.finite(m$cpo$pit) &
           is.finite(m$cpo$cpo) &
           m$cpo$cpo > 0 &
           m$cpo$failure == 0

  # adjusted PIT for discrete count data
  adjusted_pit <- m$cpo$pit[valid] -
                  0.5 * m$cpo$cpo[valid]

  # histogram information without plotting
  hist(
     adjusted_pit,
     breaks = seq(0, 1, by = 0.05),
     probability = TRUE,
     main = nm,
     xlab = "Adjusted PIT",
     ylab = "Density",
     xlim = c(0, 1),
     # keep y axis consistent for all plots
     ylim = c(0, 3),
     yaxt = "n",  
     col = "steelblue"
   )
 
 
   axis(
     side = 2,
     at = seq(0, 3, by = 0.5),
     labels = sprintf("%.1f", seq(0, 3, by = 0.5))
   )
 
   # expected density under uniform PIT distribution
   abline(h = 1, lty = 2)
}

dev.off()

par(mfrow = c(1, 1))
```

```{r}
# Adjusted PIT histograms
png(
  filename = "figures/Fig_Adjusted_PIT_Histograms.png",
  width = 2400,
  height = 1600,
  res = 300
)

par(mfrow = c(2, 3), mar = c(3, 3, 2, 1))

# keep breaks fixed
pit_breaks <- seq(0, 1, by = 0.05)

# first pass: find a common maximum relative frequency across all plots
max_rel_freq <- 0

for (nm in names(six_models)) {

  m <- six_models[[nm]]

  valid <- is.finite(m$cpo$pit) &
           is.finite(m$cpo$cpo) &
           m$cpo$cpo > 0 &
           m$cpo$failure == 0

  adjusted_pit <- m$cpo$pit[valid] -
                  0.5 * m$cpo$cpo[valid]

  h_tmp <- hist(
    adjusted_pit,
    breaks = pit_breaks,
    plot = FALSE
  )

  rel_freq_tmp <- h_tmp$counts / sum(h_tmp$counts)
  max_rel_freq <- max(max_rel_freq, rel_freq_tmp)
}

# round up a bit so all panels share the same y-axis nicely
y_max <- ceiling(max_rel_freq / 0.01) * 0.01

# second pass: make plots
for (nm in names(six_models)) {

  m <- six_models[[nm]]

  # use valid CPO/PIT observations only
  valid <- is.finite(m$cpo$pit) &
           is.finite(m$cpo$cpo) &
           m$cpo$cpo > 0 &
           m$cpo$failure == 0

  # adjusted PIT for discrete count data
  adjusted_pit <- m$cpo$pit[valid] -
                  0.5 * m$cpo$cpo[valid]

  # histogram info only
  h <- hist(
    adjusted_pit,
    breaks = pit_breaks,
    plot = FALSE
  )

  # convert counts to relative frequencies
  h$counts <- h$counts / sum(h$counts)

  # plot relative-frequency histogram
  plot(
    h,
    freq = TRUE,
    main = nm,
    xlab = "Adjusted PIT",
    ylab = "Relative frequency",
    xlim = c(0, 1),
    ylim = c(0, y_max),
    yaxt = "n",
    col = "steelblue"
  )

  axis(
    side = 2,
    at = seq(0, y_max, by = 0.02),
    labels = sprintf("%.2f", seq(0, y_max, by = 0.02))
  )

  # expected relative frequency under uniform PIT distribution
  abline(h = 1 / (length(pit_breaks) - 1), lty = 2)
}

dev.off()

par(mfrow = c(1, 1))
```

```{r}
# summary of DIC, WAIC CPO and MLIK values
model_summary <- lapply(names(six_models), function(model_name) {

  m <- six_models[[model_name]]
  valid_cpo <- is.finite(m$cpo$cpo) & m$cpo$cpo > 0

  data.frame(
    Model = model_name,
    DIC = m$dic$dic,
    WAIC = m$waic$waic,
    Negative_sum_log_CPO = -sum(log(m$cpo$cpo[valid_cpo])),
    Log_Marginal_Likelihood =
      m$mlik[1, 1]
  )

}) %>%
  dplyr::bind_rows()

model_summary
```

```{r}
# check if grraph nodes matches LSOAs

# Number of graph nodes
cat("Number of graph nodes:", g$n, "\n")
cat("Number of LSOAs:", nrow(spatial_model_variables), "\n")

cat(
  "Graph and data sizes match:",
  g$n == nrow(spatial_model_variables),
  "\n"
)

cat(
  "area_id is correctly ordered:",
  identical(
    spatial_model_variables$area_id,
    seq_len(nrow(spatial_model_variables))
  ),
  "\n"
)

cat(
  "Number of duplicated LSOA codes:",
  anyDuplicated(spatial_model_variables$lsoa21cd),
  "\n"
)
```

#### 8.3.2. CPO failures

```{r, eval=FALSE}
bym2_models <- list(
  "15-minute Public Transit" = model_15_pt_bym2,
  "60-minute Public Transit" = model_60_pt_bym2,
  "15-minute Walking" = model_15_walk_bym2,
  "60-minute Walking" = model_60_walk_bym2,
  "15-minute Cycling" = model_15_cycle_bym2,
  "60-minute Cycling" = model_60_cycle_bym2
)

cpo_failure_counts <- sapply(
  bym2_models,
  function(model) sum(model$cpo$failure > 0, na.rm = TRUE)
)

cpo_failure_counts
```

```{r, eval=FALSE}
# recompute
model_15_pt_bym2_cpo <- INLA::inla.cpo(
  model_15_pt_bym2
)

model_15_walk_bym2_cpo <- INLA::inla.cpo(
  model_15_walk_bym2
)

model_15_cycle_bym2_cpo <- INLA::inla.cpo(
  model_15_cycle_bym2
)

sum(log(model_15_pt_bym2_cpo$cpo$cpo), na.rm = TRUE)
sum(log(model_15_walk_bym2_cpo$cpo$cpo), na.rm = TRUE)
sum(log(model_15_cycle_bym2_cpo$cpo$cpo), na.rm = TRUE)
```

Barely any changes after recompute, move on...

#### 8.3.3. Residuals

Rebuild the centroids, queens, knn first

```{r}
# centroids 
coordsLSOA2 <- spatial_model_variables %>% 
  st_centroid() %>% 
  st_geometry()

plot(coordsLSOA2)
```

map

```{r}
# queens
LLSOA_nb2 <- spatial_model_variables %>%
  poly2nb(., queen = TRUE)

#or nearest neighbours
knn_LSOA2 <-coordsLSOA2 %>%
  knearneigh(., k=4)

LLSOA_knn2 <- knn_LSOA2 %>%
  knn2nb()
```

```{r, eval=FALSE}
#plot them
plot(LLSOA_nb2, st_geometry(coordsLSOA2), col="red")
```

```{r, eval=FALSE}
plot(LLSOA_knn, st_geometry(coordsLSOA2), col="blue")
```

```{r}
#create a spatial weights matrix object from these weights

LLSOA.queens_weight2 <- LLSOA_nb2 %>%
  nb2listw(., style="W")

LLSOA.knn_4_weight2 <- LLSOA_knn2 %>%
  nb2listw(., style="W")
```

**Pearson Residuals - i = (Y_i − λ̂\_i) / sqrt(λ̂\_i)**

**15 pt**

```{r}
model_15_pt_bym2_resids <- (
  # observed minus 
  spatial_model_variables$access_score_15_pt -
    # fitted value
    model_15_pt_bym2$summary.fitted.values$mean
  # divided by square root of fitted values
)/sqrt(model_15_pt_bym2$summary.fitted.values$mean)

model_profile <- spatial_model_variables %>%
  mutate(model_15_pt_bym2_resids)

#now plot the residuals
#tmap_mode("view")
#qtm(model_profile, fill = "model_15_pt_bym2_resids")
```

```{r}
# queens case
model_15_pt_bym2_resids_queen <- model_profile %>%
  st_drop_geometry()%>%
  dplyr::select(model_15_pt_bym2_resids)%>%
  pull()%>%
  moran.test(., LLSOA.queens_weight2)%>%
  tidy()

model_15_pt_bym2_resids_queen
```

```{r}
# nearest neighbour
model_15_pt_bym2_resids_nearest_neighbour <- model_profile %>%
  st_drop_geometry()%>%
  dplyr::select(model_15_pt_bym2_resids)%>%
  pull()%>%
  moran.test(., LLSOA.knn_4_weight2)%>%
  tidy()

model_15_pt_bym2_resids_nearest_neighbour
```

**60 pt**

```{r}
model_60_pt_bym2_resids <- (
  spatial_model_variables$access_score_60_pt -
    model_60_pt_bym2$summary.fitted.values$mean
)/sqrt(model_60_pt_bym2$summary.fitted.values$mean)

model_profile <- spatial_model_variables %>%
  mutate(model_60_pt_bym2_resids)

#now plot the residuals
#tmap_mode("view")
#qtm(model_profile, fill = "model_60_pt_bym2_resids")
```

```{r}
# queens case
model_60_pt_bym2_resids_queen <- model_profile %>%
  st_drop_geometry()%>%
  dplyr::select(model_60_pt_bym2_resids)%>%
  pull()%>%
  moran.test(., LLSOA.queens_weight2)%>%
  tidy()

model_60_pt_bym2_resids_queen
```

```{r}
# nearest neighbour
model_60_pt_bym2_resids_nearest_neighbour <- model_profile %>%
  st_drop_geometry()%>%
  dplyr::select(model_60_pt_bym2_resids)%>%
  pull()%>%
  moran.test(., LLSOA.knn_4_weight2)%>%
  tidy()

model_60_pt_bym2_resids_nearest_neighbour
```

**15 walk**

```{r}
model_15_walk_bym2_resids <- (
  spatial_model_variables$access_score_15_walk -
    model_15_walk_bym2$summary.fitted.values$mean
)/sqrt(model_15_walk_bym2$summary.fitted.values$mean)

model_profile <- spatial_model_variables %>%
  mutate(model_15_walk_bym2_resids)

#now plot the residuals
#tmap_mode("view")
#qtm(model_profile, fill = "model_15_walk_bym2_resids")
```

```{r}
# queens case
model_15_walk_bym2_resids_queen <- model_profile %>%
  st_drop_geometry()%>%
  dplyr::select(model_15_walk_bym2_resids)%>%
  pull()%>%
  moran.test(., LLSOA.queens_weight2)%>%
  tidy()

model_15_walk_bym2_resids_queen
```

```{r}
# nearest neighbour
model_15_walk_bym2_resids_nearest_neighbour <- model_profile %>%
  st_drop_geometry()%>%
  dplyr::select(model_15_walk_bym2_resids)%>%
  pull()%>%
  moran.test(., LLSOA.knn_4_weight2)%>%
  tidy()

model_15_walk_bym2_resids_nearest_neighbour
```

**60 walk**

```{r}
model_60_walk_bym2_resids <- (
  spatial_model_variables$access_score_60_walk -
    model_60_walk_bym2$summary.fitted.values$mean
)/sqrt(model_60_walk_bym2$summary.fitted.values$mean)

model_profile <- spatial_model_variables %>%
  mutate(model_60_walk_bym2_resids)

#now plot the residuals
#tmap_mode("view")
#qtm(model_profile, fill = "model_60_walk_bym2_resids")
```

```{r}
# queens case
model_60_walk_bym2_resids_queen <- model_profile %>%
  st_drop_geometry()%>%
  dplyr::select(model_60_walk_bym2_resids)%>%
  pull()%>%
  moran.test(., LLSOA.queens_weight2)%>%
  tidy()

model_60_walk_bym2_resids_queen
```

```{r}
# nearest neighbour
model_60_walk_bym2_resids_nearest_neighbour <- model_profile %>%
  st_drop_geometry()%>%
  dplyr::select(model_60_walk_bym2_resids)%>%
  pull()%>%
  moran.test(., LLSOA.knn_4_weight2)%>%
  tidy()

model_60_walk_bym2_resids_nearest_neighbour
```

**15 cycle**

```{r}
model_15_cycle_bym2_resids <- (
  spatial_model_variables$access_score_15_cycle -
    model_15_cycle_bym2$summary.fitted.values$mean
)/sqrt(model_15_cycle_bym2$summary.fitted.values$mean)

model_profile <- spatial_model_variables %>%
  mutate(model_15_cycle_bym2_resids)

#now plot the residuals
#tmap_mode("view")
#qtm(model_profile, fill = "model_15_cycle_bym2_resids")
```

```{r}
# queens case
model_15_cycle_bym2_resids_queen <- model_profile %>%
  st_drop_geometry()%>%
  dplyr::select(model_15_cycle_bym2_resids)%>%
  pull()%>%
  moran.test(., LLSOA.queens_weight2)%>%
  tidy()

model_15_cycle_bym2_resids_queen
```

```{r}
# nearest neighbour
model_15_cycle_bym2_resids_nearest_neighbour <- model_profile %>%
  st_drop_geometry()%>%
  dplyr::select(model_15_cycle_bym2_resids)%>%
  pull()%>%
  moran.test(., LLSOA.knn_4_weight2)%>%
  tidy()

model_15_cycle_bym2_resids_nearest_neighbour
```

**60 cycle**

```{r}
model_60_cycle_bym2_resids <- (
  spatial_model_variables$access_score_60_cycle -
    model_60_cycle_bym2$summary.fitted.values$mean
)/sqrt(model_60_cycle_bym2$summary.fitted.values$mean)

model_profile <- spatial_model_variables %>%
  mutate(model_60_cycle_bym2_resids)

#now plot the residuals
#tmap_mode("view")
#qtm(model_profile, fill = "model_60_cycle_bym2_resids")
```

```{r}
# queens case
model_60_cycle_bym2_resids_queen <- model_profile %>%
  st_drop_geometry()%>%
  dplyr::select(model_60_cycle_bym2_resids)%>%
  pull()%>%
  moran.test(., LLSOA.queens_weight2)%>%
  tidy()

model_60_cycle_bym2_resids_queen
```

```{r}
# nearest neighbour
model_60_cycle_bym2_resids_nearest_neighbour <- model_profile %>%
  st_drop_geometry()%>%
  dplyr::select(model_60_cycle_bym2_resids)%>%
  pull()%>%
  moran.test(., LLSOA.knn_4_weight2)%>%
  tidy()

model_60_cycle_bym2_resids_nearest_neighbour
```

significantly reduced compared to the GLM models, but still some spatial autocorrelation remains –\> write in discussion

```{r}
# ============================================================
# BYM2 Moran's I summary table
# ============================================================

# Function to calculate Moran's I and extract Moran's I + p-value
get_moran <- function(residuals, weights) {
  
  test <- spdep::moran.test(residuals, weights)
  
  tibble::tibble(
    moran_i = unname(test$estimate[["Moran I statistic"]]),
    p_value = test$p.value
  )
}


# Run Moran's I tests for all six BYM2 residuals
bym2_morans_table <- dplyr::bind_rows(
  
  # 15-minute PT
  dplyr::bind_cols(
    tibble::tibble(Model = "15-minute PT"),
    get_moran(
      model_15_pt_bym2_resids,
      LLSOA.queens_weight2
    ) %>%
      dplyr::rename(
        `Queens Moran's I` = moran_i,
        `Queens p-value` = p_value
      ),
    get_moran(
      model_15_pt_bym2_resids,
      LLSOA.knn_4_weight2
    ) %>%
      dplyr::rename(
        `KNN = 4 Moran's I` = moran_i,
        `KNN = 4 p-value` = p_value
      )
  ),
  
  # 60-minute PT
  dplyr::bind_cols(
    tibble::tibble(Model = "60-minute PT"),
    get_moran(
      model_60_pt_bym2_resids,
      LLSOA.queens_weight2
    ) %>%
      dplyr::rename(
        `Queens Moran's I` = moran_i,
        `Queens p-value` = p_value
      ),
    get_moran(
      model_60_pt_bym2_resids,
      LLSOA.knn_4_weight2
    ) %>%
      dplyr::rename(
        `KNN = 4 Moran's I` = moran_i,
        `KNN = 4 p-value` = p_value
      )
  ),
  
  # 15-minute Walking
  dplyr::bind_cols(
    tibble::tibble(Model = "15-minute Walking"),
    get_moran(
      model_15_walk_bym2_resids,
      LLSOA.queens_weight2
    ) %>%
      dplyr::rename(
        `Queens Moran's I` = moran_i,
        `Queens p-value` = p_value
      ),
    get_moran(
      model_15_walk_bym2_resids,
      LLSOA.knn_4_weight2
    ) %>%
      dplyr::rename(
        `KNN = 4 Moran's I` = moran_i,
        `KNN = 4 p-value` = p_value
      )
  ),
  
  # 60-minute Walking
  dplyr::bind_cols(
    tibble::tibble(Model = "60-minute Walking"),
    get_moran(
      model_60_walk_bym2_resids,
      LLSOA.queens_weight2
    ) %>%
      dplyr::rename(
        `Queens Moran's I` = moran_i,
        `Queens p-value` = p_value
      ),
    get_moran(
      model_60_walk_bym2_resids,
      LLSOA.knn_4_weight2
    ) %>%
      dplyr::rename(
        `KNN = 4 Moran's I` = moran_i,
        `KNN = 4 p-value` = p_value
      )
  ),
  
  # 15-minute Cycling
  dplyr::bind_cols(
    tibble::tibble(Model = "15-minute Cycling"),
    get_moran(
      model_15_cycle_bym2_resids,
      LLSOA.queens_weight2
    ) %>%
      dplyr::rename(
        `Queens Moran's I` = moran_i,
        `Queens p-value` = p_value
      ),
    get_moran(
      model_15_cycle_bym2_resids,
      LLSOA.knn_4_weight2
    ) %>%
      dplyr::rename(
        `KNN = 4 Moran's I` = moran_i,
        `KNN = 4 p-value` = p_value
      )
  ),
  
  # 60-minute Cycling
  dplyr::bind_cols(
    tibble::tibble(Model = "60-minute Cycling"),
    get_moran(
      model_60_cycle_bym2_resids,
      LLSOA.queens_weight2
    ) %>%
      dplyr::rename(
        `Queens Moran's I` = moran_i,
        `Queens p-value` = p_value
      ),
    get_moran(
      model_60_cycle_bym2_resids,
      LLSOA.knn_4_weight2
    ) %>%
      dplyr::rename(
        `KNN = 4 Moran's I` = moran_i,
        `KNN = 4 p-value` = p_value
      )
  )
)


# Format values for reporting
bym2_morans_table_report <- bym2_morans_table %>%
  dplyr::mutate(
    `Queens Moran's I` =
      sprintf("%.3f", `Queens Moran's I`),
    
    `Queens p-value` =
      format.pval(
        `Queens p-value`,
        digits = 3,
        eps = 0.001
      ),
    
    `KNN = 4 Moran's I` =
      sprintf("%.3f", `KNN = 4 Moran's I`),
    
    `KNN = 4 p-value` =
      format.pval(
        `KNN = 4 p-value`,
        digits = 3,
        eps = 0.001
      )
  )


# Display final table
knitr::kable(
  bym2_morans_table_report,
  align = c("l", "c", "c", "c", "c"),
  caption = "Moran's I tests of BYM2 residual spatial autocorrelation"
)
```

## 9. BiLISA - Bivariate LISA mapping

Well known method to address issues of spatial nonstationarity, allowing geographic variations in the relationships between dependent and explanatory variables –\> use `spdep` package

Each of the six dependent variables will be mapped with IMD and SUHII respectively:

```{r}
bilisa_variables <- spatial_model_variables %>%
  arrange(lsoa21cd)

# Create Queen-contiguity neighbours
bilisa_nb <- bilisa_variables %>%
  poly2nb(.,queen = T)

bilisa_weights <-  bilisa_nb %>%
  nb2listw(.,style = "W",)
```

### 9.1. Map SUHII

```{r}
# code adpated from: 
# https://r-spatial.github.io/spdep/reference/localmoran_bv.html

bilisa_15_pt_suhii <- localmoran_bv(
  x = bilisa_variables$access_score_15_pt,
  y = bilisa_variables$suhii_z,
  listw = bilisa_weights,
  nsim = 199,
  scale = TRUE,
  alternative = "two.sided",
  iseed = 42,
  zero.policy = TRUE
)

head(bilisa_15_pt_suhii)
```

```{r}
# to identify significant cluster only (keep p<0.05 only) 
bilisa_cluster <- spdep::hotspot(
  bilisa_15_pt_suhii,
  Prname = "Pr(folded) Sim",
  cutoff = 0.05,
  quadrant.type = "pysal",
  p.adjust = "none"
)
```

```{r}
bilisa_variables$bilisa_cluster <- as.character(
  bilisa_cluster
)

bilisa_variables$bilisa_cluster[
  is.na(bilisa_variables$bilisa_cluster)
] <- "Not significant"
```

```{r}
table(bilisa_variables$bilisa_cluster)
```

```{r}
tmap_mode("plot")

tm_shape(bilisa_variables) +
  tm_polygons(
    fill = "bilisa_cluster",
    fill.scale = tm_scale_categorical(),
    fill.legend = tm_legend(
      title = "BiLISA Cluster"
    ),
    col = NA
  ) +
  tm_title(
    "15-minute Public Transit Accessibility and SUHII"
  )
```

```{r}
# reorder clusters and make better labels
bilisa_variables <- bilisa_variables %>%
  mutate(
    bilisa_cluster_plot = case_when(
      bilisa_cluster == "Low-High" ~ "Low access–High SUHII",
      bilisa_cluster == "High-High" ~ "High access–High SUHII",
      bilisa_cluster == "Low-Low" ~ "Low access–Low SUHII",
      bilisa_cluster == "High-Low" ~ "High access–Low SUHII",
      TRUE ~ NA_character_
    )
  )

# set legend order
bilisa_variables$bilisa_cluster_plot <- factor(
  bilisa_variables$bilisa_cluster_plot,
  levels = c(
    "Low access–High SUHII",
    "High access–High SUHII",
    "Low access–Low SUHII",
    "High access–Low SUHII"
  )
)


cluster_cols <- c(
  "Low access–High SUHII"  = "#D73027",  # red
  "High access–High SUHII" = "#F4A582",  # light red / salmon
  "Low access–Low SUHII"   = "#91BFDB",  # light blue
  "High access–Low SUHII"  = "#4575B4"   # blue
)
```

```{r}
tmap_mode("plot")

bilisa_15_pt_suhii_suhii_map <-
  # Base: all LSOAs white
  tm_shape(bilisa_variables) +
  tm_fill(
    fill = "white"
  ) +

  # Significant clusters
  tm_shape(
    dplyr::filter(
      bilisa_variables,
      !is.na(bilisa_cluster_plot)
    )
  ) +
  tm_fill(
    fill = "bilisa_cluster_plot",
    fill.scale = tm_scale_categorical(
      values = cluster_cols
    ),
    fill.legend = tm_legend(
      title = "BiLISA cluster",
      position = tm_pos_out("right", "center"),
      frame = FALSE
    )
  ) +

  # London boundary must come last
  tm_shape(london_boundary) +
  tm_borders(
    col = "grey30",
    lwd = 1
  ) +

  tm_title(
    "15-minute PT accessibility\nand SUHII"
  ) +

  tm_layout(
    frame = FALSE,
    bg.color = "white",
    outer.margins = 0.02
  )

bilisa_15_pt_suhii_suhii_map 
```

**Loop over maps:**

```{r}
run_bilisa_map <- function(
    data,
    access_var,
    context_var,
    context_label,
    map_title,
    listw = bilisa_weights,
    boundary = london_boundary,
    nsim = 199,
    seed = 42,
    show_legend = TRUE
) {

  # Standardise only the accessibility variable
  access_z <- as.numeric(
    scale(data[[access_var]])
  )

  # SUHII or IMD is already a z-score
  context_z <- data[[context_var]]

  # Check that both variables and weights have matching lengths
  stopifnot(
    length(access_z) == length(context_z),
    length(access_z) == length(listw$neighbours)
  )

  # Bivariate Local Moran's I
  lisa_result <- spdep::localmoran_bv(
    x = access_z,
    y = context_z,
    listw = listw,
    nsim = nsim,
    scale = FALSE,
    alternative = "two.sided",
    iseed = seed,
    zero.policy = TRUE
  )

  # to identify significant cluster only (keep p<0.05 only) 
  lisa_cluster <- spdep::hotspot(
    lisa_result,
    Prname = "Pr(folded) Sim",
    cutoff = 0.05,
    quadrant.type = "pysal",
    p.adjust = "none"
  )

  # Full, readable cluster labels
  cluster_labels <- c(
    "Low-High" =
      paste0("Low access–High ", context_label),

    "High-High" =
      paste0("High access–High ", context_label),

    "Low-Low" =
      paste0("Low access–Low ", context_label),

    "High-Low" =
      paste0("High access–Low ", context_label)
  )

  cluster_levels <- unname(
    cluster_labels[
      c(
        "Low-High",
        "High-High",
        "Low-Low",
        "High-Low"
      )
    ]
  )

  cluster_colours <- stats::setNames(
    c(
      "#D73027",  # Low-High
      "#F4A582",  # High-High
      "#91BFDB",  # Low-Low
      "#4575B4"   # High-Low
    ),
    cluster_levels
  )

  # Create a separate result dataset
  result_data <- data %>%
    dplyr::mutate(
      bilisa_i = as.numeric(lisa_result[, 1]),

      bilisa_p = as.numeric(
        lisa_result[, "Pr(folded) Sim"]
      ),

      bilisa_cluster = as.character(
        lisa_cluster
      ),

      bilisa_cluster_plot = unname(
        cluster_labels[bilisa_cluster]
      ),

      bilisa_cluster_plot = factor(
        bilisa_cluster_plot,
        levels = cluster_levels
      )
    )

  # Significant areas only
  significant_data <- result_data %>%
    dplyr::filter(
      !is.na(bilisa_cluster_plot)
    )

  # Show or hide the legend
  legend_setting <- if (show_legend) {
    tm_legend(
      title = "BiLISA cluster",
      position = tm_pos_out("right", "center"),
      frame = FALSE
    )
  } else {
    tm_legend_hide()
  }

  # Create the map
  map_object <-
    tm_shape(result_data) +
    tm_fill(
      fill = "white"
    ) +

    tm_shape(significant_data) +
    tm_fill(
      fill = "bilisa_cluster_plot",

      fill.scale = tm_scale_categorical(
        values = cluster_colours
      ),

      fill.legend = legend_setting
    ) +

    # London boundary last
    tm_shape(boundary) +
    tm_borders(
      col = "grey30",
      lwd = 1
    ) +

    tm_title(
      map_title
    ) +

    tm_layout(
      frame = FALSE,
      bg.color = "white",
      outer.margins = 0.02
    )

  # Return everything
  list(
    result = lisa_result,
    clusters = lisa_cluster,
    data = result_data,
    map = map_object
  )
}
```

```{r}
bilisa_15_pt_suhii <- run_bilisa_map(
  data = bilisa_variables,
  access_var = "access_score_15_pt",
  context_var = "suhii_z",
  context_label = "SUHII",
  map_title = NULL,
  show_legend = FALSE
)

# assign as object
bilisa_15_pt_suhii_map <- bilisa_15_pt_suhii$map

# plot
#bilisa_15_pt_suhii_map
```

```{r}
bilisa_60_pt_suhii <- run_bilisa_map(
  data = bilisa_variables,
  access_var = "access_score_60_pt",
  context_var = "suhii_z",
  context_label = "SUHII",
  map_title = NULL,
  show_legend = FALSE
)

# assign as object
bilisa_60_pt_suhii_map <- bilisa_60_pt_suhii$map

# plot
#bilisa_60_pt_suhii_map
```

```{r}
bilisa_15_walk_suhii <- run_bilisa_map(
  data = bilisa_variables,
  access_var = "access_score_15_walk",
  context_var = "suhii_z",
  context_label = "SUHII",
  map_title = NULL,
  show_legend = FALSE
)

# assign as object
bilisa_15_walk_suhii_map <- bilisa_15_walk_suhii$map

# plot
#bilisa_15_walk_suhii_map
```

```{r}
bilisa_60_walk_suhii <- run_bilisa_map(
  data = bilisa_variables,
  access_var = "access_score_60_walk",
  context_var = "suhii_z",
  context_label = "SUHII",
  map_title = NULL,
  show_legend = FALSE
)

# assign as object
bilisa_60_walk_suhii_map <- bilisa_60_walk_suhii$map

# plot
#bilisa_60_walk_suhii_map
```

```{r}
bilisa_15_cycle_suhii <- run_bilisa_map(
  data = bilisa_variables,
  access_var = "access_score_15_cycle",
  context_var = "suhii_z",
  context_label = "SUHII",
  map_title = NULL,
  show_legend = FALSE
)

# assign as object
bilisa_15_cycle_suhii_map <- bilisa_15_cycle_suhii$map

# plot
#bilisa_15_cycle_suhii_map
```

```{r}
bilisa_60_cycle_suhii <- run_bilisa_map(
  data = bilisa_variables,
  access_var = "access_score_60_cycle",
  context_var = "suhii_z",
  context_label = "SUHII",
  map_title = NULL,
  show_legend = FALSE
)

# assign as object
bilisa_60_cycle_suhii_map <- bilisa_60_cycle_suhii$map

# plot
#bilisa_60_cycle_suhii_map
```

```{r}
bilisa_15_pt_imd <- run_bilisa_map(
  data = bilisa_variables,
  access_var = "access_score_15_pt",
  context_var = "imd_z",
  context_label = "IMD",
  map_title = NULL,
  show_legend = FALSE
)

# assign as object
bilisa_15_pt_imd_map <- bilisa_15_pt_imd$map

# plot
#bilisa_15_pt_imd_map
```

```{r}
bilisa_60_pt_imd <- run_bilisa_map(
  data = bilisa_variables,
  access_var = "access_score_60_pt",
  context_var = "imd_z",
  context_label = "IMD",
  map_title = NULL,
  show_legend = FALSE
)

# assign as object
bilisa_60_pt_imd_map <- bilisa_60_pt_imd$map

# plot
#bilisa_60_pt_imd_map
```

```{r}
bilisa_15_walk_imd <- run_bilisa_map(
  data = bilisa_variables,
  access_var = "access_score_15_walk",
  context_var = "imd_z",
  context_label = "IMD",
  map_title = NULL,
  show_legend = FALSE
)

# assign as object
bilisa_15_walk_imd_map <- bilisa_15_walk_imd$map

# plot
#bilisa_15_walk_imd_map
```

```{r}
bilisa_60_walk_imd <- run_bilisa_map(
  data = bilisa_variables,
  access_var = "access_score_60_walk",
  context_var = "imd_z",
  context_label = "IMD",
  map_title = NULL,
  show_legend = FALSE
)

# assign as object
bilisa_60_walk_imd_map <- bilisa_60_walk_imd$map

# plot
#bilisa_60_walk_imd_map
```

```{r}
bilisa_15_cycle_imd <- run_bilisa_map(
  data = bilisa_variables,
  access_var = "access_score_15_cycle",
  context_var = "imd_z",
  context_label = "IMD",
  map_title = NULL,
  show_legend = FALSE
)

# assign as object
bilisa_15_cycle_imd_map <- bilisa_15_cycle_imd$map

# plot
#bilisa_15_cycle_imd_map
```

```{r}
bilisa_60_cycle_imd <- run_bilisa_map(
  data = bilisa_variables,
  access_var = "access_score_60_cycle",
  context_var = "imd_z",
  context_label = "IMD",
  map_title = NULL,
  show_legend = FALSE
)

# assign as object
bilisa_60_cycle_imd_map <- bilisa_60_cycle_imd$map

# plot
#bilisa_60_cycle_imd_map
```

**Cowplot Combine Maps**

```{r}
bilisa_grobs <- list(
  tmap::tmap_grob(bilisa_15_pt_suhii_map),
  tmap::tmap_grob(bilisa_15_pt_imd_map),

  tmap::tmap_grob(bilisa_15_walk_suhii_map),
  tmap::tmap_grob(bilisa_15_walk_imd_map),

  tmap::tmap_grob(bilisa_15_cycle_suhii_map),
  tmap::tmap_grob(bilisa_15_cycle_imd_map)
)

bilisa_15_maps_combined <- cowplot::plot_grid(
  plotlist = bilisa_grobs,
  ncol = 2,
  labels = c("a", "b", "c", "d", "e", "f"),
  label_size = 12,
  align = "hv"
)

#bilisa_15_maps_combined
```

```{r}
# set up a shared legend for plotting purposes

legend_labels <- c(
  "Low access–High SUHII/IMD",
  "High access–High SUHII/IMD",
  "Low access–Low SUHII/IMD",
  "High access–Low SUHII/IMD"
)

legend_colours <- c(
  "Low access–High SUHII/IMD"  = "#D73027",
  "High access–High SUHII/IMD" = "#F4A582",
  "Low access–Low SUHII/IMD"   = "#91BFDB",
  "High access–Low SUHII/IMD"  = "#4575B4"
)

legend_data <- data.frame(
  cluster = factor(
    legend_labels,
    levels = legend_labels
  ),
  x = 1,
  y = 1:4
)

legend_plot <- ggplot(
  legend_data,
  aes(x = x, y = y, fill = cluster)
) +
  geom_point(
    shape = 22,
    size = 8,
    colour = NA
  ) +
  scale_fill_manual(
    values = legend_colours,
    breaks = legend_labels,
    drop = FALSE,
    name = "BiLISA clusters"
  ) +
  guides(
    fill = guide_legend(
      override.aes = list(
        shape = 22,
        size = 8,
        colour = NA
      )
    )
  ) +
  theme_void() +
  theme(
    legend.position = "right",
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 10),
    legend.key.height = grid::unit(0.65, "cm"),
    legend.key.width = grid::unit(0.65, "cm"),
    legend.background = element_blank()
  )

shared_legend <- cowplot::get_legend(
  legend_plot,
  legend = "right"
)
```

```{r}
# 15 min BiLISA maps
#bilisa_15_final <- cowplot::plot_grid(
#  bilisa_15_maps_combined,
#  shared_legend,
#  ncol = 2,
#  rel_widths = c(0.78, 0.22)
#)

#bilisa_15_final

# Combine the six maps with the shared legend
bilisa_15_body <- cowplot::plot_grid(
  bilisa_15_maps_combined,
  shared_legend,
  ncol = 2,
  rel_widths = c(0.82, 0.18)
)

# Create centred overall title
bilisa_15_title <- cowplot::ggdraw() +
  cowplot::draw_label(
    "15-minute Accessibility with SUHII and IMD",
    fontface = "bold",
    size = 14,
    hjust = 0.5,
    x = 0.5
  )

# Place the title above the maps
bilisa_15_final <- cowplot::plot_grid(
  bilisa_15_title,
  bilisa_15_body,
  ncol = 1,
  rel_heights = c(0.06, 0.94)
)

bilisa_15_final
```

```{r, eval= FALSE}
# save as png
ggsave(
  filename = "figures/FigD1_BiLISA_15_min_LH.png",
  plot = bilisa_15_final,
  width = 12.2,
  height = 9.5,
  units = "in",
  dpi = 300,
  bg = "white"
)
```

```{r}
# 60 min BiLISA maps
bilisa_grobs2 <- list(
  tmap::tmap_grob(bilisa_60_pt_suhii_map),
  tmap::tmap_grob(bilisa_60_pt_imd_map),

  tmap::tmap_grob(bilisa_60_walk_suhii_map),
  tmap::tmap_grob(bilisa_60_walk_imd_map),

  tmap::tmap_grob(bilisa_60_cycle_suhii_map),
  tmap::tmap_grob(bilisa_60_cycle_imd_map)
)

bilisa_60_maps_combined <- cowplot::plot_grid(
  plotlist = bilisa_grobs2,
  ncol = 2,
  labels = c("a", "b", "c", "d", "e", "f"),
  label_size = 12,
  align = "hv"
)

#bilisa_60_maps_combined
```

```{r}
#bilisa_60_final <- cowplot::plot_grid(
#  bilisa_60_maps_combined,
#  shared_legend,
#  ncol = 2,
#  rel_widths = c(0.78, 0.22)
#)

#bilisa_60_final


# Combine the six maps with the shared legend
bilisa_60_body <- cowplot::plot_grid(
  bilisa_60_maps_combined,
  shared_legend,
  ncol = 2,
  rel_widths = c(0.82, 0.18)
)

# Create centred overall title
bilisa_60_title <- cowplot::ggdraw() +
  cowplot::draw_label(
    "60-minute Accessibility with SUHII and IMD",
    fontface = "bold",
    size = 14,
    hjust = 0.5,
    x = 0.5
  )

# Place the title above the maps
bilisa_60_final <- cowplot::plot_grid(
  bilisa_60_title,
  bilisa_60_body,
  ncol = 1,
  rel_heights = c(0.06, 0.94)
)

bilisa_60_final
```

```{r, eval=FALSE}
ggsave(
  filename = "figures/FigD2_BiLISA_60_min_LH.png",
  plot = bilisa_60_final,
  width = 12.2,
  height = 9.5,
  units = "in",
  dpi = 300,
  bg = "white"
)
```

### 9.2. Distribution of BiLISA Cluster

```{r}
# Put the 12 existing BiLISA objects into a named list
bilisa_results <- list(
  "15-minute Public Transit and SUHII" = bilisa_15_pt_suhii,
  "15-minute Public Transit and IMD"   = bilisa_15_pt_imd,

  "15-minute Walking and SUHII" = bilisa_15_walk_suhii,
  "15-minute Walking and IMD"   = bilisa_15_walk_imd,

  "15-minute Cycling and SUHII" = bilisa_15_cycle_suhii,
  "15-minute Cycling and IMD"   = bilisa_15_cycle_imd,

  "60-minute Public Transit and SUHII" = bilisa_60_pt_suhii,
  "60-minute Public Transit and IMD"   = bilisa_60_pt_imd,

  "60-minute Walking and SUHII" = bilisa_60_walk_suhii,
  "60-minute Walking and IMD"   = bilisa_60_walk_imd,

  "60-minute Cycling and SUHII" = bilisa_60_cycle_suhii,
  "60-minute Cycling and IMD"   = bilisa_60_cycle_imd
)


# Extract cluster classifications from each result
bilisa_long <- bind_rows(
  lapply(
    names(bilisa_results),
    function(analysis_name) {

      cluster_values <- as.character(
        bilisa_results[[analysis_name]]$clusters
      )

      # hotspot() stores non-significant LSOAs as NA
      cluster_values[is.na(cluster_values)] <- "Not significant"

      tibble(
        Analysis = analysis_name,
        Cluster = cluster_values
      )
    }
  )
)
```

```{r}
analysis_order <- names(bilisa_results)

cluster_order <- c(
  "High-High",
  "High-Low",
  "Low-High",
  "Low-Low",
  "Not significant"
)

# Count LSOAs in each classification
bilisa_counts <- bilisa_long %>%
  dplyr::count(
    Analysis,
    Cluster,
    name = "n"
  ) %>%
  tidyr::complete(
    Analysis = analysis_order,
    Cluster = cluster_order,
    fill = list(n = 0)
  ) %>%
  dplyr::group_by(Analysis) %>%
  dplyr::mutate(
    total_lsoas = sum(n),
    percentage = (n / total_lsoas) * 100,

    table_value = sprintf(
      "%s (%.1f%%)",
      scales::comma(n),
      percentage
    )
  ) %>%
  dplyr::ungroup()

# Convert into the final 12-row table
bilisa_table <- bilisa_counts %>%
  dplyr::mutate(
    Analysis = factor(
      Analysis,
      levels = analysis_order
    ),

    Cluster = factor(
      Cluster,
      levels = cluster_order
    )
  ) %>%
  dplyr::arrange(
    Analysis,
    Cluster
  ) %>%
  dplyr::select(
    Analysis,
    Cluster,
    table_value
  ) %>%
  tidyr::pivot_wider(
    names_from = Cluster,
    values_from = table_value
  ) %>%
  dplyr::select(
    Analysis,
    `High-High`,
    `High-Low`,
    `Low-High`,
    `Low-Low`,
    `Not significant`
  ) %>%
  dplyr::mutate(
    Analysis = as.character(Analysis)
  )

bilisa_table
```

### 9.3. Low High BiLISA

```{r}
# Function to compare the Low access–High SUHII and
# Low access–High IMD clusters for the same transport mode
create_priority_two_combined <- function(suhii_object, imd_object) {

  suhii_object$data %>%
    dplyr::select(
      lsoa21cd,
      suhii_cluster = bilisa_cluster
    ) %>%

    # Join the matching IMD BiLISA classification by LSOA code
    dplyr::left_join(
      imd_object$data %>%
        sf::st_drop_geometry() %>%
        dplyr::select(
          lsoa21cd,
          imd_cluster = bilisa_cluster
        ),
      by = "lsoa21cd"
    ) %>%

    # Identify whether each LSOA belongs to either or both
    # Low access–High contextual-variable clusters
    dplyr::mutate(
      two_combined_class = dplyr::case_when(

        suhii_cluster == "Low-High" &
          imd_cluster == "Low-High" ~
          "High SUHII and IMD",

        suhii_cluster == "Low-High" ~
          "High SUHII",

        imd_cluster == "Low-High" ~
          "High IMD",

        TRUE ~ "Not Significant"
      ),

      # Set the legend order
      two_combined_class = factor(
        two_combined_class,
        levels = c(
          "High SUHII and IMD",
          "High SUHII",
          "High IMD",
          "Not Significant"
        )
      )
    )
}


# Public transport
two_combined_15_pt <- create_priority_two_combined(
  suhii_object = bilisa_15_pt_suhii,
  imd_object = bilisa_15_pt_imd
)

# Walking
two_combined_15_walk <- create_priority_two_combined(
  suhii_object = bilisa_15_walk_suhii,
  imd_object = bilisa_15_walk_imd
)

# Cycling
two_combined_15_cycle <- create_priority_two_combined(
  suhii_object = bilisa_15_cycle_suhii,
  imd_object = bilisa_15_cycle_imd
)
```

```{r}
# Colours for the two_combined classification
two_combined_colours <- c(
  "High SUHII and IMD" = "#5F2022",
  "High SUHII"         = "#AD423B",
  "High IMD"           = "#DC9179",
  "Not Significant"    = "#FBF8F2"
)


# Function to create one map with all combined
create_two_combined_map <- function(data) {

  tm_shape(data) +
    tm_fill(
      fill = "two_combined_class",
      fill.scale = tm_scale_categorical(
        values = two_combined_colours
      ),
      fill.legend = tm_legend_hide()
    ) +

    # Add Greater London boundary
    tm_shape(london_boundary) +
    tm_borders(
      col = "grey30",
      lwd = 1
    ) +

    tm_layout(
      frame = FALSE,
      bg.color = "white",
      outer.margins = 0
    )
}


# Create the three maps
two_combined_15_pt_map <- create_two_combined_map(
  two_combined_15_pt
)

two_combined_15_walk_map <- create_two_combined_map(
  two_combined_15_walk
)

two_combined_15_cycle_map <- create_two_combined_map(
  two_combined_15_cycle
)


# Convert tmap maps into grobs for cowplot
two_combined_15_grobs <- list(
  tmap::tmap_grob(two_combined_15_pt_map),
  tmap::tmap_grob(two_combined_15_walk_map),
  tmap::tmap_grob(two_combined_15_cycle_map)
)


# Combine the three maps
two_combined_15_maps <- cowplot::plot_grid(
  plotlist = two_combined_15_grobs,
  ncol = 1,
  labels = c("a", "b", "c"),
  label_size = 12,
  align = "v",
  axis = "lr"
)


# Create a shared legend
two_combined_legend_data <- data.frame(
  two_combined_class = factor(
    names(two_combined_colours),
    levels = names(two_combined_colours)
  ),
  x = 1,
  y = seq_along(two_combined_colours)
)

two_combined_legend_plot <- ggplot(
  two_combined_legend_data,
  aes(x = x, y = y, fill = two_combined_class)
) +
  geom_point(
    shape = 22,
    size = 8,
    colour = NA
  ) +
  scale_fill_manual(
    values = two_combined_colours,
    breaks = names(two_combined_colours),
    drop = FALSE,
    name = "BiLISA Clusters"
  ) +
  guides(
    fill = guide_legend(
      override.aes = list(
        shape = 22,
        size = 8,
        colour = NA
      )
    )
  ) +
  theme_void() +
  theme(
    legend.position = "right",
    legend.title = element_text(size = 11),
    legend.text = element_text(size = 9),
    legend.key.height = grid::unit(0.65, "cm"),
    legend.key.width = grid::unit(0.65, "cm")
  )


shared_two_combined_legend <- cowplot::get_legend(
  two_combined_legend_plot,
  legend = "right"
)


# Add shared legend
two_combined_15_final_body <- cowplot::plot_grid(
  two_combined_15_maps,
  shared_two_combined_legend,
  ncol = 2,
  rel_widths = c(0.76, 0.24)
)


# Add overall figure title
two_combined_15_final <- cowplot::ggdraw() +
  cowplot::draw_label(
    paste0(
      "15-minute BiLISA Clusters of\n",
      "Low Accessibility–High SUHII and/or High IMD"
    ),
    x = 0.5,
    y = 0.98,
    hjust = 0.5,
    fontface = "bold",
    size = 14
  ) +
  cowplot::draw_plot(
    two_combined_15_final_body,
    x = 0,
    y = 0,
    width = 1,
    height = 0.93
  )


# Display final figure
two_combined_15_final
```

```{r}
# save as PNG
ggsave(
  filename = "figures/Fig12_BiLISA_15_min_Combined.png",
  plot = two_combined_15_final,
  width = 9,
  height = 11,
  units = "in",
  dpi = 300,
  bg = "white"
)
```

```{r}
# Create the combined classifications

# Public transit
two_combined_60_pt <- create_priority_two_combined(
  suhii_object = bilisa_60_pt_suhii,
  imd_object = bilisa_60_pt_imd
)

# Walking
two_combined_60_walk <- create_priority_two_combined(
  suhii_object = bilisa_60_walk_suhii,
  imd_object = bilisa_60_walk_imd
)

# Cycling
two_combined_60_cycle <- create_priority_two_combined(
  suhii_object = bilisa_60_cycle_suhii,
  imd_object = bilisa_60_cycle_imd
)



# Create  three maps 
two_combined_60_pt_map <- create_two_combined_map(
  two_combined_60_pt
)

two_combined_60_walk_map <- create_two_combined_map(
  two_combined_60_walk
)

two_combined_60_cycle_map <- create_two_combined_map(
  two_combined_60_cycle
)


# Convert tmap objects into grobs for cowplot 
two_combined_60_grobs <- list(
  tmap::tmap_grob(two_combined_60_pt_map),
  tmap::tmap_grob(two_combined_60_walk_map),
  tmap::tmap_grob(two_combined_60_cycle_map)
)


# Arrange vertically:
# a = public transit
# b = walking
# c = cycling

two_combined_60_maps <- cowplot::plot_grid(
  plotlist = two_combined_60_grobs,
  ncol = 1,
  labels = c("a", "b", "c"),
  label_size = 12,
  align = "v",
  axis = "lr"
)


# Add the same shared legend  
two_combined_60_final_body <- cowplot::plot_grid(
  two_combined_60_maps,
  shared_two_combined_legend,
  ncol = 2,
  rel_widths = c(0.76, 0.24)
)


# Add the big title
two_combined_60_final <- cowplot::ggdraw() +
  cowplot::draw_label(
    paste0(
      "60-minute BiLISA Clusters of\n",
      "Low Accessibility–High SUHII and/or High IMD"
    ),
    x = 0.5,
    y = 0.98,
    hjust = 0.5,
    fontface = "bold",
    size = 14
  ) +
  cowplot::draw_plot(
    two_combined_60_final_body,
    x = 0,
    y = 0,
    width = 1,
    height = 0.93
  )


two_combined_60_final
```

```{r, eval=FALSE}
# save as PNG
ggsave(
  filename = "figures/Fig14_BiLISA_60_min_LH_Combined.png",
  plot = two_combined_60_final,
  width = 9,
  height = 11,
  units = "in",
  dpi = 300,
  bg = "white"
)
```

### 9.4. Summary of overlap L-H clusters of High SUHII and IMD

```{r}
# Put all six overlap objects into a named list
overlap_objects <- list(
  "15-minute Public Transit" = two_combined_15_pt,
  "15-minute Walking"          = two_combined_15_walk,
  "15-minute Cycling"          = two_combined_15_cycle,
  "60-minute Public Transit" = two_combined_60_pt,
  "60-minute Walking"          = two_combined_60_walk,
  "60-minute Cycling"          = two_combined_60_cycle
)

# Calculate number and percentage of London LSOAs
# that are Low-High for BOTH SUHII and IMD
overlap_summary <- dplyr::bind_rows(
  
  lapply(names(overlap_objects), function(x) {
    
    data <- overlap_objects[[x]] %>%
      sf::st_drop_geometry()
    
    tibble::tibble(
      Accessibility = x,
      
      overlap_n = sum(
        data$two_combined_class == "High SUHII and IMD",
        na.rm = TRUE
      ),
      
      total_lsoas = nrow(data),
      
      overlap_percentage = 
        overlap_n / total_lsoas * 100
    )
    
  })
)

overlap_summary
```
